---
title: "Basic_outputs_plots"
author: "Dries Lorré"
date: "2026-03-11"
output: html_document
---

In this script, the species-specific biomass, species-specific catch and fleet-specific effort maps are processed and plotted.

Packages
```{r}
library(terra)
library(dplyr)
library(stringr)
library(purrr)
library(ggplot2)
library(tidyr)
```
Main directories
```{r}
scenario_dirs <- c(
  Baseline  = "C:/Users/dlorr/OneDrive - UGent/Documents/EwE output/SBNS_CC_adapted_spatial_baseline_test",
  CC        = "C:/Users/dlorr/OneDrive - UGent/Documents/EwE output/SBNS_CC_adapted_spatial_CC_test",
  OWF       = "C:/Users/dlorr/OneDrive - UGent/Documents/EwE output/SBNS_CC_adapted_spatial_OWF_test",
  OWF_CC    = "C:/Users/dlorr/OneDrive - UGent/Documents/EwE output/SBNS_CC_adapted_spatial_OWF_CC_test"
)

# Folder name inside each scenario
asc_subfolder <- file.path("Test_Env_resps", "asc")

# Output folder
out_dir <- file.path(getwd(), "Basic_output_plots")

if (!dir.exists(out_dir)) {
  dir.create(out_dir, recursive = TRUE)
}

cat("Output figures will be saved in:\n", out_dir, "\n")
```

Keep time steps for last 10 years
```{r}
# Last timestep = 1920
# Yearly maps are every 12 steps
# Last 10 years = 1812, 1824, ..., 1920
target_steps <- seq(1812, 1920, by = 12)
target_steps_chr <- sprintf("%05d", target_steps)
```

Definition of the metrics
```{r}
metrics <- list(
  biomass = list(
    prefix = "EcospaceMapBiomass-",
    type   = "group"
  ),
  catch = list(
    prefix = "EcospaceMapCatch-",
    type   = "group"
  ),
  discards = list(
    prefix = "EcospaceMapDiscards-",
    type   = "group"
  ),
  habitat_capacity = list(
    prefix = "LayerHabitatCapacity-",
    type   = "group"
  ),
  effort = list(
    prefix = "EcospaceMapEffort-",
    type   = "fleet"
  )
)
```

Helper functions
```{r}
# Extract timestep from filename
extract_timestep <- function(x) {
  as.integer(str_extract(basename(x), "(?<=-)\\d{5}(?=\\.asc$)"))
}

# Extract group/fleet name from filename:
# removes prefix and trailing "-00012.asc"
extract_entity_name <- function(x, prefix) {
  nm <- basename(x)
  nm <- str_remove(nm, paste0("^", stringr::fixed(prefix)))
  nm <- str_remove(nm, "-\\d{5}\\.asc$")
  nm
}

# Build full asc folder path for a scenario
get_asc_dir <- function(scenario_path) {
  file.path(scenario_path, asc_subfolder)
}

# Read and average rasters for a single group/fleet over selected timesteps
average_entity_rasters <- function(files) {
  r_list <- lapply(files, rast)
  r_stack <- rast(r_list)
  app(r_stack, fun = mean, na.rm = TRUE)
}

# Convert raster to dataframe for ggplot
raster_to_df <- function(r, entity_name, scenario_name, metric_name) {
  df <- as.data.frame(r, xy = TRUE, na.rm = FALSE)
  names(df)[3] <- "value"
  df$entity <- entity_name
  df$scenario <- scenario_name
  df$metric <- metric_name
  df
}

# Process one metric for one scenario
process_metric_scenario <- function(scenario_name, scenario_path, metric_name, metric_info) {
  
  asc_dir <- get_asc_dir(scenario_path)
  
  if (!dir.exists(asc_dir)) {
    warning("ASC directory not found for scenario: ", scenario_name, " -> ", asc_dir)
    return(NULL)
  }
  
  all_files <- list.files(asc_dir, pattern = "\\.asc$", full.names = TRUE)
  
  # Keep only files for this metric
  metric_files <- all_files[str_detect(basename(all_files), paste0("^", fixed(metric_info$prefix)))]
  
  if (length(metric_files) == 0) {
    warning("No files found for metric '", metric_name, "' in scenario: ", scenario_name)
    return(NULL)
  }
  
  # Extract timestep and keep only last 10 years
  file_info <- tibble(
    file = metric_files,
    timestep = extract_timestep(metric_files),
    entity = extract_entity_name(metric_files, metric_info$prefix)
  ) %>%
    filter(timestep %in% target_steps)
  
  if (nrow(file_info) == 0) {
    warning("No matching last-10-year files found for metric '", metric_name,
            "' in scenario: ", scenario_name)
    return(NULL)
  }
  
  # Check whether entities have complete 10-year series
  completeness <- file_info %>%
    count(entity, name = "n_files") %>%
    mutate(complete = n_files == length(target_steps))
  
  incomplete_entities <- completeness %>%
    filter(!complete)
  
  if (nrow(incomplete_entities) > 0) {
    message("Scenario: ", scenario_name,
            " | Metric: ", metric_name,
            " | Some entities do not have all 10 target timesteps.")
    print(incomplete_entities)
  }
  
  # Average raster for each entity
  entity_dfs <- file_info %>%
    arrange(entity, timestep) %>%
    group_split(entity) %>%
    lapply(function(df_ent) {
      ent_name <- unique(df_ent$entity)
      message("Processing ", scenario_name, " | ", metric_name, " | ", ent_name)
      
      r_mean <- average_entity_rasters(df_ent$file)
      raster_to_df(r_mean, ent_name, scenario_name, metric_name)
    })
  
  bind_rows(entity_dfs)
}

# Function to make a faceted plot
plot_metric_maps <- function(df_metric, scenario_name, metric_name, out_dir,
                             ncol = NULL, width = 16, height = 10) {
  
  if (is.null(df_metric) || nrow(df_metric) == 0) {
    warning("No data to plot for ", scenario_name, " - ", metric_name)
    return(NULL)
  }
  
  n_entities <- dplyr::n_distinct(df_metric$entity)
  
  if (is.null(ncol)) {
    ncol <- ceiling(sqrt(n_entities))
  }
  
  if (scenario_name == "Baseline") {
    
    p <- ggplot(df_metric, aes(x = x, y = y, fill = plot_value)) +
      geom_raster() +
      coord_equal() +
      facet_wrap(~ entity, ncol = ncol) +
      scale_fill_viridis_c(
        option = "magma",
        na.value = "grey90"
      ) +
      labs(
        title = paste0(str_to_title(gsub("_", " ", metric_name)),
                       " maps (mean of last 10 years)"),
        subtitle = "Baseline scenario | absolute values",
        x = "Longitude",
        y = "Latitude",
        fill = "Mean value"
      ) +
      theme_bw() +
      theme(
        plot.title = element_text(face = "bold", size = 14),
        plot.subtitle = element_text(size = 10),
        strip.text = element_text(size = 8),
        panel.grid = element_blank(),
        legend.position = "right"
      )
    
  } else {
    
    p <- ggplot(df_metric, aes(x = x, y = y, fill = plot_value)) +
      geom_raster() +
      coord_equal() +
      facet_wrap(~ entity, ncol = ncol) +
      scale_fill_gradient2(
        low = "steelblue",
        mid = "white",
        high = "firebrick",
        midpoint = 0,
        na.value = "grey90",
        labels = function(x) paste0(x, "%")
      ) +
      labs(
        title = paste0(str_to_title(gsub("_", " ", metric_name)),
                       " maps (mean of last 10 years)"),
        subtitle = paste0("Scenario: ", scenario_name, " | % change relative to Baseline"),
        x = "Longitude",
        y = "Latitude",
        fill = "% change"
      ) +
      theme_bw() +
      theme(
        plot.title = element_text(face = "bold", size = 14),
        plot.subtitle = element_text(size = 10),
        strip.text = element_text(size = 8),
        panel.grid = element_blank(),
        legend.position = "right"
      )
  }
  
  print(p)
  
  out_file <- file.path(
    out_dir,
    paste0("MapGrid_", metric_name, "_", scenario_name, "_last10yrs_mean.png")
  )
  
  ggsave(out_file, p, width = width, height = height, dpi = 300)
  
  return(p)
}
```

Process data for all scenrio's
```{r}
all_results <- list()

for (sc_name in names(scenario_dirs)) {
  sc_path <- scenario_dirs[[sc_name]]
  
  message("====================================================")
  message("SCENARIO: ", sc_name)
  message("====================================================")
  
  all_results[[sc_name]] <- list()
  
  for (met_name in names(metrics)) {
    message("---- Metric: ", met_name)
    
    res <- process_metric_scenario(
      scenario_name = sc_name,
      scenario_path = sc_path,
      metric_name   = met_name,
      metric_info   = metrics[[met_name]]
    )
    
    all_results[[sc_name]][[met_name]] <- res
  }
}

```

Baseline Plotting:
```{r}
library(patchwork)
library(scales)
```


```{r}
# Flatten all results
all_df <- bind_rows(
  lapply(all_results, function(x) bind_rows(x))
)

# Keep only baseline
baseline_df <- all_df %>%
  filter(scenario == "Baseline")

baseline_habcap_df <- baseline_df %>%
  filter(metric == "habitat_capacity")

# Read in study area land mass for plotting
library(sf)

land_main <- readRDS(
  "land_main_study_area.rds"
)
```

Helper functions
```{r}
#-----------------------------------------------------------
# Prepare baseline data for one metric
#-----------------------------------------------------------
prepare_baseline_metric <- function(df, metric_name) {
  
  df %>%
    filter(metric == metric_name) %>%
    group_by(entity) %>%
    mutate(
      mean_abs = mean(value, na.rm = TRUE),
      rel_value = ifelse(!is.na(mean_abs) & mean_abs != 0, value / mean_abs, NA_real_),
      rel_min = min(rel_value, na.rm = TRUE),
      rel_max = max(rel_value, na.rm = TRUE)
    ) %>%
    ungroup()
}


#-----------------------------------------------------------
# Build one subplot for one entity
#-----------------------------------------------------------
# make_baseline_subplot <- function(df_entity, metric_name,
#                                   show_x = FALSE,
#                                   show_y = FALSE) {
#   
#   ent_name <- unique(df_entity$entity)
#   mean_abs <- unique(df_entity$mean_abs)
#   rel_min  <- unique(df_entity$rel_min)
#   rel_max  <- unique(df_entity$rel_max)
#   
#   # safeguards
#   if (length(mean_abs) > 1) mean_abs <- mean(mean_abs, na.rm = TRUE)
#   if (length(rel_min)  > 1) rel_min  <- min(rel_min, na.rm = TRUE)
#   if (length(rel_max)  > 1) rel_max  <- max(rel_max, na.rm = TRUE)
#   
#   # sparse axis breaks
#   x_breaks <- pretty(range(df_entity$x, na.rm = TRUE), n = 3)
#   y_breaks <- pretty(range(df_entity$y, na.rm = TRUE), n = 3)
#   
#   p <- ggplot(df_entity, aes(x = x, y = y, fill = rel_value)) +
#     geom_raster() +
#     coord_equal() +
#     scale_x_continuous(breaks = x_breaks) +
#     scale_y_continuous(breaks = y_breaks) +
#     scale_fill_distiller(
#       palette = "Spectral",
#       direction = -1,
#       limits = c(rel_min, rel_max),
#       breaks = pretty(c(rel_min, rel_max), n = 3),
#       labels = label_number(accuracy = 0.1),
#       na.value = "grey80",
#       guide = guide_colorbar(
#         title.position = "top",
#         barheight = unit(10, "mm"),
#         barwidth  = unit(4, "mm"),
#         frame.colour = "black",
#         ticks.colour = "black"
#       )
#     ) +
#     labs(
#       title = ent_name,
#       subtitle = paste0("Mean = ", signif(mean_abs, 4)),
#       x = if (show_x) "Longitude" else NULL,
#       y = if (show_y) "Latitude" else NULL,
#       fill = "Relative\nvalue"
#     ) +
#     theme_bw(base_size = 11) +
#     theme(
#       plot.title = element_text(face = "bold", size = 13),
#       plot.subtitle = element_text(size = 13),
#       axis.title = element_text(size = 13),
#       axis.text = element_text(size = 13),
#       legend.title = element_text(size = 12),
#       legend.text = element_text(size = 12),
#       panel.grid = element_blank(),
#       
#       # place legend INSIDE panel
#       legend.position = c(0.99, 0.02),
#       legend.justification = c(1, 0),
#       
#       # make legend readable but compact
#       legend.direction = "vertical",
#       legend.box.background = element_rect(
#         colour = "black",
#         fill = scales::alpha("white", 0.75)
#       ),
#       legend.background = element_rect(
#         fill = scales::alpha("white", 0.75),
#         colour = NA
#       ),
#       legend.margin = margin(2, 2, 2, 2),
#       legend.key.height = unit(0.45, "cm"),
#       legend.key.width  = unit(0.30, "cm")
#     )
#   
#   if (!show_x) {
#     p <- p + theme(
#       axis.text.x = element_blank(),
#       axis.ticks.x = element_blank()
#     )
#   }
#   
#   if (!show_y) {
#     p <- p + theme(
#       axis.text.y = element_blank(),
#       axis.ticks.y = element_blank()
#     )
#   }
#   
#   return(p)
# }

# --NEW-------------------------------------------------------------------
make_baseline_subplot <- function(df_entity, metric_name,
                                  show_x = FALSE,
                                  show_y = FALSE) {
  
  ent_name <- unique(df_entity$entity)
  mean_abs <- unique(df_entity$mean_abs)
  
  # safeguards
  if (length(mean_abs) > 1) mean_abs <- mean(mean_abs, na.rm = TRUE)
  
  # robust panel-specific limits based on rel_value
  panel_min <- quantile(df_entity$rel_value, probs = 0.01, na.rm = TRUE)
  panel_max <- quantile(df_entity$rel_value, probs = 0.99, na.rm = TRUE)
  
  if (!is.finite(panel_min)) panel_min <- min(df_entity$rel_value, na.rm = TRUE)
  if (!is.finite(panel_max)) panel_max <- max(df_entity$rel_value, na.rm = TRUE)
  
  if (panel_min == panel_max) {
    panel_max <- panel_min + 1e-6
  }
  
  # sparse axis breaks
  x_breaks <- pretty(range(df_entity$x, na.rm = TRUE), n = 3)
  y_breaks <- pretty(range(df_entity$y, na.rm = TRUE), n = 3)
  
  p <- ggplot(df_entity, aes(x = x, y = y, fill = rel_value)) +
    geom_raster() +
    geom_sf(
      data = land_main,
      inherit.aes = FALSE,
      fill = "grey85",
      colour = "grey45",
      linewidth = 0.2
      ) +
    coord_sf(
      xlim = range(df_entity$x, na.rm = TRUE),
      ylim = range(df_entity$y, na.rm = TRUE),
      expand = FALSE
      ) +
    scale_x_continuous(breaks = x_breaks) +
    scale_y_continuous(breaks = y_breaks) +
    scale_fill_distiller(
      palette = "Spectral",
      direction = -1,
      limits = c(panel_min, panel_max),
      oob = scales::squish,
      breaks = pretty(c(panel_min, panel_max), n = 3),
      labels = label_number(accuracy = 0.1),
      na.value = "white",
      guide = guide_colorbar(
        title.position = "top",
        barheight = unit(10, "mm"),
        barwidth  = unit(4, "mm"),
        frame.colour = "black",
        ticks.colour = "black"
      )
    ) +
    labs(
      title = ent_name,
      subtitle = paste0("Mean = ", signif(mean_abs, 4)),
      x = if (show_x) "Longitude" else NULL,
      y = if (show_y) "Latitude" else NULL,
      fill = "Relative\nvalue"
    ) +
    theme_bw(base_size = 11) +
    theme(
      plot.title = element_text(face = "bold", size = 13),
      plot.subtitle = element_text(size = 13),
      axis.title = element_text(size = 13),
      axis.text = element_text(size = 13),
      legend.title = element_text(size = 12),
      legend.text = element_text(size = 12),
      panel.grid = element_blank(),
      
      # place legend INSIDE panel
      legend.position = c(0.99, 0.02),
      legend.justification = c(1, 0),
      
      # make legend readable but compact
      legend.direction = "vertical",
      legend.box.background = element_rect(
        colour = "black",
        fill = scales::alpha("white", 0.75)
      ),
      legend.background = element_rect(
        fill = scales::alpha("white", 0.75),
        colour = NA
      ),
      legend.margin = margin(2, 2, 2, 2),
      legend.key.height = unit(0.45, "cm"),
      legend.key.width  = unit(0.30, "cm")
    )
  
  if (!show_x) {
    p <- p + theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank()
    )
  }
  
  if (!show_y) {
    p <- p + theme(
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    )
  }
  
  return(p)
}
#-----------------------------------------------------------
# Build one habitat capacity subplot for one entity
#-----------------------------------------------------------
make_baseline_habcap_subplot <- function(df_entity,
                                         show_x = FALSE,
                                         show_y = FALSE) {
  
  ent_name <- unique(df_entity$entity)
  
  # sparse axis breaks
  x_breaks <- pretty(range(df_entity$x, na.rm = TRUE), n = 3)
  y_breaks <- pretty(range(df_entity$y, na.rm = TRUE), n = 3)
  
  p <- ggplot(df_entity, aes(x = x, y = y, fill = value)) +
    geom_raster() +
    geom_sf(
      data = land_main,
      inherit.aes = FALSE,
      fill = "grey85",
      colour = "grey45",
      linewidth = 0.2
      ) +
    coord_sf(
      xlim = range(df_entity$x, na.rm = TRUE),
      ylim = range(df_entity$y, na.rm = TRUE),
      expand = FALSE
      ) +
    scale_x_continuous(breaks = x_breaks) +
    scale_y_continuous(breaks = y_breaks) +
    # scale_fill_gradient(
    #   low = "white",
    #   high = "#1B5F6A",
    #   limits=c(0,1.2),
    #   breaks = c(0, 1.2),
    #   labels = c("LOW", "HIGH"),
    scale_fill_gradientn(
  colours = (RColorBrewer::brewer.pal(11, "Spectral")[6:11]),
  limits = c(0, 1.2),
  breaks = c(0, 1.2),
  labels = c("LOW", "HIGH"),
      # labels = label_number(accuracy = 0.1),
      na.value = "white",
      guide = guide_colorbar(
        title.position = "top",
        title.hjust = 0.5,
        direction = "horizontal",
        barheight = unit(4, "mm"),
        barwidth  = unit(70, "mm"),
        frame.colour = "black",
        ticks.colour = "black"
      )
    ) +
    labs(
      title = ent_name,
      x = if (show_x) "Longitude" else NULL,
      y = if (show_y) "Latitude" else NULL,
      fill = "Habitat capacity"
    ) +
    theme_bw(base_size = 11) +
    theme(
      plot.title = element_text(face = "bold", size = 13),
      axis.title = element_text(size = 13),
      axis.text = element_text(size = 13),
      legend.title = element_text(size = 12),
      legend.text = element_text(size = 12),
      panel.grid = element_blank()
    )
  
  if (!show_x) {
    p <- p + theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank()
    )
  }
  
  if (!show_y) {
    p <- p + theme(
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    )
  }
  
  return(p)
}

#-----------------------------------------------------------
# Split vector into pages of max n items
#-----------------------------------------------------------
split_into_pages <- function(x, n = 20) {
  split(x, ceiling(seq_along(x) / n))
}


#-----------------------------------------------------------
# Make paginated baseline plots for one metric
#-----------------------------------------------------------
plot_baseline_metric_pages <- function(df_metric, metric_name, out_dir,
                                       ncol = 4, nrow = 5,
                                       width = 15.5, height = 18) {
  
  entities <- unique(df_metric$entity)
  entity_pages <- split_into_pages(entities, n = ncol * nrow)
  
  page_plots <- list()
  
  for (i in seq_along(entity_pages)) {
    
    ents_this_page <- entity_pages[[i]]
    n_panels <- length(ents_this_page)
    n_rows_used <- ceiling(n_panels / ncol)
    
    subplot_list <- list()
    
    for (j in seq_along(ents_this_page)) {
      
      ent <- ents_this_page[j]
      df_ent <- df_metric %>% filter(entity == ent)
      
      row_pos <- ceiling(j / ncol)
      col_pos <- j %% ncol
      if (col_pos == 0) col_pos <- ncol
      
      show_x <- row_pos == n_rows_used
      show_y <- col_pos == 1
      
      subplot_list[[j]] <- make_baseline_subplot(
        df_ent,
        metric_name,
        show_x = show_x,
        show_y = show_y
      )
    }
    
    combined_plot <- wrap_plots(subplot_list, ncol = ncol, nrow = n_rows_used)
    
    page_title <- paste0(
      str_to_title(gsub("_", " ", metric_name)),
      " | Baseline spatial distribution (relative to layer mean) | Plot ",
      i, "/", length(entity_pages)
    )
    
    combined_plot <- combined_plot +
      plot_annotation(title = page_title)
    
    print(combined_plot)
    
    out_file <- file.path(
      out_dir,
      paste0("Baseline_", metric_name, "_relative_distribution_page_", i, ".png")
    )
    
    ggsave(out_file, combined_plot, width = width, height = height, dpi = 800)
    
    page_plots[[i]] <- combined_plot
  }
  
  return(page_plots)
}

#-----------------------------------------------------------
# Make paginated baseline habitat capacity plots
# with shared horizontal legend at bottom
#-----------------------------------------------------------
plot_baseline_habcap_pages <- function(df_metric, out_dir,
                                       ncol = 4, nrow = 5,
                                       width = 15.5, height = 18) {
  
  entities <- unique(df_metric$entity)
  entity_pages <- split_into_pages(entities, n = ncol * nrow)
  
  page_plots <- list()
  
  for (i in seq_along(entity_pages)) {
    
    ents_this_page <- entity_pages[[i]]
    
    subplot_list <- list()
    
    n_panels <- length(ents_this_page)
    n_rows_used <- ceiling(n_panels / ncol)
    
    for (j in seq_along(ents_this_page)) {
      
      ent <- ents_this_page[j]
      df_ent <- df_metric %>% filter(entity == ent)
      
      row_pos <- ceiling(j / ncol)
      col_pos <- j %% ncol
      if (col_pos == 0) col_pos <- ncol
      
      show_x <- row_pos == n_rows_used
      show_y <- col_pos == 1
      
      subplot_list[[j]] <- make_baseline_habcap_subplot(
        df_ent,
        show_x = show_x,
        show_y = show_y
      )
    }
    
    n_panels <- length(ents_this_page)
n_rows_used <- ceiling(n_panels / ncol)

combined_plot <- wrap_plots(subplot_list, ncol = ncol, nrow = n_rows_used, guides = "collect") &
      theme(
        legend.position = "bottom",
        legend.direction = "horizontal"
      )
    
    page_title <- paste0(
      "Habitat Capacity | Baseline | Plot ",
      i, "/", length(entity_pages)
    )
    
    combined_plot <- combined_plot +
      plot_annotation(title = page_title)
    
    print(combined_plot)
    
    out_file <- file.path(
      out_dir,
      paste0("Baseline_habitat_capacity_page_", i, ".png")
    )
    
    ggsave(out_file, combined_plot, width = width, height = height, dpi = 800)
    
    page_plots[[i]] <- combined_plot
  }
  
  return(page_plots)
}
```

Plot Baseline metrics
```{r}
############################################################
# MAKE BASELINE PLOTS
############################################################

# Plot for Biomass, Catch, Discards and Effort
baseline_plot_list <- list()

metrics_baseline <- c("biomass", "catch", "discards", "effort")
# add "habitat_capacity" here too if you want it in the same style:
# metrics_baseline <- c("biomass", "catch", "discards", "effort", "habitat_capacity")

for (met_name in metrics_baseline) {
  
  message("====================================================")
  message("BASELINE PLOTS FOR: ", met_name)
  message("====================================================")
  
  df_metric <- prepare_baseline_metric(baseline_df, met_name)
  
  baseline_plot_list[[met_name]] <- plot_baseline_metric_pages(
    df_metric   = df_metric,
    metric_name = met_name,
    out_dir     = out_dir,
    ncol        = 4,
    nrow        = 5
  )
}

# Habitat Capacity plot
baseline_habcap_plots <- plot_baseline_habcap_pages(
  df_metric = baseline_habcap_df,
  out_dir   = out_dir,
  ncol      = 4,
  nrow      = 5,
  width     = 15.5,
  height    = 18
)
```





Build plot dataframe with baseline absolute values and % change for other scenarios
```{r}
all_df <- bind_rows(
  lapply(all_results, function(x) bind_rows(x))
)

baseline_ref <- all_df %>%
  filter(scenario == "Baseline") %>%
  select(metric, entity, x, y, baseline_value = value)

plot_df <- all_df %>%
  left_join(baseline_ref, by = c("metric", "entity", "x", "y")) %>%
  mutate(
    plot_value = case_when(
      scenario == "Baseline" ~ value,
      scenario != "Baseline" & !is.na(baseline_value) & baseline_value != 0 ~
        ((value - baseline_value) / baseline_value) * 100,
      TRUE ~ NA_real_
    )
  )

# Optional diagnostic
plot_df %>%
  group_by(scenario, metric) %>%
  summarise(
    n_cells = n(),
    n_na_plot = sum(is.na(plot_value)),
    .groups = "drop"
  )
```

Plots
```{r}
# plot_list <- list()
# 
# for (sc_name in unique(plot_df$scenario)) {
#   
#   plot_list[[sc_name]] <- list()
#   
#   for (met_name in unique(plot_df$metric)) {
#     
#     df_sub <- plot_df %>%
#       filter(scenario == sc_name, metric == met_name)
#     
#     p <- plot_metric_maps(
#       df_metric     = df_sub,
#       scenario_name = sc_name,
#       metric_name   = met_name,
#       out_dir       = out_dir
#     )
#     
#     plot_list[[sc_name]][[met_name]] <- p
#   }
# }
```

SCENARIO plots
```{r}
library(patchwork)
library(RColorBrewer)
library(scales)
```


```{r}
############################################################
# SCENARIO PLOTS
# % change relative to baseline
############################################################

# Flatten all processed results
all_df <- bind_rows(
  lapply(all_results, function(x) bind_rows(x))
)

# Baseline reference per metric/entity/grid cell
baseline_ref <- all_df %>%
  filter(scenario == "Baseline") %>%
  select(metric, entity, x, y, baseline_value = value)

# Scenario data with % change vs baseline
scenario_plot_df <- all_df %>%
  filter(scenario != "Baseline") %>%
  left_join(baseline_ref, by = c("metric", "entity", "x", "y")) %>%
  mutate(
    pct_change = case_when(
      baseline_value == 0 & value == 0 ~ 0,
      baseline_value == 0 & value != 0 ~ NA_real_,
      !is.na(baseline_value) ~ ((value - baseline_value) / baseline_value) * 100,
      TRUE ~ NA_real_
    ),
    scenario = factor(scenario, levels = c("CC", "OWF", "OWF_CC"),
                      labels = c("CC", "OWF", "OWF + CC"))
  )
```

Helper function to prepare one metric
```{r}
prepare_scenario_metric <- function(df, metric_name, cap_percent = 100) {
  
  df_metric <- df %>%
    filter(metric == metric_name)
  
  max_abs_change <- df_metric %>%
    summarise(max_abs = max(abs(pct_change), na.rm = TRUE)) %>%
    pull(max_abs)
  
  if (!is.finite(max_abs_change)) max_abs_change <- cap_percent
  
  max_abs_change <- min(max_abs_change, cap_percent)
  
  df_metric %>%
    mutate(
      pct_change_plot = pmin(pmax(pct_change, -max_abs_change), max_abs_change),
      scale_limit = max_abs_change
    )
}
```

Subplot function
```{r}
# make_scenario_subplot <- function(df_panel,
#                                   show_x = FALSE,
#                                   show_y = FALSE,
#                                   show_title = TRUE) {
#   
#   scen_name <- unique(df_panel$scenario)
#   
#   # actual min and max per panel
#   panel_min <- min(df_panel$pct_change, na.rm = TRUE)
#   panel_max <- max(df_panel$pct_change, na.rm = TRUE)
#   
#   # safeguard
#   if (!is.finite(panel_min)) panel_min <- -1
#   if (!is.finite(panel_max)) panel_max <- 1
#   
#   # avoid zero-width scale
#   if (panel_min == panel_max) {
#     panel_min <- panel_min - 1
#     panel_max <- panel_max + 1
#   }
#   
#   x_breaks <- pretty(range(df_panel$x, na.rm = TRUE), n = 3)
#   y_breaks <- pretty(range(df_panel$y, na.rm = TRUE), n = 3)
#   
#   p <- ggplot(df_panel, aes(x = x, y = y, fill = pct_change)) +
#     geom_raster() +
#     coord_equal() +
#     scale_x_continuous(breaks = x_breaks) +
#     scale_y_continuous(breaks = y_breaks) +
#     scale_fill_gradient2(
#       low = "#E67E22",    # orange for decrease
#       mid = "white",
#       high = "#2C7FB8",   # blue for increase
#       midpoint = 0,
#       limits = c(panel_min, panel_max),
#       breaks = pretty(c(panel_min, panel_max), n = 3),
#       labels = function(x) paste0(round(x, 1), "%"),
#       na.value = "grey85",
#       guide = guide_colorbar(
#         title.position = "top",
#         barheight = unit(13, "mm"),
#         barwidth  = unit(4.5, "mm"),
#         frame.colour = "black",
#         ticks.colour = "black"
#       )
#     ) +
#     labs(
#       title = if (show_title) as.character(scen_name) else NULL,
#       x = if (show_x) "Longitude" else NULL,
#       y = if (show_y) "Latitude" else NULL,
#       fill = "% change"
#     ) +
#     theme_bw(base_size = 11) +
#     theme(
#       plot.title = element_text(face = "bold", size = 12, hjust = 0.5),
#       axis.title = element_text(size = 12),
#       axis.text = element_text(size = 12),
#       legend.title = element_text(size = 11),
#       legend.text = element_text(size = 11),
#       panel.grid = element_blank(),
#       
#       # legend inside panel
#       legend.position = c(0.99, 0.02),
#       legend.justification = c(1, 0),
#       
#       legend.box.background = element_rect(
#         colour = "black",
#         fill = scales::alpha("white", 0.75)
#       ),
#       legend.background = element_rect(
#         fill = scales::alpha("white", 0.75),
#         colour = NA
#       ),
#       legend.margin = margin(2, 2, 2, 2),
#       legend.key.height = unit(0.4, "cm"),
#       legend.key.width  = unit(0.25, "cm")
#     )
#   
#   if (!show_x) {
#     p <- p + theme(
#       axis.text.x = element_blank(),
#       axis.ticks.x = element_blank()
#     )
#   }
#   
#   if (!show_y) {
#     p <- p + theme(
#       axis.text.y = element_blank(),
#       axis.ticks.y = element_blank()
#     )
#   }
#   
#   p
# }
# 

# NEW ----------------------------------------------------------------
make_scenario_subplot <- function(df_panel,
                                  show_x = FALSE,
                                  show_y = FALSE,
                                  show_title = TRUE) {
  
  scen_name <- unique(df_panel$scenario)
  
  # robust panel-specific limits
  panel_min <- quantile(df_panel$pct_change, probs = 0.01, na.rm = TRUE)
  panel_max <- quantile(df_panel$pct_change, probs = 0.99, na.rm = TRUE)
  
  # safeguard
  if (!is.finite(panel_min)) panel_min <- min(df_panel$pct_change, na.rm = TRUE)
  if (!is.finite(panel_max)) panel_max <- max(df_panel$pct_change, na.rm = TRUE)
  
  # avoid zero-width scale
  if (panel_min == panel_max) {
    panel_min <- panel_min - 1
    panel_max <- panel_max + 1
  }
  
  x_breaks <- pretty(range(df_panel$x, na.rm = TRUE), n = 3)
  y_breaks <- pretty(range(df_panel$y, na.rm = TRUE), n = 3)
  
  p <- ggplot(df_panel, aes(x = x, y = y, fill = pct_change)) +
        geom_raster() +
    geom_sf(
      data = land_main,
      inherit.aes = FALSE,
      fill = "grey85",
      colour = "grey45",
      linewidth = 0.2
      ) +
    coord_sf(
      xlim = range(df_panel$x, na.rm = TRUE),
      ylim = range(df_panel$y, na.rm = TRUE),
      expand = FALSE
      ) +
    scale_x_continuous(breaks = x_breaks) +
    scale_y_continuous(breaks = y_breaks) +
    scale_fill_gradient2(
      low = "#E67E22",    # orange for decrease
      mid = "white",
      high = "#2C7FB8",   # blue for increase
      midpoint = 0,
      limits = c(panel_min, panel_max),
      oob = scales::squish,
      breaks = pretty(c(panel_min, panel_max), n = 3),
      labels = function(x) paste0(round(x, 1), "%"),
      na.value = "white",
      guide = guide_colorbar(
        title.position = "top",
        barheight = unit(13, "mm"),
        barwidth  = unit(4.5, "mm"),
        frame.colour = "black",
        ticks.colour = "black"
      )
    ) +
    labs(
      title = if (show_title) as.character(scen_name) else NULL,
      x = if (show_x) "Longitude" else NULL,
      y = if (show_y) "Latitude" else NULL,
      fill = "% change"
    ) +
    theme_bw(base_size = 11) +
    theme(
      plot.title = element_text(face = "bold", size = 12, hjust = 0.5),
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 12),
      legend.title = element_text(size = 11),
      legend.text = element_text(size = 11),
      panel.grid = element_blank(),
      
      # legend inside panel
      legend.position = c(0.99, 0.02),
      legend.justification = c(1, 0),
      
      legend.box.background = element_rect(
        colour = "black",
        fill = scales::alpha("white", 0.75)
      ),
      legend.background = element_rect(
        fill = scales::alpha("white", 0.75),
        colour = NA
      ),
      legend.margin = margin(2, 2, 2, 2),
      legend.key.height = unit(0.4, "cm"),
      legend.key.width  = unit(0.25, "cm")
    )
  
  if (!show_x) {
    p <- p + theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank()
    )
  }
  
  if (!show_y) {
    p <- p + theme(
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    )
  }
  
  p
}
```

Plot function
```{r}
plot_scenario_metric_pages <- function(df_metric, metric_name, out_dir,
                                       ncol = 4, nrow_entities = 5,
                                       width = 15.5, height = 18) {
  
  entities <- unique(df_metric$entity)
  entity_pages <- split(entities, ceiling(seq_along(entities) / nrow_entities))
  
  page_plots <- list()
  
  for (i in seq_along(entity_pages)) {
    
    ents_this_page <- entity_pages[[i]]
    
    # ------------------------------------------------------
    # Top header row with scenario names
    # ------------------------------------------------------
    header_plots <- list(
      ggplot() + theme_void(),  # empty left spacer
      ggplot() +
        annotate("text", x = 0.5, y = 0.5, label = "CC",
                 fontface = "bold", size = 7) +
        xlim(0, 1) + ylim(0, 1) + theme_void(),
      ggplot() +
        annotate("text", x = 0.5, y = 0.5, label = "OWF",
                 fontface = "bold", size = 7) +
        xlim(0, 1) + ylim(0, 1) + theme_void(),
      ggplot() +
        annotate("text", x = 0.5, y = 0.5, label = "OWF + CC",
                 fontface = "bold", size = 7) +
        xlim(0, 1) + ylim(0, 1) + theme_void()
    )
    
    header_row <- wrap_plots(header_plots, ncol = 4) +
      plot_layout(widths = c(0.9, 3, 3, 3))
    
    # ------------------------------------------------------
    # Entity blocks
    # ------------------------------------------------------
    entity_blocks <- list()
    
    for (r in seq_along(ents_this_page)) {
      
      ent <- ents_this_page[r]
      df_ent <- df_metric %>% filter(entity == ent)
      
      # Species/fleet title above row
      title_plot <- ggplot() +
        annotate("text",
                 x = 0, y = 0.5,
                 label = ent,
                 hjust = 0,
                 fontface = "bold",
                 size = 5.5) +
        xlim(0, 1) + ylim(0, 1) +
        theme_void()
      
      # Three scenario panels
      row_plots <- list()
      
      for (c in 1:3) {
        
        scen <- levels(df_metric$scenario)[c]
        df_panel <- df_ent %>% filter(scenario == scen)
        
        row_plots[[c]] <- make_scenario_subplot(
          df_panel,
          show_x = (r == length(ents_this_page)),
          show_y = (c == 1),
          show_title = FALSE   # IMPORTANT: titles now handled by header row
        )
      }
      
      row_plot <- wrap_plots(row_plots, ncol = 3)
      
      entity_block <- title_plot / row_plot +
        plot_layout(heights = c(0.12, 1))
      
      entity_blocks[[r]] <- entity_block
    }
    
    maps_block <- wrap_plots(entity_blocks, ncol = 1)
    
    final_plot <- header_row / maps_block +
      plot_layout(heights = c(0.04, 1)) +
      plot_annotation(
        title = paste0(
          str_to_title(gsub("_", " ", metric_name)),
          " | % change relative to Baseline | Plot ",
          i, "/", length(entity_pages)
        )
      )
    
    print(final_plot)
    
    out_file <- file.path(
      out_dir,
      paste0("ScenarioComparison_", metric_name, "_page_", i, ".png")
    )
    
    ggsave(out_file, final_plot, width = width, height = height, dpi = 300)
    
    page_plots[[i]] <- final_plot
  }
  
  return(page_plots)
}
```

Run plot function for all metrics
```{r}
############################################################
# MAKE SCENARIO PLOTS
############################################################

scenario_plot_list <- list()

metrics_scenarios <- c("biomass", "catch", "discards", "effort", "habitat_capacity")

for (met_name in metrics_scenarios) {
  
  message("====================================================")
  message("SCENARIO PLOTS FOR: ", met_name)
  message("====================================================")
  
  df_metric <- prepare_scenario_metric(
    scenario_plot_df,
    metric_name = met_name,
    cap_percent = 100
  )
  
  scenario_plot_list[[met_name]] <- plot_scenario_metric_pages(
    df_metric      = df_metric,
    metric_name    = met_name,
    out_dir        = out_dir,
    ncol           = 4,
    nrow_entities  = 5,
    width          = 13,
    height         = 18
  )
}
```

