---
title: "Ecological_Indicator_Processing_plots"
author: "Dries Lorré"
date: "2026-03-12"
output: html_document
---

In this script, the data of ecological indicators are processed and plotted.

Packages
```{r}
library(dplyr)
library(stringr)
library(ggplot2)
library(readr)
library(purrr)
library(tidyr)
library(patchwork)
library(scales)
library(sf)
library(terra)
```

Directories
```{r}
scenario_dirs <- c(
  Baseline  = "C:/Users/dlorr/OneDrive - UGent/Documents/EwE output/SBNS_CC_adapted_spatial_baseline_test",
  CC        = "C:/Users/dlorr/OneDrive - UGent/Documents/EwE output/SBNS_CC_adapted_spatial_CC_test",
  OWF       = "C:/Users/dlorr/OneDrive - UGent/Documents/EwE output/SBNS_CC_adapted_spatial_OWF_test",
  OWF_CC    = "C:/Users/dlorr/OneDrive - UGent/Documents/EwE output/SBNS_CC_adapted_spatial_OWF_CC_test"
)

csv_subfolder <- "Test_Env_resps"

out_dir <- file.path(getwd(), "Ecological_indicator_plots")
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

# load pre-cropped land object
land_main <- readRDS(
  "land_main_study_area.rds"
)
land_main_df <- land_main |>
  st_cast("MULTIPOLYGON") |>
  st_cast("POLYGON") |>
  mutate(poly_id = row_number())

land_main_df_plot <- do.call(
  rbind,
  lapply(seq_len(nrow(land_main_df)), function(i) {
    coords <- st_coordinates(land_main_df[i, ])
    data.frame(
      x = coords[, 1],
      y = coords[, 2],
      poly_id = i
    )
  })
)
# use one reference raster to guarantee identical CRS to the plotted rasters
# ref_rast <- rast("OWF_MPA_2150.asc")
# crs(ref_rast)
# crs(land_main)
# reproject land to raster CRS
# land_main <- st_transform(land_main, crs(ref_rast))
```

Setting up targets
```{r}
# last 10 yearly outputs
target_steps <- seq(1812, 1920, by = 12)

# folder inside each scenario directory
csv_subfolder <- "Test_Env_resps"

# indicator groups for the baseline figure
indicator_groups <- list(
  "Biomass-based" = c(
    "Total B",
    "Commercial B",
    "Fish B",
    "Invertebrates B",
    "Demersal B",
    "Pelagic B"
  ),
  "Catch-based" = c(
    "Total C",
    "Fish C",
    "Invertebrate C",
    "Demersal C",
    "Pelagic C"
  ),
  "Trophic level-based" = c(
    "TL community",
    "TL community 3,25",
    "TL catch"
  )
)

# flattened order
selected_indicators <- unlist(indicator_groups, use.names = FALSE)
```

Helper functions
```{r}
extract_step <- function(x) {
  as.integer(stringr::str_extract(basename(x), "(?<=Ecospace)\\d+(?=\\.csv$)"))
}

get_csv_dir <- function(scenario_path) {
  file.path(scenario_path, csv_subfolder)
}

list_target_csvs <- function(scenario_path, steps = target_steps) {
  csv_dir <- get_csv_dir(scenario_path)
  
  files <- list.files(
    csv_dir,
    pattern = "^biodiv_ind_Ecospace\\d+\\.csv$",
    full.names = TRUE
  )
  
  tibble(file = files) %>%
    mutate(step = extract_step(file)) %>%
    filter(step %in% steps) %>%
    arrange(step)
}

read_indicator_csv <- function(file) {
  readr::read_csv(file, show_col_types = FALSE)
}
```

Baseline scenario: processing
```{r}
baseline_files <- list_target_csvs(scenario_dirs["Baseline"])

baseline_raw <- lapply(seq_len(nrow(baseline_files)), function(i) {
  df <- read_indicator_csv(baseline_files$file[i])
  df$step <- baseline_files$step[i]
  df
}) %>%
  bind_rows()

# quick check: are all requested indicators present?
missing_indicators <- setdiff(selected_indicators, names(baseline_raw))
missing_indicators
```

```{r}
baseline_mean <- baseline_raw %>%
  select(Column, Row, Lon, Lat, step, all_of(selected_indicators)) %>%
  group_by(Column, Row, Lon, Lat) %>%
  summarise(
    across(all_of(selected_indicators), ~ mean(.x, na.rm = TRUE)),
    .groups = "drop"
  )

baseline_long <- baseline_mean %>%
  pivot_longer(
    cols = all_of(selected_indicators),
    names_to = "indicator",
    values_to = "value"
  ) %>%
  mutate(
    indicator = factor(indicator, levels = selected_indicators)
  )
```

Baseline: subplot function
```{r}
make_baseline_indicator_subplot <- function(df_panel,
                                            show_x = FALSE,
                                            show_y = FALSE) {
    
  # ensure full raster grid (this is the fix)
  df_panel <- df_panel %>%
    tidyr::complete(
      Lon = unique(Lon),
      Lat = unique(Lat)
    )
  
  # panel-specific limits
  panel_min <- quantile(df_panel$value, probs = 0.01, na.rm = TRUE)
  panel_max <- quantile(df_panel$value, probs = 0.99, na.rm = TRUE)
  
  # safeguard
  if (!is.finite(panel_min)) panel_min <- 0
  if (!is.finite(panel_max)) panel_max <- max(df_panel$value, na.rm = TRUE)
  
  if (panel_min == panel_max) {
    panel_max <- panel_min + 1e-6
  }
  
  # squish values above the chosen upper limit into the top color
  p <- ggplot(df_panel, aes(x = Lon, y = Lat, fill = value)) +
    geom_raster() +
geom_polygon(
  data = land_main_df_plot,
  aes(x = x, y = y, group = poly_id),
  inherit.aes = FALSE,
  fill = "grey85",
  colour = "grey45",
  linewidth = 0.2
) +
coord_equal(expand = FALSE) +
    scale_x_continuous(breaks = pretty(range(df_panel$Lon, na.rm = TRUE), n = 3)) +
    scale_y_continuous(breaks = pretty(range(df_panel$Lat, na.rm = TRUE), n = 3)) +
    scale_fill_distiller(
      palette = "Spectral",
      direction = -1,
      limits = c(panel_min, panel_max),
      oob = scales::squish,
      breaks = pretty(c(panel_min, panel_max), n = 3),
      na.value = "white",
      guide = guide_colorbar(
        title.position = "top",
        barheight = unit(15, "mm"),
        barwidth  = unit(5, "mm"),
        frame.colour = "black",
        ticks.colour = "black"
      )
    ) +
    labs(
      fill = NULL,
      x = if (show_x) "Longitude" else NULL,
      y = if (show_y) "Latitude" else NULL
    ) +
    theme_bw(base_size = 11) +
    theme(
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 12),
      legend.title = element_blank(),
      legend.text = element_text(size = 11),
      panel.grid = element_blank(),
      panel.background = element_rect(fill = "white", colour = NA),
      legend.position = c(0.98, 0.02),
      legend.justification = c(1, 0),
      legend.box.background = element_rect(
        colour = "black",
        fill = scales::alpha("white", 0.75)
      ),
      legend.background = element_rect(
        fill = scales::alpha("white", 0.75),
        colour = NA
      ),
      plot.margin = margin(2, 2, 2, 2)
    )
  
  if (!show_x) {
    p <- p + theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank()
    )
  }
  
  if (!show_y) {
    p <- p + theme(
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    )
  }
  
  p
}
```

Baseline: grouped plot
```{r}
plot_grouped_baseline_figure <- function(df_long, indicator_groups, out_dir,
                                         filename = "Baseline_ecological_indicators.png",
                                         width = 13, height = 16) {
  
  group_blocks <- list()
  group_heights <- c()
  
  for (grp_name in names(indicator_groups)) {
    
    inds <- indicator_groups[[grp_name]]
    df_grp <- df_long %>% filter(indicator %in% inds)
    
    n_inds <- length(inds)
    ncol <- 3
    n_rows_used <- ceiling(n_inds / ncol)
    
    # -----------------------------
    # group title row
    # -----------------------------
    group_title <- ggplot() +
      annotate("text",
               x = 0.02, y = 0.15,
               label = grp_name,
               hjust = 0,
               fontface = "bold",
               size = 6) +
      coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), clip = "off") +
      theme_void() +
      theme(
        plot.margin = margin(0, 0, -4, 0)
      )
    
    # -----------------------------
    # indicator blocks
    # -----------------------------
    indicator_blocks <- lapply(seq_along(inds), function(i) {
      ind <- inds[i]
      df_panel <- df_grp %>% filter(indicator == ind)
      
      row_pos <- ceiling(i / ncol)
      col_pos <- i %% ncol
      if (col_pos == 0) col_pos <- ncol
      
      ind_title <- ggplot() +
        annotate("text",
                 x = 0.02, y = 0.05,
                 label = ind,
                 hjust = 0,
                 fontface = "bold",
                 size = 4.6) +
        coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), clip = "off") +
        theme_void() +
        theme(
          plot.margin = margin(0, 0, -8, 0)
        )
      
      map_plot <- make_baseline_indicator_subplot(
        df_panel,
        show_x = row_pos == n_rows_used,
        show_y = col_pos == 1
      )
      
      ind_title / map_plot + plot_layout(heights = c(0.02, 0.1))
    })
    
    maps_block <- wrap_plots(
      indicator_blocks,
      ncol = ncol,
      nrow = n_rows_used
    )
    
    # group block: title row + maps
    group_block <- group_title / maps_block +
      plot_layout(heights = c(0.035, 1))
    
    group_blocks[[grp_name]] <- group_block
    group_heights <- c(group_heights, n_rows_used)
  }
  
  final_plot <- wrap_plots(group_blocks, ncol = 1) +
    plot_layout(heights = group_heights)
  
  print(final_plot)
  
  ggsave(
    filename = file.path(out_dir, filename),
    plot = final_plot,
    width = width,
    height = height,
    dpi = 800
  )
  
  final_plot
}
```

Baseline figure
```{r}
baseline_indicator_figure <- plot_grouped_baseline_figure(
  df_long = baseline_long,
  indicator_groups = indicator_groups,
  out_dir = out_dir,
  filename = "Baseline_ecological_indicators.png",
  width = 13,
  height = 16
)
```
----------------------------------------------------------
SCENARIO PLOTS

First process all scenario's
```{r}
process_scenario_indicators <- function(scenario_name, scenario_path, selected_indicators) {
  
  file_tbl <- list_target_csvs(scenario_path)
  
  df_raw <- lapply(seq_len(nrow(file_tbl)), function(i) {
    df <- read_indicator_csv(file_tbl$file[i])
    df$step <- file_tbl$step[i]
    df
  }) %>%
    bind_rows()
  
  df_mean <- df_raw %>%
    select(Column, Row, Lon, Lat, step, all_of(selected_indicators)) %>%
    group_by(Column, Row, Lon, Lat) %>%
    summarise(
      across(all_of(selected_indicators), ~ mean(.x, na.rm = TRUE)),
      .groups = "drop"
    ) %>%
    mutate(scenario = scenario_name)
  
  df_long <- df_mean %>%
    pivot_longer(
      cols = all_of(selected_indicators),
      names_to = "indicator",
      values_to = "value"
    ) %>%
    mutate(
      indicator = factor(indicator, levels = selected_indicators)
    )
  
  df_long
}

all_indicator_results <- list()

for (sc_name in names(scenario_dirs)) {
  message("Processing scenario: ", sc_name)
  
  all_indicator_results[[sc_name]] <- process_scenario_indicators(
    scenario_name = sc_name,
    scenario_path = scenario_dirs[[sc_name]],
    selected_indicators = selected_indicators
  )
}

all_indicators_df <- bind_rows(all_indicator_results)
```

% change vs baseline
```{r}
baseline_ref <- all_indicators_df %>%
  filter(scenario == "Baseline") %>%
  select(Column, Row, Lon, Lat, indicator, baseline_value = value)

scenario_change_df <- all_indicators_df %>%
  filter(scenario != "Baseline") %>%
  left_join(
    baseline_ref,
    by = c("Column", "Row", "Lon", "Lat", "indicator")
  ) %>%
  mutate(
    pct_change = case_when(
      baseline_value == 0 & value == 0 ~ 0,
      baseline_value == 0 & value != 0 ~ NA_real_,
      TRUE ~ ((value - baseline_value) / baseline_value) * 100
    ),
    scenario = factor(
      scenario,
      levels = c("CC", "OWF", "OWF_CC"),
      labels = c("CC", "OWF", "OWF + CC")
    )
  )
```

plot function
```{r}
make_scenario_indicator_subplot <- function(df_panel,
                                            show_x = FALSE,
                                            show_y = FALSE) {
  
  # ensure full raster grid (this is the fix)
  df_panel <- df_panel %>%
    tidyr::complete(
      Lon = unique(Lon),
      Lat = unique(Lat)
    )
  # panel-specific limits
  panel_min <- quantile(df_panel$pct_change, probs = 0.01, na.rm = TRUE)
  panel_max <- quantile(df_panel$pct_change, probs = 0.99, na.rm = TRUE)
  
  # safeguards
  if (!is.finite(panel_min)) panel_min <- min(df_panel$pct_change, na.rm = TRUE)
  if (!is.finite(panel_max)) panel_max <- max(df_panel$pct_change, na.rm = TRUE)
  
  if (!is.finite(panel_min)) panel_min <- -1
  if (!is.finite(panel_max)) panel_max <- 1
  
  if (panel_min == panel_max) {
    panel_min <- panel_min - 1
    panel_max <- panel_max + 1
  }
  
  p <- ggplot(df_panel, aes(x = Lon, y = Lat, fill = pct_change)) +
    geom_raster() +
    geom_polygon(
      data = land_main_df_plot,
      aes(x = x, y = y, group = poly_id),
      inherit.aes = FALSE,
      fill = "grey85",
      colour = "grey45",
      linewidth = 0.2
      ) +
    coord_equal(expand = FALSE) +
    scale_x_continuous(breaks = pretty(range(df_panel$Lon, na.rm = TRUE), n = 3)) +
    scale_y_continuous(breaks = pretty(range(df_panel$Lat, na.rm = TRUE), n = 3)) +
    scale_fill_gradient2(
      low = "#E67E22",
      mid = "white",
      high = "#2C7FB8",
      midpoint = 0,
      limits = c(panel_min, panel_max),
      oob = scales::squish,
      breaks = pretty(c(panel_min, panel_max), n = 3),
      labels = function(x) paste0(round(x, 1), "%"),
      na.value = "white",
      guide = guide_colorbar(
        title.position = "top",
        barheight = unit(15, "mm"),
        barwidth  = unit(5, "mm"),
        frame.colour = "black",
        ticks.colour = "black"
      )
    ) +
    labs(
      fill = NULL,
      x = if (show_x) "Longitude" else NULL,
      y = if (show_y) "Latitude" else NULL
    ) +
    theme_bw(base_size = 11) +
    theme(
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 12),
      legend.title = element_text(size = 11),
      legend.text = element_text(size = 11),
      panel.grid = element_blank(),
      panel.background = element_rect(
        fill = "white", colour = NA),
      legend.position = c(0.98, 0.02),
      legend.justification = c(1, 0),
      legend.box.background = element_rect(
        colour = "black",
        fill = scales::alpha("white", 0.75)
      ),
      legend.background = element_rect(
        fill = scales::alpha("white", 0.75),
        colour = NA
      ),
      plot.margin = margin(2, 2, 2, 2)
    )
  
  if (!show_x) {
    p <- p + theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank()
    )
  }
  
  if (!show_y) {
    p <- p + theme(
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    )
  }
  
  p
}
```

Grouped scenario plots
```{r}
plot_grouped_scenario_figure <- function(df_change, indicator_groups, group_name,
                                         out_dir, filename,
                                         width = 14, height = 16) {
  
  inds <- indicator_groups[[group_name]]
  
  df_grp <- df_change %>%
    filter(indicator %in% inds) %>%
    mutate(indicator = factor(indicator, levels = inds))
  
  # -----------------------------
  # scenario header row
  # -----------------------------
  header_row <- wrap_plots(
    list(
      ggplot() + theme_void(),
      ggplot() +
        annotate("text", x = 0.375, y = 0.2, label = "   CC   ",
                 fontface = "bold", size = 6) +
        xlim(0, 1) + ylim(0, 1) + theme_void(),
      ggplot() +
        annotate("text", x = 0.375, y = 0.2, label = "   OWF  ",
                 fontface = "bold", size = 6) +
        xlim(0, 1) + ylim(0, 1) + theme_void(),
      ggplot() +
        annotate("text", x = 0.375, y = 0.2, label = "OWF + CC",
                 fontface = "bold", size = 6) +
        xlim(0, 1) + ylim(0, 1) + theme_void()
    ),
    ncol = 4
  ) +
    plot_layout(widths = c(0.9, 3, 3, 3))
  
  # -----------------------------
  # indicator rows
  # -----------------------------
  row_blocks <- lapply(seq_along(inds), function(i) {
    ind <- inds[i]
    df_ind <- df_grp %>% filter(indicator == ind)
    
    # indicator title
    row_title <- ggplot() +
      annotate("text",
               x = 0.02, y = 0.25,
               label = ind,
               hjust = 0,
               fontface = "bold",
               size = 4.6) +
      coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), clip = "off") +
      theme_void() +
      theme(
        plot.margin = margin(0, 0, -8, 0)
      )
    
    scen_levels <- c("CC", "OWF", "OWF + CC")
    
    row_maps <- lapply(seq_along(scen_levels), function(j) {
      sc <- scen_levels[j]
      df_panel <- df_ind %>% filter(scenario == sc)
      
      make_scenario_indicator_subplot(
        df_panel,
        show_x = i == length(inds),
        show_y = j == 1
      )
    })
    
    maps_row <- wrap_plots(row_maps, ncol = 3)
    
    row_title / maps_row + plot_layout(heights = c(0.02, 0.1))
  })
  
  maps_block <- wrap_plots(row_blocks, ncol = 1)
  
  final_plot <- header_row / maps_block +
    plot_layout(heights = c(0.04, 1))
  
  print(final_plot)
  
  ggsave(
    filename = file.path(out_dir, filename),
    plot = final_plot,
    width = width,
    height = height,
    dpi = 800
  )
  
  final_plot
}
```

Make all plots
```{r}
# Manuscript figures:
indicator_groups_ms <- list(
  
  "Biomass-based" = c(
    "Total B",
    "Fish B",
    "Invertebrates B"
  ),
  
  "Catch-based" = c(
    "Total C",
    "Fish C",
    "Invertebrate C"
  ),
  
  "Trophic level-based" = c(
    "TL community",
    "TL community 3,25",
    "TL catch"
  )
)

# Supplementary figues
indicator_groups_sup <- list(
  
  "Biomass-based" = c(
    "Commercial B",
    "Demersal B",
    "Pelagic B"
  ),
  
  "Catch-based" = c(
    "Demersal C",
    "Pelagic C"
  )
)


scenario_ms_plots <- list()

for (grp in names(indicator_groups_ms)) {
  
  scenario_ms_plots[[grp]] <- plot_grouped_scenario_figure(
    df_change = scenario_change_df,
    indicator_groups = indicator_groups_ms,
    group_name = grp,
    out_dir = out_dir,
    filename = paste0("Scenario_", grp, "_MANUSCRIPT.png"),
    width = 14,
    height = 10
  )
}

scenario_sup_plots <- list()

for (grp in names(indicator_groups_sup)) {
  
  scenario_sup_plots[[grp]] <- plot_grouped_scenario_figure(
    df_change = scenario_change_df,
    indicator_groups = indicator_groups_sup,
    group_name = grp,
    out_dir = out_dir,
    filename = paste0("Scenario_", grp, "_SUPPLEMENTARY.png"),
    width = 14,
    height = 10
  )
}
# scenario_group_plot_list[["Biomass-based"]] <- plot_grouped_scenario_figure(
#   df_change = scenario_change_df,
#   indicator_groups = indicator_groups,
#   group_name = "Biomass-based",
#   out_dir = out_dir,
#   filename = "Scenario_ecological_indicators_Biomass_based.png",
#   width = 14,
#   height = 16
# )

# scenario_group_plot_list[["Catch-based"]] <- plot_grouped_scenario_figure(
#   df_change = scenario_change_df,
#   indicator_groups = indicator_groups,
#   group_name = "Catch-based",
#   out_dir = out_dir,
#   filename = "Scenario_ecological_indicators_Catch_based.png",
#   width = 14,
#   height = 14
# )



scenario_group_plot_list <- list()

scenario_group_plot_list[["Trophic level-based"]] <- plot_grouped_scenario_figure(
  df_change = scenario_change_df,
  indicator_groups = indicator_groups,
  group_name = "Trophic level-based",
  out_dir = out_dir,
  filename = "Scenario_ecological_indicators_Trophic_level_based.png",
  width = 14,
  height = 10
)
```

Include zonal summaries
```{r}
library(terra)
owf_rast <- rast("OWF_MPA_2150.asc")
# owf_rast
# ext(owf_rast)
# res(owf_rast)
# crs(owf_rast)

# Read binary OWF raster
owf_rast <- rast("OWF_MPA_2150.asc")

# Ensure binary 0/1
owf_bin <- ifel(owf_rast > 0, 1, 0)

# Build a 3-cell spillover ring using a 7x7 moving window
# This is a "3-cell neighbourhood" buffer in raster-cell terms
w <- matrix(1, nrow = 5, ncol = 5)

owf_buffered <- focal(
  owf_bin,
  w = w,
  fun = max,
  na.policy = "omit"
)

spillover_bin <- ifel(owf_buffered == 1 & owf_bin == 0, 1, 0)
outside_bin   <- ifel(owf_buffered == 0, 1, 0)

# Convert to dataframe on the same grid
owf_df <- as.data.frame(owf_bin, xy = TRUE, na.rm = FALSE)
names(owf_df) <- c("Lon", "Lat", "in_owf")

spill_df <- as.data.frame(spillover_bin, xy = TRUE, na.rm = FALSE)
names(spill_df) <- c("Lon", "Lat", "in_spill")

out_df <- as.data.frame(outside_bin, xy = TRUE, na.rm = FALSE)
names(out_df) <- c("Lon", "Lat", "in_outside")

zone_df <- owf_df %>%
  left_join(spill_df, by = c("Lon", "Lat")) %>%
  left_join(out_df,   by = c("Lon", "Lat")) %>%
  mutate(
    zone = case_when(
      in_owf == 1 ~ "OWF",
      in_spill == 1 ~ "Spillover",
      in_outside == 1 ~ "Outside OWF",
      TRUE ~ NA_character_
    )
  ) %>%
  select(Lon, Lat, zone)

# new code using col and row
nr <- nrow(owf_bin)
nc <- ncol(owf_bin)

owf_vals   <- as.vector(values(owf_bin, mat = FALSE))
spill_vals <- as.vector(values(spillover_bin, mat = FALSE))
out_vals   <- as.vector(values(outside_bin, mat = FALSE))

zone_df_rc <- expand.grid(
  Column = 1:nc,
  Row = 1:nr
) %>%
  as_tibble() %>%
  mutate(
    in_owf = owf_vals,
    in_spill = spill_vals,
    in_outside = out_vals,
    zone = case_when(
      in_owf == 1 ~ "OWF",
      in_spill == 1 ~ "Spillover",
      in_outside == 1 ~ "Outside OWF",
      TRUE ~ NA_character_
    )
  ) %>%
  select(Column, Row, zone)
```

Assign zones to data
```{r}
zone_df_join <- zone_df %>%
  mutate(
    Lon = round(Lon, 6),
    Lat = round(Lat, 6)
  )

scenario_change_zones <- scenario_change_df %>%
  mutate(
    Lon = round(Lon, 6),
    Lat = round(Lat, 6)
  ) %>%
  left_join(zone_df_join, by = c("Lon", "Lat"))

scenario_change_zones <- scenario_change_df %>%
  left_join(zone_df_rc, by = c("Column", "Row"))
```


Calculate zonal summaries
```{r}
# Overall summaries for all scenarios
zone_summary_overall <- scenario_change_zones %>%
  group_by(indicator, scenario) %>%
  summarise(
    mean_pct_change = mean(pct_change, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(zone = "Overall")

# Zonal summaries only for OWF-based scenarios
zone_summary_zones <- scenario_change_zones %>%
  filter(
    scenario %in% c("OWF", "OWF + CC"),
    !is.na(zone)
  ) %>%
  group_by(indicator, scenario, zone) %>%
  summarise(
    mean_pct_change = mean(pct_change, na.rm = TRUE),
    .groups = "drop"
  )

zone_summary <- bind_rows(zone_summary_overall, zone_summary_zones) %>%
  mutate(
    zone = factor(zone, levels = c("OWF", "Spillover", "Outside OWF", "Overall"))
  )


# NEW SUMMARY
zone_summary_overall <- scenario_change_zones %>%
  group_by(indicator, scenario) %>%
  summarise(
    mean_pct_change = mean(pct_change, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(zone = "Overall")

zone_summary_zones <- scenario_change_zones %>%
  filter(
    scenario %in% c("OWF", "OWF + CC"),
    !is.na(zone)
  ) %>%
  group_by(indicator, scenario, zone) %>%
  summarise(
    mean_pct_change = mean(pct_change, na.rm = TRUE),
    .groups = "drop"
  )

zone_summary <- bind_rows(zone_summary_overall, zone_summary_zones) %>%
  mutate(
    zone = factor(zone, levels = c("OWF", "Spillover", "Outside OWF", "Overall"))
  )
```

Plot for visual check
```{r}
# coordinates from the raster
xy_df <- as.data.frame(owf_bin, xy = TRUE, na.rm = FALSE) %>%
  select(x, y) %>%
  rename(Lon = x, Lat = y)

zone_plot_df <- bind_cols(
  zone_df_rc,
  xy_df
) %>%
  mutate(
    zone = factor(zone, levels = c("OWF", "Spillover", "Outside OWF"))
  )

zone_check_plot <- ggplot(zone_plot_df, aes(x = Lon, y = Lat, fill = zone)) +
  geom_raster() +
  coord_equal() +
  scale_x_continuous(breaks = pretty(range(zone_plot_df$Lon, na.rm = TRUE), n = 3)) +
  scale_y_continuous(breaks = pretty(range(zone_plot_df$Lat, na.rm = TRUE), n = 3)) +
  scale_fill_manual(
    values = c(
      "OWF" = "#1f78b4",
      "Spillover" = "#33a02c",
      "Outside OWF" = "grey85"
    ),
    na.value = "white"
  ) +
  labs(
    title = "OWF zones",
    x = "Longitude",
    y = "Latitude",
    fill = "Zone"
  ) +
  theme_bw(base_size = 12) +
  theme(
    panel.grid = element_blank(),
    legend.position = "right"
  )

print(zone_check_plot)

ggsave(
  file.path(out_dir, "OWF_zone_check_map.png"),
  zone_check_plot,
  width = 8,
  height = 5,
  dpi = 800
)
```
Barplots
```{r}
make_zone_barplot <- function(df_zone_panel) {
  
  if (nrow(df_zone_panel) == 0) {
    return(
      ggplot() +
        theme_void()
    )
  }
  
df_zone_panel <- df_zone_panel %>%
  mutate(
    zone_label = case_when(
      zone == "OWF" ~ "OWF",
      zone == "Spillover" ~ "SPILL",
      zone == "Outside OWF" ~ "OUT",
      zone == "Overall" ~ "ALL",
      TRUE ~ as.character(zone)
    ),
    zone_label = factor(zone_label, levels = c("ALL","OUT","OWF","SPILL")),
    change_dir = ifelse(mean_pct_change < 0, "decrease", "increase")
  )
  
ggplot(df_zone_panel, aes(x = zone_label, y = mean_pct_change, fill = change_dir)) +
  geom_col(width = 0.4) +
  geom_hline(yintercept = 0, linewidth = 0.3) +
  scale_x_discrete(
  limits = c("ALL","OUT","OWF","SPILL"),
  labels = function(x){
    if(length(unique(df_zone_panel$zone_label)) == 1){
      ifelse(x == "ALL","ALL","")
    } else {
      x
    }
  }
) +
  scale_fill_manual(
    values = c(
      "decrease" = "#E67E22",
      "increase" = "#2C7FB8"
    ),
    guide = "none"
  ) +
    labs(x = NULL, y = NULL) +
    theme_bw(base_size = 11) +
    theme(
      axis.title = element_blank(),
      axis.text.x = element_text(size = 12, angle = 0, hjust = 0.5),
      axis.text.y = element_text(size = 12),
      axis.ticks.x = element_blank(),
      panel.grid = element_blank(),
      plot.margin = margin(-12, 2, 0, 2)
    )
}
```

```{r}
make_scenario_indicator_panel <- function(df_panel,
                                          df_zone_panel,
                                          show_x = FALSE,
                                          show_y = FALSE) {
  
  # ensure full raster grid (this is the fix)
  df_panel <- df_panel %>%
    tidyr::complete(
      Lon = unique(Lon),
      Lat = unique(Lat)
    )
  # robust panel-specific limits
  panel_min <- quantile(df_panel$pct_change, probs = 0.01, na.rm = TRUE)
  panel_max <- quantile(df_panel$pct_change, probs = 0.99, na.rm = TRUE)
  
  if (!is.finite(panel_min)) panel_min <- min(df_panel$pct_change, na.rm = TRUE)
  if (!is.finite(panel_max)) panel_max <- max(df_panel$pct_change, na.rm = TRUE)
  
  if (!is.finite(panel_min)) panel_min <- -1
  if (!is.finite(panel_max)) panel_max <- 1
  
  if (panel_min == panel_max) {
    panel_min <- panel_min - 1
    panel_max <- panel_max + 1
  }
  
  map_plot <- ggplot(df_panel, aes(x = Lon, y = Lat, fill = pct_change)) +
    geom_raster() +
    geom_polygon(
      data = land_main_df_plot,
      aes(x = x, y = y, group = poly_id),
      inherit.aes = FALSE,
      fill = "grey85",
      colour = "grey45",
      linewidth = 0.2
      ) +
    coord_equal(expand = FALSE) +
    scale_x_continuous(breaks = pretty(range(df_panel$Lon, na.rm = TRUE), n = 3)) +
    scale_y_continuous(breaks = pretty(range(df_panel$Lat, na.rm = TRUE), n = 3)) +
    scale_fill_gradient2(
      low = "#E67E22",
      mid = "white",
      high = "#2C7FB8",
      midpoint = 0,
      limits = c(panel_min, panel_max),
      oob = scales::squish,
      breaks = pretty(c(panel_min, panel_max), n = 3),
      labels = function(x) paste0(round(x, 1), "%"),
      na.value = "white",
      guide = guide_colorbar(
        title.position = "top",
        barheight = unit(13, "mm"),
        barwidth  = unit(4.5, "mm"),
        frame.colour = "black",
        ticks.colour = "black"
      )
    ) +
    labs(
      fill = NULL,
      x = "Longitude",
      y = "Latitude"
    ) +
    theme_bw(base_size = 11) +
    theme(
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 12),
      legend.title = element_blank(),
      legend.text = element_text(size = 11),
      panel.grid = element_blank(),
      panel.background = element_rect(fill = "white", colour = NA),
      legend.position = c(0.98, 0.02),
      legend.justification = c(1, 0),
      legend.box.background = element_rect(
        colour = "black",
        fill = scales::alpha("white", 0.75)
      ),
      legend.background = element_rect(
        fill = scales::alpha("white", 0.75),
        colour = NA
      ),
      plot.margin = margin(2, 2, -6, 2)
    )
  
  # Keep axis space equal across panels; hide unused elements with transparency
  if (!show_x) {
    map_plot <- map_plot + theme(
      axis.title.x = element_text(colour = NA),
      axis.text.x  = element_text(colour = NA),
      axis.ticks.x = element_line(colour = NA)
    )
  }
  
  if (!show_y) {
    map_plot <- map_plot + theme(
      axis.title.y = element_text(colour = NA),
      axis.text.y  = element_text(colour = NA),
      axis.ticks.y = element_line(colour = NA)
    )
  }
  
  bar_plot <- make_zone_barplot(df_zone_panel)
  
  map_plot / bar_plot + plot_layout(heights = c(1, 0.6))
}
```

```{r}
plot_grouped_scenario_figure <- function(df_change, indicator_groups, group_name,
                                         out_dir, filename,
                                         width = 14, height = 16) {
  
  inds <- indicator_groups[[group_name]]
  
  df_grp <- df_change %>%
    filter(indicator %in% inds) %>%
    mutate(indicator = factor(indicator, levels = inds))
  
  # -----------------------------
  # scenario header row
  # -----------------------------
  header_row <- wrap_plots(
    list(
      ggplot() + theme_void(),
      ggplot() +
        annotate("text", x = 0.375, y = 0.2, label = "   CC   ",
                 fontface = "bold", size = 6) +
        xlim(0, 1) + ylim(0, 1) + theme_void(),
      ggplot() +
        annotate("text", x = 0.375, y = 0.2, label = "   OWF  ",
                 fontface = "bold", size = 6) +
        xlim(0, 1) + ylim(0, 1) + theme_void(),
      ggplot() +
        annotate("text", x = 0.375, y = 0.2, label = "OWF + CC",
                 fontface = "bold", size = 6) +
        xlim(0, 1) + ylim(0, 1) + theme_void()
    ),
    ncol = 4
  ) +
    plot_layout(widths = c(0.9, 3, 3, 3))
  
  # -----------------------------
  # indicator rows
  # -----------------------------
  row_blocks <- lapply(seq_along(inds), function(i) {
    ind <- inds[i]
    df_ind <- df_grp %>% filter(indicator == ind)
    
    # indicator title
    row_title <- ggplot() +
      annotate("text",
               x = 0.02, y = 0.25,
               label = ind,
               hjust = 0,
               fontface = "bold",
               size = 4.6) +
      coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), clip = "off") +
      theme_void() +
      theme(
        plot.margin = margin(0, 0, -8, 0)
      )
    
    scen_levels <- c("CC", "OWF", "OWF + CC")
    
    row_maps <- lapply(seq_along(scen_levels), function(j) {
      sc <- scen_levels[j]
      
      df_panel <- df_ind %>% filter(scenario == sc)
      
      df_zone_panel <- zone_summary %>%
        filter(
          indicator == ind,
          scenario == sc
          )
      make_scenario_indicator_panel(
        df_panel = df_panel,
        df_zone_panel = df_zone_panel,
        show_x = i == length(inds),
        show_y = j == 1
        )
      })
    
    maps_row <- wrap_plots(row_maps, ncol = 3)
    
    row_title / maps_row + plot_layout(heights = c(0.02, 0.1))
  })
  
  maps_block <- wrap_plots(row_blocks, ncol = 1)
  
  final_plot <- header_row / maps_block +
    plot_layout(heights = c(0.04, 1))
  
  print(final_plot)
  
  ggsave(
    filename = file.path(out_dir, filename),
    plot = final_plot,
    width = width,
    height = height,
    dpi = 800
  )
  
  final_plot
}
```

test plot
```{r}
test_ind <- "Total B"
test_sc  <- "CC"

df_panel_test <- scenario_change_df %>%
  filter(indicator == test_ind, scenario == test_sc)

df_zone_test <- zone_summary %>%
  filter(indicator == test_ind, scenario == test_sc)

test_plot <- make_scenario_indicator_panel(
  df_panel = df_panel_test,
  df_zone_panel = df_zone_test,
  show_x = TRUE,
  show_y = TRUE
)

print(test_plot)
```
Plot for indicators

```{r}
scenario_group_plot_list <- list()

scenario_group_plot_list[["Biomass-based"]] <- plot_grouped_scenario_figure(
  df_change = scenario_change_df,
  indicator_groups = indicator_groups,
  group_name = "Biomass-based",
  out_dir = out_dir,
  filename = "Scenario_ecological_indicators_Biomass_based.png",
  width = 14,
  height = 16
)

scenario_group_plot_list[["Catch-based"]] <- plot_grouped_scenario_figure(
  df_change = scenario_change_df,
  indicator_groups = indicator_groups,
  group_name = "Catch-based",
  out_dir = out_dir,
  filename = "Scenario_ecological_indicators_Catch_based.png",
  width = 14,
  height = 14
)

scenario_group_plot_list[["Trophic level-based"]] <- plot_grouped_scenario_figure(
  df_change = scenario_change_df,
  indicator_groups = indicator_groups,
  group_name = "Trophic level-based",
  out_dir = out_dir,
  filename = "Scenario_ecological_indicators_Trophic_level_based.png",
  width = 14,
  height = 10
)
```

```{r}
############################################################
# MANUSCRIPT FIGURES
############################################################

scenario_ms_plot_list <- list()

scenario_ms_plot_list[["Biomass-based"]] <- plot_grouped_scenario_figure(
  df_change = scenario_change_df,
  indicator_groups = indicator_groups_ms,
  group_name = "Biomass-based",
  out_dir = out_dir,
  filename = "Scenario_ecological_indicators_Biomass_based_MANUSCRIPT.png",
  width = 14,
  height = 14.5
)

scenario_ms_plot_list[["Catch-based"]] <- plot_grouped_scenario_figure(
  df_change = scenario_change_df,
  indicator_groups = indicator_groups_ms,
  group_name = "Catch-based",
  out_dir = out_dir,
  filename = "Scenario_ecological_indicators_Catch_based_MANUSCRIPT.png",
  width = 14,
  height = 14.5
)

scenario_ms_plot_list[["Trophic level-based"]] <- plot_grouped_scenario_figure(
  df_change = scenario_change_df,
  indicator_groups = indicator_groups_ms,
  group_name = "Trophic level-based",
  out_dir = out_dir,
  filename = "Scenario_ecological_indicators_Trophic_level_based_MANUSCRIPT.png",
  width = 14,
  height = 14.5
)

############################################################
# SUPPLEMENTARY FIGURES
############################################################

scenario_sup_plot_list <- list()

scenario_sup_plot_list[["Biomass-based"]] <- plot_grouped_scenario_figure(
  df_change = scenario_change_df,
  indicator_groups = indicator_groups_sup,
  group_name = "Biomass-based",
  out_dir = out_dir,
  filename = "Scenario_ecological_indicators_Biomass_based_SUPPLEMENTARY.png",
  width = 14,
  height = 14.5
)

scenario_sup_plot_list[["Catch-based"]] <- plot_grouped_scenario_figure(
  df_change = scenario_change_df,
  indicator_groups = indicator_groups_sup,
  group_name = "Catch-based",
  out_dir = out_dir,
  filename = "Scenario_ecological_indicators_Catch_based_SUPPLEMENTARY.png",
  width = 14,
  height = 14.5
)
```
Table of absolute and relative changes according to zones (see also code zonal summaries above)
```{r}
# =========================================================
# SUMMARY TABLE: zonal absolute values + % change vs zonal baseline
# =========================================================

library(dplyr)
library(tidyr)
library(readr)

# ---------------------------------------------------------
# 1. Attach zones to ALL scenarios, including Baseline
# ---------------------------------------------------------
all_indicators_zones <- all_indicators_df %>%
  left_join(zone_df_rc, by = c("Column", "Row")) %>%
  mutate(
    zone = case_when(
      zone == "OWF" ~ "OWF",
      zone == "Spillover" ~ "SPILL",
      zone == "Outside OWF" ~ "OUT",
      TRUE ~ NA_character_
    )
  )

# ---------------------------------------------------------
# 2. Build a table with both:
#    - ALL = whole study area
#    - OWF / SPILL / OUT = zonal subsets
# ---------------------------------------------------------

# Whole study area
summary_all <- all_indicators_zones %>%
  group_by(indicator, scenario) %>%
  summarise(
    abs_value = mean(value, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(zone = "ALL")

# Zonal subsets
summary_zones <- all_indicators_zones %>%
  filter(!is.na(zone)) %>%
  group_by(indicator, scenario, zone) %>%
  summarise(
    abs_value = mean(value, na.rm = TRUE),
    .groups = "drop"
  )

summary_abs <- bind_rows(summary_all, summary_zones)

# ---------------------------------------------------------
# 3. Extract baseline zonal means
# ---------------------------------------------------------
baseline_zone_ref <- summary_abs %>%
  filter(scenario == "Baseline") %>%
  select(
    indicator,
    zone,
    baseline_abs_value = abs_value
  )

# ---------------------------------------------------------
# 4. Calculate % change relative to the baseline of the SAME zone
# ---------------------------------------------------------
summary_vs_baseline <- summary_abs %>%
  filter(scenario != "Baseline") %>%
  left_join(baseline_zone_ref, by = c("indicator", "zone")) %>%
  mutate(
    pct_change = case_when(
      baseline_abs_value == 0 & abs_value == 0 ~ 0,
      baseline_abs_value == 0 & abs_value != 0 ~ NA_real_,
      TRUE ~ ((abs_value - baseline_abs_value) / baseline_abs_value) * 100
    )
  )

# ---------------------------------------------------------
# 5. Optional: keep a logical indicator order
# ---------------------------------------------------------
summary_vs_baseline <- summary_vs_baseline %>%
  mutate(
    indicator = factor(indicator, levels = selected_indicators),
    scenario = factor(scenario, levels = c("CC", "OWF", "OWF_CC")),
    zone = factor(zone, levels = c("ALL", "OWF", "SPILL", "OUT"))
  ) %>%
  arrange(indicator, scenario, zone)

# ---------------------------------------------------------
# 6. Formatting function for table cells
#    Example: 12.34 (+5.6%)
# ---------------------------------------------------------
format_summary_cell <- function(abs_value, pct_change,
                                abs_digits = 2,
                                pct_digits = 1) {
  
  abs_txt <- formatC(abs_value, format = "f", digits = abs_digits)
  
  pct_txt <- ifelse(
    is.na(pct_change),
    "NA",
    paste0(
      ifelse(pct_change > 0, "+", ""),
      formatC(pct_change, format = "f", digits = pct_digits),
      "%"
    )
  )
  
  paste0(abs_txt, " (", pct_txt, ")")
}

summary_vs_baseline_fmt <- summary_vs_baseline %>%
  mutate(
    cell = format_summary_cell(abs_value, pct_change)
  )

# ---------------------------------------------------------
# 7. Wide table:
#    rows = indicators
#    columns = scenario x zone
# ---------------------------------------------------------
summary_table_wide <- summary_vs_baseline_fmt %>%
  mutate(
    scenario = recode(
      scenario,
      "OWF_CC" = "OWF + CC"
    )
  ) %>%
  unite(col_name, scenario, zone, sep = "_") %>%
  select(indicator, col_name, cell) %>%
  pivot_wider(
    names_from = col_name,
    values_from = cell
  ) %>%
  select(
    indicator,
    CC_ALL, CC_OWF, CC_SPILL, CC_OUT,
    OWF_ALL, OWF_OWF, OWF_SPILL, OWF_OUT,
    `OWF + CC_ALL`, `OWF + CC_OWF`, `OWF + CC_SPILL`, `OWF + CC_OUT`
  ) %>%
  mutate(
    indicator = as.character(indicator)
  )

summary_table_wide

```

Table showing absolute change and relative change.
```{r}
# =========================================================
# PREP: add absolute change
# =========================================================

summary_vs_baseline <- summary_vs_baseline %>%
  mutate(
    abs_change = abs_value - baseline_abs_value
  )
```

```{r}
# =========================================================
# TABLE: only relative changes (%)
# =========================================================

format_pct_cell <- function(x, digits = 1) {
  ifelse(
    is.na(x),
    "NA",
    paste0(
      ifelse(x > 0, "+", ""),
      formatC(x, format = "f", digits = digits),
      "%"
    )
  )
}

summary_table_pct <- summary_vs_baseline %>%
  mutate(
    scenario = recode(scenario, "OWF_CC" = "OWF + CC"),
    pct_cell = format_pct_cell(pct_change)
  ) %>%
  unite(col_name, scenario, zone, sep = "_") %>%
  select(indicator, col_name, pct_cell) %>%
  pivot_wider(names_from = col_name, values_from = pct_cell) %>%
  select(
    indicator,
    CC_ALL, CC_OWF, CC_SPILL, CC_OUT,
    OWF_ALL, OWF_OWF, OWF_SPILL, OWF_OUT,
    `OWF + CC_ALL`, `OWF + CC_OWF`, `OWF + CC_SPILL`, `OWF + CC_OUT`
  ) %>%
  mutate(indicator = as.character(indicator))

summary_table_pct
```

```{r}
# =========================================================
# TABLE: only absolute changes (+/- X)
# =========================================================

format_abs_change_cell <- function(x, digits = 2) {
  ifelse(
    is.na(x),
    "NA",
    paste0(
      ifelse(x > 0, "+", ""),
      formatC(x, format = "f", digits = digits)
    )
  )
}

summary_table_abs_change <- summary_vs_baseline %>%
  mutate(
    scenario = recode(scenario, "OWF_CC" = "OWF + CC"),
    abs_change_cell = format_abs_change_cell(abs_change)
  ) %>%
  unite(col_name, scenario, zone, sep = "_") %>%
  select(indicator, col_name, abs_change_cell) %>%
  pivot_wider(names_from = col_name, values_from = abs_change_cell) %>%
  select(
    indicator,
    CC_ALL, CC_OWF, CC_SPILL, CC_OUT,
    OWF_ALL, OWF_OWF, OWF_SPILL, OWF_OUT,
    `OWF + CC_ALL`, `OWF + CC_OWF`, `OWF + CC_SPILL`, `OWF + CC_OUT`
  ) %>%
  mutate(indicator = as.character(indicator))

summary_table_abs_change
```

```{r}
# =========================================================
# TABLE: absolute change + relative change
# =========================================================

format_abs_and_pct_cell <- function(abs_change, pct_change,
                                    abs_digits = 2,
                                    pct_digits = 1) {
  
  abs_txt <- ifelse(
    is.na(abs_change),
    "NA",
    paste0(
      ifelse(abs_change > 0, "+", ""),
      formatC(abs_change, format = "f", digits = abs_digits)
    )
  )
  
  pct_txt <- ifelse(
    is.na(pct_change),
    "NA",
    paste0(
      ifelse(pct_change > 0, "+", ""),
      formatC(pct_change, format = "f", digits = pct_digits),
      "%"
    )
  )
  
  paste0(abs_txt, " (", pct_txt, ")")
}

summary_table_abs_and_pct <- summary_vs_baseline %>%
  mutate(
    scenario = recode(scenario, "OWF_CC" = "OWF + CC"),
    cell = format_abs_and_pct_cell(abs_change, pct_change)
  ) %>%
  unite(col_name, scenario, zone, sep = "_") %>%
  select(indicator, col_name, cell) %>%
  pivot_wider(names_from = col_name, values_from = cell) %>%
  select(
    indicator,
    CC_ALL, CC_OWF, CC_SPILL, CC_OUT,
    OWF_ALL, OWF_OWF, OWF_SPILL, OWF_OUT,
    `OWF + CC_ALL`, `OWF + CC_OWF`, `OWF + CC_SPILL`, `OWF + CC_OUT`
  ) %>%
  mutate(indicator = as.character(indicator))

summary_table_abs_and_pct
```

Save tables
```{r}
# =========================================================
# EXPORT TABLES
# =========================================================

write_csv(summary_table_pct,
          file.path("Ecological_indicator_summary_table_pct_only.csv"))

write_csv(summary_table_abs_change,
          file.path("Ecological_indicator_summary_table_abs_change_only.csv"))

write_csv(summary_table_abs_and_pct,
          file.path("Ecological_indicator_summary_table_abs_and_pct.csv"))
```


Combining with ENA table
```{r}
# =========================================================
# COMBINE ECOLOGICAL + ENA TABLES
# Two sub-rows per indicator: absolute change and % change
# =========================================================

library(dplyr)
library(tidyr)
library(readr)
library(stringr)

# ---------------------------------------------------------
# 1. Read ENA numeric table
# ---------------------------------------------------------
ena_numeric <- read_csv("ENA_summary_table_numeric.csv")
```


```{r}
# ---------------------------------------------------------
# 2. Reshape ENA numeric table correctly
#    The file is effectively one row per indicator × source-zone,
#    with sparse values spread across wide columns.
#    We pivot longer and keep only non-NA values.
# ---------------------------------------------------------
ena_from_numeric <- ena_numeric %>%
  pivot_longer(
    cols = -c(indicator, baseline_abs_value),
    names_to = c("scenario", "zone", ".value"),
    names_pattern = "^(CC|OWF|OWF \\+ CC)_(ALL|OWF|SPILL|OUT)_(abs_value|abs_change|pct_change)$"
  ) %>%
  filter(
    !is.na(abs_value) | !is.na(abs_change) | !is.na(pct_change)
  ) %>%
  group_by(indicator, scenario, zone) %>%
  summarise(
    baseline_abs_value = first(na.omit(baseline_abs_value)),
    abs_value = first(na.omit(abs_value)),
    abs_change = first(na.omit(abs_change)),
    pct_change = first(na.omit(pct_change)),
    .groups = "drop"
  ) %>%
  mutate(
    indicator_group = "ENA"
  )

ena_from_numeric

# ---------------------------------------------------------
# 3. Prepare ecological indicators
#    Uses the object already present in your ecological script
# ---------------------------------------------------------
eco_master <- summary_vs_baseline %>%
  mutate(
    scenario = recode(scenario, "OWF_CC" = "OWF + CC"),
    indicator_group = case_when(
      indicator %in% c("Total B", "Fish B", "Invertebrates B",
                       "Commercial B", "Demersal B", "Pelagic B") ~ "Biomass",
      indicator %in% c("Total C", "Fish C", "Invertebrates C",
                       "Commercial C", "Demersal C", "Pelagic C") ~ "Catch",
      indicator %in% c("TL community", "TL catch", "TL2") ~ "Trophic level",
      TRUE ~ "Other"
    )
  ) %>%
  select(indicator_group, indicator, scenario, zone, abs_change, pct_change)

# ---------------------------------------------------------
# 4. Prepare ENA indicators
# ---------------------------------------------------------
ena_master <- ena_from_numeric %>%
  select(indicator_group, indicator, scenario, zone, abs_change, pct_change)

# ---------------------------------------------------------
# 5. Combine both
# ---------------------------------------------------------
combined_master <- bind_rows(eco_master, ena_master) %>%
  mutate(
    scenario = factor(scenario, levels = c("CC", "OWF", "OWF + CC")),
    zone = factor(zone, levels = c("ALL", "OWF", "SPILL", "OUT"))
  )
```


```{r}
# ---------------------------------------------------------
# 6. Formatting helpers
# ---------------------------------------------------------
format_abs_change <- function(x, digits = 2) {
  ifelse(
    is.na(x),
    "NA",
    paste0(
      ifelse(x > 0, "+", ""),
      formatC(x, format = "f", digits = digits)
    )
  )
}

format_pct_change <- function(x, digits = 1) {
  ifelse(
    is.na(x),
    "NA",
    paste0(
      ifelse(x > 0, "+", ""),
      formatC(x, format = "f", digits = digits),
      "%"
    )
  )
}
```


```{r}
# ---------------------------------------------------------
# 7. Build two-subrow table
# ---------------------------------------------------------
combined_long <- bind_rows(
  combined_master %>%
    mutate(
      change_type = "Δ",
      value_fmt = format_abs_change(abs_change, digits = 2)
    ),
  combined_master %>%
    mutate(
      change_type = "%",
      value_fmt = format_pct_change(pct_change, digits = 1)
    )
) %>%
  arrange(indicator_group, indicator, scenario, zone, change_type)

combined_table_two_rows <- combined_long %>%
  unite(col_name, scenario, zone, sep = "_") %>%
  select(indicator_group, indicator, change_type, col_name, value_fmt) %>%
  pivot_wider(
    names_from = col_name,
    values_from = value_fmt
  ) %>%
  select(
    indicator_group, indicator, change_type,
    CC_ALL, CC_OWF, CC_SPILL, CC_OUT,
    OWF_ALL, OWF_OWF, OWF_SPILL, OWF_OUT,
    `OWF + CC_ALL`, `OWF + CC_OWF`, `OWF + CC_SPILL`, `OWF + CC_OUT`
  )

combined_table_two_rows
```

```{r}
combined_table_two_rows_pretty <- combined_table_two_rows %>%
  group_by(indicator_group, indicator) %>%
  mutate(
    indicator_display = ifelse(row_number() == 1, as.character(indicator), "")
  ) %>%
  ungroup() %>%
  select(
    indicator_group,
    indicator_display,
    change_type,
    CC_ALL, CC_OWF, CC_SPILL, CC_OUT,
    OWF_ALL, OWF_OWF, OWF_SPILL, OWF_OUT,
    `OWF + CC_ALL`, `OWF + CC_OWF`, `OWF + CC_SPILL`, `OWF + CC_OUT`
  ) %>%
  rename(indicator = indicator_display)

combined_table_two_rows_pretty

write_csv(
  combined_table_two_rows_pretty,
  file.path("Combined_summary_table_two_rows_pretty.csv")
)
```

Now make a figure of this table
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(scales)

# ---------------------------------------------------------
# Prepare plotting data
# ---------------------------------------------------------
plot_df <- bind_rows(
  combined_master %>%
    mutate(
      change_type = "%",
      value_num = pct_change,
      label = ifelse(
        is.na(pct_change),
        "NA",
        paste0(
          ifelse(pct_change > 0, "+", ""),
          formatC(pct_change, format = "f", digits = 1),
          "%"
        )
      )
    ),
  combined_master %>%
    mutate(
      change_type = "Δ",
      value_num = abs_change,
      label = ifelse(
        is.na(abs_change),
        "NA",
        paste0(
          ifelse(abs_change > 0, "+", ""),
          formatC(abs_change, format = "f", digits = 2)
        )
      )
    )
) %>%
  mutate(
    scenario = factor(scenario, levels = c("CC", "OWF", "OWF + CC")),
    zone = factor(zone, levels = c("ALL", "OWF", "SPILL", "OUT")),
    col_id = paste(scenario, zone, sep = "_")
  )
```

```{r}
# ---------------------------------------------------------
# Row ordering
# ---------------------------------------------------------
indicator_order <- c(
  "Total B", "Fish B", "Invertebrates B", "Demersal B", "Pelagic B", "Commercial B",
  "Total C", "Fish C", "Invertebrate C", "Demersal C", "Pelagic C",
  "TL community", "TL community 3,25", "TL catch",
  "FCI", "H", "RDC", "SOI"
)

plot_df <- plot_df %>%
  mutate(
    indicator = factor(indicator, levels = indicator_order),
    change_type = factor(change_type, levels = c("%", "Δ"))
  ) %>%
  arrange(indicator, change_type) %>%
  mutate(
    row_id = paste(indicator, change_type, sep = " | ")
  )

row_levels <- plot_df %>%
  distinct(indicator, change_type, row_id) %>%
  arrange(indicator, change_type) %>%
  pull(row_id)

plot_df <- plot_df %>%
  mutate(
    row_id = factor(row_id, levels = rev(row_levels))
  )
# Fill only the % rows
plot_df <- plot_df %>%
  mutate(
    fill_value = ifelse(change_type == "%", value_num, NA_real_)
  )
```

fill color cap
```{r}
fill_limit <- max(abs(plot_df$fill_value), na.rm = TRUE)

# OR

fill_limit <- 25

plot_df <- plot_df %>%
  mutate(
    fill_value_capped = pmax(pmin(fill_value, fill_limit), -fill_limit)
  )
```

```{r}
p_table_heatmap <- ggplot(plot_df, aes(x = col_id, y = row_id)) +
  geom_tile(aes(fill = fill_value_capped), color = "grey70", linewidth = 0.3) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(
    low = "#E67E22",
    mid = "white",
    high = "#2C7FB8",
    midpoint = 0,
    limits = c(-fill_limit, fill_limit),
    oob = squish,
    na.value = "grey95",
    name = "% change"
  ) +
  scale_x_discrete(
    labels = c(
      "CC_ALL" = "ALL", "CC_OWF" = "OWF", "CC_SPILL" = "SPILL", "CC_OUT" = "OUT",
      "OWF_ALL" = "ALL", "OWF_OWF" = "OWF", "OWF_SPILL" = "SPILL", "OWF_OUT" = "OUT",
      "OWF + CC_ALL" = "ALL", "OWF + CC_OWF" = "OWF", "OWF + CC_SPILL" = "SPILL", "OWF + CC_OUT" = "OUT"
    )
  ) +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(size = 10, face = "bold"),
    axis.text.y = element_text(size = 10),
    legend.position = "right"
  )

p_table_heatmap
p_table_heatmap +
  annotate("text", x = 2.5, y = length(row_levels) + 1.2, label = "Climate change", fontface = "bold", size = 4) +
  annotate("text", x = 6.5, y = length(row_levels) + 1.2, label = "Offshore wind farms", fontface = "bold", size = 4) +
  annotate("text", x = 10.5, y = length(row_levels) + 1.2, label = "Offshore wind farms + climate change", fontface = "bold", size = 4) +
  geom_vline(xintercept = c(4.5, 8.5), linewidth = 0.7) + coord_cartesian(clip = "off") +
  theme(plot.margin = margin(20, 20, 20, 20))
```
Plot only % change
```{r}
# =========================================================
# HEATMAP FIGURE: percentage change only
# Rows = indicators
# Columns = scenario x zone
# =========================================================

library(dplyr)
library(ggplot2)
library(scales)

# ---------------------------------------------------------
# 1. Prepare plotting dataframe
# ---------------------------------------------------------
plot_pct <- combined_master %>%
  mutate(
    scenario = factor(scenario, levels = c("CC", "OWF", "OWF + CC")),
    zone = factor(zone, levels = c("ALL", "OWF", "SPILL", "OUT")),
    col_id = factor(
      paste(scenario, zone, sep = "_"),
      levels = c(
        "CC_ALL", "CC_OWF", "CC_SPILL", "CC_OUT",
        "OWF_ALL", "OWF_OWF", "OWF_SPILL", "OWF_OUT",
        "OWF + CC_ALL", "OWF + CC_OWF", "OWF + CC_SPILL", "OWF + CC_OUT"
      )
    ),
    label = ifelse(
      is.na(pct_change),
      "NA",
      paste0(
        ifelse(pct_change > 0, "+", ""),
        formatC(pct_change, format = "f", digits = 1),
        "%"
      )
    )
  )

# ---------------------------------------------------------
# 2. Set indicator order
# Adjust if your exact names differ slightly
# ---------------------------------------------------------
indicator_order <- c(
  "Total B", "Fish B", "Invertebrates B", "Demersal B", "Pelagic B", "Commercial B",
  "Total C", "Fish C", "Invertebrate C", "Demersal C", "Pelagic C",
  "TL community", "TL community 3,25", "TL catch",
  "FCI", "H", "RDC", "SOI"
)

plot_pct <- plot_pct %>%
  mutate(
    indicator = factor(indicator, levels = indicator_order)
  ) %>%
  arrange(indicator) %>%
  mutate(
    row_id = factor(as.character(indicator), levels = rev(indicator_order))
  )

# ---------------------------------------------------------
# 3. Add indicator groups for separators / optional styling
# ---------------------------------------------------------
plot_pct <- plot_pct %>%
  mutate(
    indicator_group = case_when(
      indicator %in% c("Total B", "Fish B", "Invertebrates B", "Demersal B", "Pelagic B", "Commercial B") ~ "Biomass",
      indicator %in% c("Total C", "Fish C", "Invertebrate C", "Demersal C", "Pelagic C") ~ "Catch",
      indicator %in% c("TL community", "TL community 3,25", "TL catch") ~ "Trophic level",
      indicator %in% c("FCI", "H", "RDC", "SOI") ~ "ENA",
      TRUE ~ "Other"
    )
  )

# ---------------------------------------------------------
# 4. Determine asymmetric color caps
# lower cap = strongest decrease excluding -100
# upper cap = strongest increase
# ---------------------------------------------------------
neg_cap <- plot_pct %>%
  filter(!is.na(pct_change), pct_change != -100) %>%
  summarise(cap = min(pct_change)) %>%
  pull(cap)

pos_cap <- plot_pct %>%
  filter(!is.na(pct_change)) %>%
  summarise(cap = max(pct_change)) %>%
  pull(cap)

# safety fallback
if (!is.finite(neg_cap)) neg_cap <- -1
if (!is.finite(pos_cap)) pos_cap <- 1

# cap values for plotting
plot_pct <- plot_pct %>%
  mutate(
    fill_value = case_when(
      is.na(pct_change) ~ NA_real_,
      pct_change < neg_cap ~ neg_cap,
      pct_change > pos_cap ~ pos_cap,
      TRUE ~ pct_change
    ),
    fontface_lab = ifelse(!is.na(pct_change) & pct_change == -100, "bold", "plain")
  )

# ---------------------------------------------------------
# 5. Separator positions
# ---------------------------------------------------------
# horizontal separators between groups
hline_df <- data.frame(
  y = c(12.5, 7.5, 4.5)   # after Biomass, Catch, Trophic level
)

# ---------------------------------------------------------
# 6. Plot
# ---------------------------------------------------------
p_pct_heatmap <- ggplot(plot_pct, aes(x = col_id, y = row_id)) +
  geom_tile(aes(fill = fill_value), color = "grey70", linewidth = 0.35) +
  geom_text(
    aes(label = label, fontface = fontface_lab),
    size = 3.2
  ) +
  scale_fill_gradientn(
    colours = c("#E67E22", "white", "#2C7FB8"),
    values = scales::rescale(c(neg_cap, 0, pos_cap)),
    limits = c(neg_cap, pos_cap),
    oob = scales::squish,
    na.value = "grey95",
    name = "% change"
  ) +
  scale_x_discrete(
    labels = c(
      "CC_ALL" = "ALL", "CC_OWF" = "OWF", "CC_SPILL" = "SPILL", "CC_OUT" = "OUT",
      "OWF_ALL" = "ALL", "OWF_OWF" = "OWF", "OWF_SPILL" = "SPILL", "OWF_OUT" = "OUT",
      "OWF + CC_ALL" = "ALL", "OWF + CC_OWF" = "OWF", "OWF + CC_SPILL" = "SPILL", "OWF + CC_OUT" = "OUT"
    ),
    expand = c(0, 0)
  ) +
  scale_y_discrete(expand = c(0, 0)) +
  geom_vline(xintercept = c(4.5, 8.5), linewidth = 0.8) +
  geom_hline(data = hline_df, aes(yintercept = y), inherit.aes = FALSE, linewidth = 0.8) +
  annotate(
    "text", x = 2.5, y = length(indicator_order) + 1.1,
    label = "Climate change", fontface = "bold", size = 4
  ) +
  annotate(
    "text", x = 6.5, y = length(indicator_order) + 1.1,
    label = "Offshore wind farms", fontface = "bold", size = 4
  ) +
  annotate(
    "text", x = 10.5, y = length(indicator_order) + 1.1,
    label = "Offshore wind farms + climate change", fontface = "bold", size = 4
  ) +
  coord_cartesian(clip = "off") +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(size = 10, face = "bold"),
    axis.text.y = element_text(size = 10),
    legend.position = "right",
    plot.margin = margin(30, 20, 20, 20)
  )

p_pct_heatmap

ggsave("Indicator_Summary_heatmap_1.png", p_pct_heatmap,widt = 9, dpi = 800)
```
Integrated figure
```{r}
library(dplyr)
library(ggplot2)
library(scales)
library(cowplot)
```
Prepare the plotting dataframe once
```{r}
# =========================================================
# PREPARE DATA FOR GROUPED HEATMAP FIGURES
# =========================================================

plot_pct <- combined_master %>%
  mutate(
    scenario = factor(scenario, levels = c("CC", "OWF", "OWF + CC")),
    zone = factor(zone, levels = c("ALL", "OWF", "SPILL", "OUT")),
    col_id = factor(
      paste(scenario, zone, sep = "_"),
      levels = c(
        "CC_ALL", "CC_OWF", "CC_SPILL", "CC_OUT",
        "OWF_ALL", "OWF_OWF", "OWF_SPILL", "OWF_OUT",
        "OWF + CC_ALL", "OWF + CC_OWF", "OWF + CC_SPILL", "OWF + CC_OUT"
      )
    ),
    label = ifelse(
      is.na(pct_change),
      "NA",
      paste0(
        ifelse(pct_change > 0, "+", ""),
        formatC(pct_change, format = "f", digits = 1),
        "%"
      )
    ),
    indicator_group = case_when(
      indicator %in% c("Total B", "Fish B", "Invertebrates B", "Demersal B", "Pelagic B", "Commercial B") ~ "Biomass",
      indicator %in% c("Total C", "Fish C", "Invertebrate C", "Demersal C", "Pelagic C") ~ "Catch",
      indicator %in% c("TL community", "TL community 3,25", "TL catch") ~ "Trophic level",
      indicator %in% c("FCI", "H", "RDC", "SOI") ~ "Ecological Network Analysis",
      TRUE ~ "Other"
    )
  )
```

Define the row order within groups
```{r}
group_orders <- list(
  "Biomass" = c("Total B", "Fish B", "Invertebrates B", "Demersal B", "Pelagic B", "Commercial B"),
  "Catch" = c("Total C", "Fish C", "Invertebrate C", "Demersal C", "Pelagic C"),
  "Trophic level" = c("TL community", "TL community 3,25", "TL catch"),
  "Ecological Network Analysis" = c("FCI", "H", "RDC", "SOI")
)
```

Plotting function
```{r}
make_pct_heatmap_group <- function(data, group_name, show_top_axis = TRUE, show_legend = TRUE) {
  
  x_levels <- c(
  "CC_ALL", "CC_OWF", "CC_SPILL", "CC_OUT",
  "OWF_ALL", "OWF_OWF", "OWF_SPILL", "OWF_OUT",
  "OWF + CC_ALL", "OWF + CC_OWF", "OWF + CC_SPILL", "OWF + CC_OUT"
)
  
  df <- data %>%
  filter(indicator_group == group_name) %>%
  mutate(
    indicator = factor(indicator, levels = group_orders[[group_name]]),
    row_id = factor(as.character(indicator), levels = rev(group_orders[[group_name]])),
    col_id = factor(as.character(col_id), levels = x_levels)
  )
  
  neg_cap <- df %>%
    filter(!is.na(pct_change), pct_change != -100) %>%
    summarise(cap = min(pct_change)) %>%
    pull(cap)
  
  pos_cap <- df %>%
    filter(!is.na(pct_change)) %>%
    summarise(cap = max(pct_change)) %>%
    pull(cap)
  
  if (!is.finite(neg_cap)) neg_cap <- -1
  if (!is.finite(pos_cap)) pos_cap <- 1
  
df <- df %>%
  mutate(
    is_minus100 = !is.na(pct_change) & pct_change == -100,
    fill_value = case_when(
      is.na(pct_change) ~ NA_real_,
      pct_change == -100 ~ NA_real_,
      pct_change < neg_cap ~ neg_cap,
      pct_change > pos_cap ~ pos_cap,
      TRUE ~ pct_change
    ),
    fontface_lab = ifelse(is_minus100, "bold", "plain")
  )

p <- ggplot(df, aes(x = col_id, y = row_id)) +
  geom_tile(
    data = df %>% filter(is_minus100),
    fill = "grey70",
    color = "grey70",
    linewidth = 0.35
  ) +
  geom_tile(
    data = df %>% filter(!is_minus100),
    aes(fill = fill_value),
    color = "grey70",
    linewidth = 0.35
  ) +
  geom_text(aes(label = label, fontface = fontface_lab), size = 3.2) +
  scale_fill_gradientn(
    colours = c("#E67E22", "white", "#2C7FB8"),
    values = scales::rescale(c(neg_cap, 0, pos_cap)),
    limits = c(neg_cap, pos_cap),
    oob = scales::squish,
    na.value = "grey95",
    name = "% change"
  ) +
    scale_x_discrete(
  limits = x_levels,
  labels = c(
    "CC_ALL" = "ALL", "CC_OWF" = "OWF", "CC_SPILL" = "SPILL", "CC_OUT" = "OUT",
    "OWF_ALL" = "ALL", "OWF_OWF" = "OWF", "OWF_SPILL" = "SPILL", "OWF_OUT" = "OUT",
    "OWF + CC_ALL" = "ALL", "OWF + CC_OWF" = "OWF", "OWF + CC_SPILL" = "SPILL", "OWF + CC_OUT" = "OUT"
  ),
  expand = c(0, 0),
  position = "top"
) +
    scale_y_discrete(expand = c(0, 0)) +
    geom_vline(xintercept = c(4.5, 8.5), linewidth = 0.8) +
    labs(x = NULL, y = NULL, title = group_name) +
    theme_minimal(base_size = 11) +
    theme(
      panel.grid = element_blank(),
      plot.title = element_text(face = "bold", hjust = 0, size = 12),
      axis.text.y = element_text(size = 10),
      axis.text.x.top = element_text(size = 10, face = "bold"),
      axis.ticks.x.top = element_blank(),
      axis.text.x.bottom = element_blank(),
      axis.ticks.x.bottom = element_blank(),
      plot.margin = margin(8, 20, 10, 20),
      legend.position = "right",
      legend.title = element_text(size = 10),
      legend.text = element_text(size = 9)
    )
  
  if (!show_legend) {
    p <- p + theme(legend.position = "none")
  } else {
    p <- p +
      guides(
        fill = guide_colorbar(
          barheight = unit(2.8, "cm"),
          barwidth = unit(0.45, "cm"),
          ticks.colour = "black",
          frame.colour = "black"
        )
      )
  }
  
  p
}
```

Separate scenario headers
```{r}
library(ggplot2)

scenario_header_plot <- ggplot() +
  annotate("text", x = 2.5, y = 1, label = "CC", fontface = "bold", size = 5) +
  annotate("text", x = 6.5, y = 1, label = "OWF", fontface = "bold", size = 5) +
  annotate("text", x = 10.5, y = 1, label = "OWF + CC", fontface = "bold", size = 5) +
  geom_vline(xintercept = c(4.5, 8.5), linewidth = 0.8) +
  scale_x_continuous(limits = c(0.5, 12.5), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0.5, 1.5), expand = c(0, 0)) +
  theme_void() +
  theme(
    plot.margin = margin(0, 20, -5, 20)
  )

scenario_header_plot
```


Build the 4 plots
```{r}
p_biomass <- make_pct_heatmap_group(plot_pct, "Biomass", show_top_axis = TRUE, show_legend = FALSE)
p_catch   <- make_pct_heatmap_group(plot_pct, "Catch", show_top_axis = TRUE, show_legend = FALSE)
p_tl      <- make_pct_heatmap_group(plot_pct, "Trophic level", show_top_axis = TRUE, show_legend = FALSE)
p_ena     <- make_pct_heatmap_group(plot_pct, "Ecological Network Analysis", show_top_axis = TRUE, show_legend = FALSE)
```

Align and combine them
```{r}
library(cowplot)

aligned_plots <- align_plots(
  scenario_header_plot,
  p_biomass, p_catch, p_tl, p_ena,
  align = "v",
  axis = "lr"
)

combined_heatmap <- plot_grid(
  aligned_plots[[1]],
  aligned_plots[[2]],
  aligned_plots[[3]],
  aligned_plots[[4]],
  aligned_plots[[5]],
  ncol = 1,
  rel_heights = c(0.55, 6, 5, 3, 4)
)

combined_heatmap

ggsave(
  filename = file.path(out_dir, "Pct_change_heatmap_grouped.png"),
  plot = combined_heatmap,
  width = 8.5,
  height = 9,
  dpi = 800,
  bg = "white"
)
```

