---
title: "Input_layers_processing"
author: "Dries Lorré"
date: "2026-02-09"
output: html_document
---

NOTE: all data should be stored as ESRI ASCII raster maps after processing. This will allow both smooth Ecospace base map and STDF input

Packages
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(terra)
```

Input layers (paths)
```{r}
# Base folder (as before)
base_dir  <- "C:/Users/dlorr/OneDrive - UGent/Documents/PhD/EwE_OWF/Spatial data"

path_sbns <- file.path(base_dir, "SBNS_boundary_epsg3857.shp")
path_temp <- file.path(base_dir, "Temperature")
path_depth <- file.path(base_dir, "Depth")

out_dir <- file.path(base_dir, "Processed")
if (!dir.exists(out_dir)) dir.create(out_dir)
```

PP - template (for raster structure - resolution - CRS)
```{r}
# Study area polygon
sbns <- vect(path_sbns)

temp_files <- list.files(path_temp, pattern = "\\.nc$", full.names = TRUE)
if (length(temp_files) == 0) {
  stop("No NetCDF files found in the Temperature folder.")
}

# Use the first SST file as grid template
temp_template <- rast(temp_files[1])
temp_template
crs(temp_template)
res(temp_template)
ext(temp_template)

# Reproject SBNS to SST CRS if needed
if (crs(sbns) != crs(temp_template)) {
  sbns <- project(sbns, temp_template)
}

# Crop + mask SST to SBNS; this defines the *working grid* for the study area
temp_sbns <- crop(temp_template, sbns)
temp_sbns <- mask(temp_sbns, sbns)

temp_sbns
res(temp_sbns)   # this is the resolution your base map will use
ext(temp_sbns)

# Base grid = first SST layer (surface temperature monthly mean)
temp_grid <- temp_sbns[[1]]
names(temp_grid) <- "SST"
plot(temp_grid, main = "SST (base grid) over SBNS")
```
Mosaic and project depth to SST grid
```{r}
# --- Depth tiles ---
depth_files <- list.files(path_depth, pattern = "\\.tif$", full.names = TRUE)
if (length(depth_files) == 0) {
  stop("No .tif depth tiles found in the Depth folder.")
}

# Read each tile separately
depth_list <- lapply(depth_files, rast)

# Mosaic all tiles into one bathymetry raster
# (assumes same CRS & resolution; extents may differ/adjacent)
depth_mosaic <- do.call(mosaic, c(depth_list, fun = "mean"))
depth_mosaic
crs(depth_mosaic)
res(depth_mosaic)
ext(depth_mosaic)

# Quick plot to check the full “puzzle”
plot(depth_mosaic, main = "Mosaic of depth tiles")

# Project + resample bathymetry onto SST grid
depth_sst <- project(depth_mosaic, temp_grid, method = "bilinear")

# Restrict to SBNS polygon
depth_sst <- crop(depth_sst, sbns)
depth_sst <- mask(depth_sst, sbns)

names(depth_sst) <- "depth"
plot(depth_sst, main = "Bathymetry on SST grid")
depth_range <- as.vector(global(depth_sst, fun = range, na.rm = TRUE))
depth_range
```
Align depth and SST + remove islands
```{r}
## 1) SST main connected sea body (no islands) - as you already have

# SST base grid (first layer) – you already have this
# temp_grid <- temp_sbns[[1]]
# names(temp_grid) <- "SST"

# Mask where SST has data (1) vs no data (NA)
sst_mask <- !is.na(temp_grid)
sst_mask[sst_mask == 0] <- NA
names(sst_mask) <- "sst_sea"

plot(sst_mask, main = "SST mask (1 = sea cells with SST data)")

# Connected patches (4-neighbour, as you set)
sst_clumps <- patches(sst_mask, directions = 4)

# Frequency of patch IDs
clump_freq <- freq(sst_clumps)
clump_freq

# ID of the largest connected sea patch
main_id <- clump_freq$value[which.max(clump_freq$count)]
main_id

# Mask for the main connected SST sea body
sst_main_mask <- sst_clumps == main_id
sst_main_mask[sst_main_mask == 0] <- NA
names(sst_main_mask) <- "sst_main"

plot(sst_main_mask, main = "Main connected SST sea patch")
```


```{r}
## 2) Depth on SST grid, absolute values, restricted to main SST patch

# depth_sst: you already computed this
# depth_sst <- project(depth_mosaic, temp_grid, method = "bilinear")
# depth_sst <- crop(depth_sst, sbns)
# depth_sst <- mask(depth_sst, sbns)
# names(depth_sst) <- "depth"

# Convert depth to absolute values (all bathymetry positive)
depth_sst_abs <- abs(depth_sst)
names(depth_sst_abs) <- "depth_m"

# Optional: quick check
global(depth_sst_abs, fun = range, na.rm = TRUE)

# Sea mask just for plotting (non-NA = sea)
sea_mask <- !is.na(depth_sst_abs)
sea_mask[sea_mask == 0] <- NA
names(sea_mask) <- "sea"

plot(sea_mask, main = "Depth-based sea mask (1 = non-NA depth)")


# Enforce SST-based main patch on depth
# -> guarantees that the same main body is used for all env layers
depth_base <- mask(depth_sst_abs, sst_main_mask)
names(depth_base) <- "depth_m"

plot(depth_base, main = "Depth (abs, SST grid, main SBNS patch only)")
global(depth_base, fun = range, na.rm = TRUE)
```

DEPTH
```{r}
## 3) Final “Depth” layer for your software: positive depth, 0 = land/inactive

# Depth with 0 = land/inactive
depth_base_0 <- depth_base
depth_base_0[is.na(depth_base_0)] <- 0
names(depth_base_0) <- "depth_m"

plot(depth_base_0, main = "Depth (m, abs) on SST grid; 0 = land/inactive")

# Save outputs
writeRaster(
  depth_base_0,
  filename  = file.path(out_dir, "Depth.asc"),
  overwrite = TRUE
)

writeRaster(
  depth_base_0,
  filename  = file.path(out_dir, "Depth_SBNS_Basegrid.tif"),
  overwrite = TRUE
)

```

PRIMARY PRODUCTION
```{r}
path_pp <- file.path(base_dir, "PrimaryProduction_1km_98-25")

pp_files <- list.files(path_pp, pattern = "\\.nc$", full.names = TRUE)
length(pp_files)

# Helper to extract year and month from filenames like:
# 19980101-19980131_cmems_obs-oc_atl_bgc-pp_my_l4-multi-1km_P1M.nc
get_year_month <- function(f) {
  b <- basename(f)
  year <- as.integer(substr(b, 1, 4))
  month <- as.integer(substr(b, 5, 6))
  list(year = year, month = month)
}

ym <- lapply(pp_files, get_year_month)
years  <- sapply(ym, `[[`, "year")
months <- sapply(ym, `[[`, "month")

# Keep only Jan 1998 – Dec 2024
keep <- (years > 1998 | (years == 1998 & months >= 1)) &
        (years < 2025 | (years == 2024 & months <= 12))

pp_files_sel <- pp_files[keep]
length(pp_files_sel)
```
Template for pp calculation
```{r}
# Template grid = depth_base (geometry of base layer)
template <- depth_base
template
res(template)
ext(template)

# Init accumulators with zeros
sum_PP   <- template
count_PP <- template

values(sum_PP)   <- 0
values(count_PP) <- 0
```

```{r}
terraOptions(progress = 1)  # show progress in long loops

for (f in pp_files_sel) {
  cat("Processing:", basename(f), "\n")
  
  # Read only the PP subdataset (mgC m^-2 day^-1)
  # (If this errors, you can try sds(f) to list subdatasets)
  pp_raw <- rast(f, subds = "PP")
  
  # Project PP onto the base grid (SST/depth grid)
  pp_proj <- project(pp_raw, template, method = "bilinear")
  
  # Restrict to active cells (same as depth_base extent & NA pattern)
  pp_proj <- mask(pp_proj, template)
  
  # Mask of non-NA PP values
  non_na <- !is.na(pp_proj)
  
  # Replace NA with 0 so they don't affect the sum
  pp_proj_na0 <- pp_proj
  pp_proj_na0[is.na(pp_proj_na0)] <- 0
  
  # Accumulate sum of PP
  sum_PP <- sum_PP + pp_proj_na0
  
  # Accumulate count of months with valid PP
  # TRUE -> 1, FALSE -> 0, so summing works
  count_PP <- count_PP + non_na
}

sum_PP
count_PP
```
Compute mean PP between 1998 and 2024
```{r}
# Avoid division by zero / no-data cells
mean_PP <- sum_PP / count_PP
mean_PP[count_PP == 0] <- NA

names(mean_PP) <- "PP_mean_mg_m2_d"

plot(mean_PP, main = "Mean PP (1998–2024) [mgC m^-2 d^-1]")
global(mean_PP, fun = range, na.rm = TRUE)


```

Convert to mgC/m³/day
```{r}
# Sanity
global(depth_base, fun = range, na.rm = TRUE)

# Volume-based PP: mgC m^-3 day^-1
# PP_vol <- mean_PP / depth_base
PP_vol <- mean_PP / 10

# Where depth <= 0 or depth is NA, set PP to NA
# PP_vol[is.na(depth_base) | depth_base <= 0] <- NA

names(PP_vol) <- "PP_mean_mg_m3_d"

# --- NEW: cap values at 3000 ---
# PP_vol <- clamp(PP_vol, upper = 300)

plot(PP_vol, main = "Mean PP (1998–2024) [mgC m^-3 d^-1]")
global(PP_vol, fun = range, na.rm = TRUE)

# Volume-based mean PP for the software
writeRaster(
  PP_vol,
  filename  = file.path(out_dir, "PP_mean_1998_2024_mgC_m3_d.tif"),
  overwrite = TRUE
)

writeRaster(
  PP_vol,
  filename  = file.path(out_dir, "PP_mean_1998_2024_mgC_m3_d.asc"),
  overwrite = TRUE
)
```
Ecospace Base Relative Primary Production
```{r}
# Correct: get numeric mean
mean_PP <- sum_PP / count_PP
mean_PP[count_PP == 0] <- NA  # treat “no data” as NA
names(mean_PP) <- "PP_mean_mg_m2_d"

PP_vol <- mean_PP / 10
names(PP_vol) <- "PP_mean_mg_m3_d"
plot(PP_vol)

pp_mean <- global(PP_vol, "mean", na.rm = TRUE)$mean
RPP_mean <- PP_vol / pp_mean
names(RPP_mean) <- "RPP_mean"

RPP_mean_0 <- RPP_mean
RPP_mean_0[is.na(RPP_mean_0)] <- 0

plot(RPP_mean_0)
# Check if values make sense:
plot(RPP_mean > 1, main = "Cells with above-average PP")

writeRaster(
  RPP_mean_0,
  filename = file.path(out_dir, "RelativePimProd_basegrid.tif"),
  overwrite = TRUE
)

writeRaster(
  RPP_mean_0,
  filename = file.path(out_dir, "RelativePimProd_base_mean.asc"),
  overwrite = TRUE
)
```

Anomaly of PP (Nutrients)
```{r}
library(terra)
library(dplyr)
library(stringr)

# 1) Read anomaly CSV
anom_raw <- read.csv("PP_anomaly.csv", stringsAsFactors = FALSE)

# 2) Identify which rows are actual years vs months
months_vec <- c(
  "January", "February", "March", "April", "May", "June",
  "July", "August", "September", "October", "November", "December"
)

anom_years <- anom_raw %>%
  filter(!Name %in% months_vec) %>%          # keep only year rows
  mutate(
    Year        = as.integer(Name),
    anom_factor = PO4_y_copernicus_2        # multiplicative factor
  ) %>%
  arrange(Year)

# Check range
range(anom_years$Year)
head(anom_years)

# Factor for 1991
f_1991 <- anom_years$anom_factor[anom_years$Year == 1991]

if (length(f_1991) != 1 || is.na(f_1991)) {
  stop("Could not find a unique anomaly for year 1991.")
}

anom_years <- anom_years %>%
  mutate(rel_factor = anom_factor / f_1991)

# Sanity: should be 1
anom_years$rel_factor[anom_years$Year == 1991]

# Otherwise, reload from file:
RPP_base <- rast(file.path(out_dir, "RelativePimProd_base_mean.asc"))
names(RPP_base) <- "RPP_base"

# Optional quick check
plot(RPP_base, main = "Baseline Relative PP (climatology)")
global(RPP_base, "mean", na.rm = TRUE)
```


```{r}
# ---- 1. Years to produce: FULL model horizon ----
years_vec <- 1991:2150

# Last year that has an anomaly in the CSV
max_year_in_csv <- max(anom_years$Year, na.rm = TRUE)
last_rel_factor <- anom_years$rel_factor[anom_years$Year == max_year_in_csv]

if (length(last_rel_factor) != 1 || is.na(last_rel_factor)) {
  stop("Could not determine a unique last_rel_factor from anomaly data.")
}

# ---- 2. Output dir ----
out_rpp_dir <- file.path(out_dir, "RPP_relative_yearly")
if (!dir.exists(out_rpp_dir)) dir.create(out_rpp_dir, recursive = TRUE)

# ---- 3. Loop over all years 1991–2150 ----
for (yr in years_vec) {
  
  if (yr %in% anom_years$Year) {
    # Year covered by CSV: use its own rel_factor
    f_rel <- anom_years$rel_factor[anom_years$Year == yr]
  } else {
    # Year beyond CSV range: reuse last available factor (e.g. 2025)
    f_rel <- last_rel_factor
  }
  
  # Safety check
  if (length(f_rel) != 1 || is.na(f_rel)) {
    warning("Year ", yr, ": rel_factor missing or ambiguous, skipping.")
    next
  }
  
  # Multiply baseline map by this year's factor
  r_year <- RPP_base * f_rel
  
  # NA → 0 for Ecospace
  r_year_0 <- r_year
  r_year_0[is.na(r_year_0)] <- 0
  names(r_year_0) <- paste0("RPP_rel_", yr)
  
  # Write as ASCII (and optional GeoTIFF)
  out_asc <- file.path(out_rpp_dir, paste0("RPP_rel_", yr, ".asc"))
  out_tif <- file.path(out_rpp_dir, paste0("RPP_rel_", yr, ".tif"))
  
  writeRaster(r_year_0, filename = out_asc, overwrite = TRUE)
  writeRaster(r_year_0, filename = out_tif, overwrite = TRUE)
  
  cat("Written:", basename(out_asc), " (factor =", f_rel, ")\n")
}
```



DISTANCE TO LAND
```{r}
# First build land mark
# depth_base_0: >0 = sea, 0 = land/inactive
land_mask <- depth_base_0

# Sea cells (depth > 0) -> NA
land_mask[land_mask > 0] <- NA

# Land / inactive cells (0) -> 1
land_mask[land_mask == 0] <- 1

names(land_mask) <- "land"

plot(land_mask, main = "Land mask (1 = land, NA = sea)")

# Project land mask to a metric CRS (for accurate distances)
land_mask_proj <- project(land_mask, "EPSG:3035", method = "near")

crs(land_mask_proj)
res(land_mask_proj)
plot(land_mask_proj, main = "Land mask in EPSG:3035")

# Distance to land
# Distance from sea (NA) to nearest land cell (non-NA)
dist_m_proj <- distance(land_mask_proj)

names(dist_m_proj) <- "dist_m_to_coast"
plot(dist_m_proj, main = "Distance to coast [m] (projected)")

# Back to base grid (lon/lat + SBNS extent)
dist_m <- project(dist_m_proj, depth_base, method = "bilinear")
names(dist_m) <- "dist_m_to_coast"

plot(dist_m, main = "Distance to coast [m] on base grid")

# To km
dist_km <- dist_m / 1000
names(dist_km) <- "dist_km_to_coast"

plot(dist_km, main = "Distance to coast [km] on base grid")
global(dist_km, fun = range, na.rm = TRUE)

# Version 1: NA on land/inactive, distance (km) on sea
dist_km_sea <- dist_km
dist_km_sea[depth_base_0 == 0] <- NA
names(dist_km_sea) <- "dist_km_to_coast"

plot(dist_km_sea, main = "Distance to coast [km] (sea cells only)")

# Version 2 (for software): 0 on land/inactive, distance (km) on sea
dist_km_0 <- dist_km
dist_km_0[depth_base_0 == 0] <- 0
names(dist_km_0) <- "dist_km_to_coast"

plot(dist_km_0, main = "Distance to coast [km]; 0 = land/inactive")


# Store data
# GeoTIFF
writeRaster(
  dist_km_0,
  filename  = file.path(out_dir, "DistCoast_SBNS_Basegrid_km.tif"),
  overwrite = TRUE
)

# ESRI ASCII (if you want it as model input)
writeRaster(
  dist_km_0,
  filename  = file.path(out_dir, "DistCoast_km.asc"),
  overwrite = TRUE
)
```

COASTAL - MARINE
```{r}
# 12 nautical miles in km
nm_to_km    <- 1.852
coastal_thr <- 12 * nm_to_km  # 22.224 km
coastal_thr

# If needed: rebuild dist_km with NA on land
dist_km <- dist_km_0
dist_km[depth_base_0 == 0] <- NA
names(dist_km) <- "dist_km_to_coast"

sea_cells <- depth_base_0 > 0   # TRUE on active sea cells

template <- depth_base_0
values(template) <- 0

coastal_cells <- sea_cells & (dist_km <= coastal_thr)
marine_cells  <- sea_cells & (dist_km >  coastal_thr)

# Quick sanity check counts
sum(coastal_cells, na.rm = TRUE)
sum(marine_cells,  na.rm = TRUE)


# SEABIRDS
seabirds_cm <- template  # starts at 0 everywhere

seabirds_cm[coastal_cells] <- 1
seabirds_cm[marine_cells]  <- 0.75

names(seabirds_cm) <- "coastal_marine_seabirds"

plot(seabirds_cm, main = "Coastal–marine (seabirds)\n1 = coastal, 0.75 = marine, 0 = land")


# CRANGON
crangon_cm <- template  # 0 everywhere by default

crangon_cm[coastal_cells] <- 1
# marine_cells remain 0

names(crangon_cm) <- "coastal_marine_crangon"

plot(crangon_cm, main = "Coastal–marine (crangon)\n1 = coastal, 0 = marine/land")

# MEIOFAUNA
meiofauna_cm <- template  # 0 everywhere

meiofauna_cm[coastal_cells] <- 1
meiofauna_cm[marine_cells]  <- 0.5

names(meiofauna_cm) <- "coastal_marine_meiofauna"

plot(meiofauna_cm, main = "Coastal–marine (meiofauna)\n1 = coastal, 0.5 = marine, 0 = land")

# Seabirds
writeRaster(
  seabirds_cm,
  filename  = file.path(out_dir, "CoastalMarine_Seabirds.tif"),
  overwrite = TRUE
)
writeRaster(
  seabirds_cm,
  filename  = file.path(out_dir, "CoastalMarine_Seabirds.asc"),
  overwrite = TRUE
)

# Crangon
writeRaster(
  crangon_cm,
  filename  = file.path(out_dir, "CoastalMarine_Crangon.tif"),
  overwrite = TRUE
)
writeRaster(
  crangon_cm,
  filename  = file.path(out_dir, "CoastalMarine_Crangon.asc"),
  overwrite = TRUE
)

# Meiofauna
writeRaster(
  meiofauna_cm,
  filename  = file.path(out_dir, "CoastalMarine_Meiofauna.tif"),
  overwrite = TRUE
)
writeRaster(
  meiofauna_cm,
  filename  = file.path(out_dir, "CoastalMarine_Meiofauna.asc"),
  overwrite = TRUE
)
```


Habitats
(make sure all are 1, this will make OWF ingestion later on much easier)
```{r}
path_folk <- file.path(
  base_dir,
  "SedimentHabitats",
  "seabed_substrate_250k.shp"
)

folk <- vect(path_folk)
plot(folk)
folk
crs(folk)
names(folk)

# reproject and clip
# Make sure CRS matches the base grid (depth/SST grid)
if (crs(folk) != crs(depth_base_0)) {
  folk <- project(folk, depth_base_0)
}

# Optional: crop to SBNS or base grid extent to speed things up
folk_sbns <- crop(folk, depth_base_0)
folk_sbns
plot(folk_sbns)
```
grid cell of polygons with IDs
```{r}
template  <- depth_base_0          # base grid
sea_mask  <- template > 0          # TRUE on active sea cells
n_cells   <- ncell(template)
n_cells
# Create one polygon per raster cell (in current CRS)
grid <- as.polygons(template, values = FALSE, dissolve = FALSE)
grid$cell_id <- 1:nrow(grid)      # 1 : n_cells
nrow(grid)

path_folk <- file.path(
  base_dir,
  "SedimentHabitats",
  "seabed_substrate_250k.shp"
)

folk <- vect(path_folk)

# Reproject Folk to base grid CRS if needed
if (crs(folk) != crs(template)) {
  folk <- project(folk, template)
}

# Clip Folk to SBNS/base extent
folk_sbns <- crop(folk, template)

library(terra)
library(dplyr)

folk_field <- "folk_5cl"

# Intersections: each row is (cell polygon ∩ sediment polygon)
iv <- intersect(grid, folk_sbns)
iv
nrow(iv)       # this SHOULD now be >> 60, likely several thousand

# Compute overlap area per intersection piece in m² (geodesic)
iv$area_m2 <- expanse(iv, unit = "m")

# Extract attributes to a data.frame
iv_df <- as.data.frame(iv[, c("cell_id", folk_field, "area_m2")])
names(iv_df)[2] <- "folk_5cl"
head(iv_df)

dom_tbl <- iv_df %>%
  group_by(cell_id, folk_5cl) %>%
  summarise(area_m2 = sum(area_m2), .groups = "drop") %>%
  group_by(cell_id) %>%
  slice_max(area_m2, n = 1, with_ties = FALSE) %>%
  ungroup()

dom_tbl
nrow(dom_tbl)  # number of cells that intersect at least one sediment polygon
folk_vals <- c(
  "1. Mud to muddy Sand",
  "2. Sand",
  "3. Coarse-grained sediment",
  "4. Mixed sediment",
  "5. Rock & boulders"
)

# Map label -> numeric ID 1..5
class_to_id <- setNames(seq_along(folk_vals), folk_vals)

# Vector of habitat IDs per cell (NA initial)
hab_id_vec <- rep(NA_integer_, n_cells)

# Add numeric IDs to dom_tbl
dom_tbl$hab_id <- class_to_id[dom_tbl$folk_5cl]

# Assign IDs by cell_id
hab_id_vec[dom_tbl$cell_id] <- dom_tbl$hab_id

# Convert to raster aligned with template
hab_id <- template
values(hab_id) <- hab_id_vec

# Keep only sea cells
hab_id[!sea_mask] <- NA

names(hab_id) <- "Folk5_id_dominant"

plot(hab_id, main = "Dominant Folk-5 ID (1–5) per sea cell")

```

Fill cells without habitat (nearest habitat)
```{r}
# Sea cells with no class assigned yet
na_sea <- is.na(hab_id) & sea_mask
sum(values(na_sea), na.rm = TRUE)  # how many sea cells without habitat?

# Distance to each class
dist_layers <- list()

for (i in seq_along(folk_vals)) {
  cat("Computing distance to class ID", i, ":", folk_vals[i], "\n")
  
  is_class <- hab_id == i
  dist_i   <- distance(is_class)    # distance to nearest cell of this class
  
  dist_layers[[i]] <- dist_i
}

dist_stack <- rast(dist_layers)
names(dist_stack) <- paste0("dist_to_id_", seq_along(folk_vals))

# For each cell, ID of the closest class
nearest_id <- which.min(dist_stack)

# Fill NA sea cells with nearest class
hab_filled <- hab_id
hab_filled[na_sea] <- nearest_id[na_sea]

# Land/inactive remain NA
hab_filled[!sea_mask] <- NA

names(hab_filled) <- "Folk5_id_final"

plot(hab_filled, main = "Final Folk-5 ID (dominant + nearest fill)")

```
Build 5 binary habitat rasters
```{r}
folk_labels <- c(
  "1. Mud to muddy Sand"       = "Mud_MuddySand",
  "2. Sand"                    = "Sand",
  "3. Coarse-grained sediment" = "CoarseSed",
  "4. Mixed sediment"          = "MixedSed",
  "5. Rock & boulders"         = "RockBoulders"
)

hab_rasters <- list()

for (i in seq_along(folk_vals)) {
  cl    <- folk_vals[i]
  short <- folk_labels[[cl]]
  
  r_cl <- template
  values(r_cl) <- 0
  
  # 1 where this class is the final ID AND sea
  r_cl[sea_mask & (hab_filled == i)] <- 1
  
  names(r_cl) <- paste0("Folk5_", short)
  hab_rasters[[short]] <- r_cl
}

# check if they sum to 1
hab_stack <- rast(hab_rasters)
hab_sum   <- sum(hab_stack)
names(hab_sum) <- "Folk5_sum"

plot(hab_sum, main = "Sum of Folk-5 binary layers")

sum_values <- values(hab_sum)
sea_idx    <- which(values(sea_mask))

table(sum_values[sea_idx], useNA = "ifany")

# Quick plot to check rasters
par(mfrow = c(2, 3))

for (nm in names(hab_rasters)) {
  plot(hab_rasters[[nm]], 
       main = nm, 
       col = c("white", "dodgerblue"))
}
```

Export to 5 habitat rasters
```{r}
for (short in names(hab_rasters)) {
  r_cl <- hab_rasters[[short]]
  fname_stub <- paste0("Folk5_", short)
  
  writeRaster(
    r_cl,
    filename  = file.path(out_dir, paste0(fname_stub, ".tif")),
    overwrite = TRUE
  )
  
  writeRaster(
    r_cl,
    filename  = file.path(out_dir, paste0(fname_stub, ".asc")),
    overwrite = TRUE
  )
}
```
--------------------------------------------------------------------------
Spatial Temporal Data (Habitat + OWFs)
Need maps from 1991 to 2150
first make OWF presence or not
Then make 2 times 5 habitat layers:
1) Habitat-name-(without)-OWF indicating for each habitat its presence (will be 1 where habitat is present, 0 where absent, between 0 and 1 where overlapping with OWF area)
2) For each habitat also make Habitat_OWF (%OWF in cell in that habitat) Will be 0 where OWFs are not present, between 0 and 1 where present in that habitat

If data available for owf:
area
N turbines
--> calculate substrate proportion to cell

If no data available for owf
--> use avergage  turbine density (calculated from owfs for which data is available)
--> estimate substrate proportion to cell

Parameter initialisatoin
```{r}
# --- Parameters from ter Hofstede et al. (2023) + Püts-style 1 m lateral
# Scour / armour rock area per turbine [m²]
A_scour <- 530.2      
# Representative average monopile diameter [m]
D_mp <- 7             
# Lateral surface of 1 m monopile [m²]
A_lat_1m <- pi * D_mp   # ≈ 21.99
# Total hard-substrate area per turbine [m²]
A_hard_turb <- A_scour + A_lat_1m  # ≈ 552.2 m²
A_hard_turb
```
Farm-specific owf density (rho)
```{r}
library(terra)
library(dplyr)

# Base grid
template <- depth_base_0
sea_mask <- template > 0

# Path to OWF polygons
path_owf <- file.path(
  base_dir,
  "OffshoreWindFarms",
  "EMODnet_HA_Energy_WindFarms_pg_20250825.shp"
)

# Load OWF polygons
owf <- vect(path_owf)

# Project to base grid CRS
if (crs(owf) != crs(template)) {
  owf <- project(owf, template)
}

# Restrict to SBNS / model domain
owf_sbns <- crop(owf, template)

# Inspect fields once (just to see N_TURBINES etc.)
names(owf_sbns)

# Ensure turbine count column is numeric; adjust "N_TURBINES" if your field is named differently
owf_sbns$N_TURBINES <- as.numeric(owf_sbns$N_TURBINES)

# Compute geometric farm area [m²] for each OWF polygon
owf_sbns$A_farm_m2 <- expanse(owf_sbns, unit = "m")

# Compute global mean turbindensity
valid <- !is.na(owf_sbns$N_TURBINES) & owf_sbns$N_TURBINES > 0 &
         !is.na(owf_sbns$A_farm_m2)  & owf_sbns$A_farm_m2  > 0

sum(valid)  # how many farms have proper turbine counts?

# Total turbines and total area for valid polygons
tot_turbines <- sum(owf_sbns$N_TURBINES[valid])
tot_area_m2  <- sum(owf_sbns$A_farm_m2[valid])

# Global mean turbine density from *your* EMODnet data [turbines per m²]
dens_mean_per_m2 <- tot_turbines / tot_area_m2

# In more intuitive units: turbines per km²
dens_mean_per_km2 <- dens_mean_per_m2 * 1e6

dens_mean_per_m2
dens_mean_per_km2

```
calculate farm-specific hard substrate density
```{r}
# Initialise rho with NA
rho <- rep(NA_real_, nrow(owf_sbns))

# 1) Farms with explicit N_TURBINES: farm-specific rho
rho[valid] <- (owf_sbns$N_TURBINES[valid] * A_hard_turb) /
              owf_sbns$A_farm_m2[valid]

# 2) Farms without N_TURBINES: fallback using EMODnet-derived mean density
fallback_idx <- !valid & !is.na(owf_sbns$A_farm_m2) & owf_sbns$A_farm_m2 > 0

rho_fallback <- dens_mean_per_m2 * A_hard_turb  # m²/m² (dimensionless)
rho[fallback_idx] <- rho_fallback

# Just in case we still have NA (strange geometries), set to 0
rho[is.na(rho)] <- 0

owf_sbns$rho <- rho

summary(owf_sbns$rho)

```

Cell-fraction that is OWF hard substrate:
```{r}
# Build grid polygons from template
grid <- as.polygons(template, values = FALSE, dissolve = FALSE)
grid$cell_id <- 1:nrow(grid)

# Cell area [m²]
A_cell <- expanse(grid[1, ], unit = "m")

# Intersect grid with OWFs
iv <- intersect(grid, owf_sbns)

iv$A_int_m2 <- expanse(iv, unit = "m")

iv_df <- as.data.frame(iv[, c("cell_id", "rho", "A_int_m2")])

hard_by_cell <- iv_df %>%
  mutate(f_piece = (A_int_m2 / A_cell) * rho) %>%
  group_by(cell_id) %>%
  summarise(f_hard = sum(f_piece), .groups = "drop")

hard_by_cell$f_hard <- pmin(hard_by_cell$f_hard, 1)

f_hard_vec <- rep(0, ncell(template))
f_hard_vec[hard_by_cell$cell_id] <- hard_by_cell$f_hard

f_hard_r <- template
values(f_hard_r) <- f_hard_vec
names(f_hard_r) <- "OWF_hard_frac"

# land/inactive as 0
f_hard_r[!sea_mask] <- 0

plot(f_hard_r, main = "Fraction of cell that is OWF hard substrate")
global(f_hard_r, fun = range, na.rm = TRUE)

```

OWF presence raster (MPA layer in Ecospace)
```{r}
owf_presence <- rasterize(
  owf_sbns,
  template,
  field      = 1,
  background = 0,
  touches    = TRUE
)
owf_presence[!sea_mask] <- 0
names(owf_presence) <- "OWF_presence"

plot(owf_presence)


# Store to .asc file for Ecospace input of MPA
library(terra)

years_all <- 1991:2150

# Output dir for MPA rasters
out_mpa_dir <- file.path(base_dir, "Processed", "OWF_MPA")
dir.create(out_mpa_dir, recursive = TRUE, showWarnings = FALSE)

for (y in years_all) {
  if (y < 2025) {
    # No MPA yet
    r <- template
    values(r) <- 0
  } else {
    # MPA = OWF presence
    r <- owf_presence
  }
  
  # Ensure land/inactive are 0
  r[!sea_mask] <- 0
  names(r) <- "OWF_MPA"
  
  # Save as ASCII for Ecospace
  writeRaster(
    r,
    filename  = file.path(out_mpa_dir, paste0("OWF_MPA_", y, ".asc")),
    overwrite = TRUE
  )
}
```

Habitat_noOWF and Habitat_OWF per folk5 class
```{r}
out_hab_dir <- file.path(base_dir, "Processed", "Habitats_OWF")
dir.create(out_hab_dir, recursive = TRUE, showWarnings = FALSE)

hab_owf_rasters <- list()

for (nm in names(hab_rasters)) {
  H <- hab_rasters[[nm]]   # 0/1 raster for this Folk5 class
  
  # Remaining soft habitat
  H_noOWF <- H * (1 - f_hard_r)
  names(H_noOWF) <- paste0("Folk5_", nm, "_noOWF")
  
  # OWF hard substrate
  H_OWF <- H * f_hard_r
  names(H_OWF) <- paste0("Folk5_", nm, "_OWF")
  
  hab_owf_rasters[[paste0(nm, "_noOWF")]] <- H_noOWF
  hab_owf_rasters[[paste0(nm, "_OWF")]]   <- H_OWF
  
  writeRaster(
    H_noOWF,
    filename  = file.path(out_hab_dir, paste0("Folk5_", nm, "_noOWF.asc")),
    overwrite = TRUE
  )
  writeRaster(
    H_OWF,
    filename  = file.path(out_hab_dir, paste0("Folk5_", nm, "_OWF.asc")),
    overwrite = TRUE
  )
}

```
Store Habitat data 1991-2150
```{r}
years_all <- 1991:2150

# Output directory for temporal habitat layers
out_hab_time_dir <- file.path(base_dir, "Processed", "Habitats_OWF_temporal")
dir.create(out_hab_time_dir, recursive = TRUE, showWarnings = FALSE)

for (nm in names(hab_rasters)) {
  # Base (clean) habitat raster (0/1)
  H_base <- hab_rasters[[nm]]
  
  # OWF-adjusted rasters
  H_noOWF_OWF <- hab_owf_rasters[[paste0(nm, "_noOWF")]]  # Folk5_<nm>_noOWF
  H_OWF       <- hab_owf_rasters[[paste0(nm, "_OWF")]]    # Folk5_<nm>_OWF
  
  # Sanity check: alignment
  stopifnot(
    all.equal(ext(H_base), ext(template)),
    nrow(H_base) == nrow(template),
    ncol(H_base) == ncol(template)
  )
  
  for (y in years_all) {
    if (y < 2025) {
      # No OWFs yet: only base habitat, no OWF hard substrate
      r_noOWF <- H_base
      r_OWF   <- template
      values(r_OWF) <- 0
    } else {
      # OWFs active: use OWF-adjusted layers
      r_noOWF <- H_noOWF_OWF
      r_OWF   <- H_OWF
    }
    
    # Enforce 0 on land/inactive (Ecospace convention)
    r_noOWF[!sea_mask] <- 0
    r_OWF[!sea_mask]   <- 0
    
    names(r_noOWF) <- paste0("Folk5_", nm, "_noOWF")
    names(r_OWF)   <- paste0("Folk5_", nm, "_OWF")
    
    # File names: include habitat + year
    fn_noOWF <- file.path(
      out_hab_time_dir,
      paste0("Folk5_", nm, "_noOWF_", y, ".asc")
    )
    fn_OWF <- file.path(
      out_hab_time_dir,
      paste0("Folk5_", nm, "_OWF_", y, ".asc")
    )
    
    writeRaster(r_noOWF, filename = fn_noOWF, overwrite = TRUE)
    writeRaster(r_OWF,   filename = fn_OWF,   overwrite = TRUE)
  }
}

```

```{r}
hab <- "Sand"                    # change to: Mud, Sand, Mixed, Coarse, Rock
years_to_plot <- c(1999, 2030)

rasters_to_plot <- list()

for (y in years_to_plot) {
  f_no   <- file.path(out_hab_time_dir, paste0("Folk5_", hab, "_noOWF_", y, ".asc"))
  f_owf  <- file.path(out_hab_time_dir, paste0("Folk5_", hab, "_OWF_", y, ".asc"))
  
  r_no  <- rast(f_no)
  r_owf <- rast(f_owf)
  
  rasters_to_plot[[paste0("noOWF_", y)]] <- r_no
  rasters_to_plot[[paste0("OWF_",  y)]] <- r_owf
}

# Plot
par(mfrow = c(2, length(years_to_plot)))

for (nm in names(rasters_to_plot)) {
  plot(rasters_to_plot[[nm]], main = nm)
}

```


Mean turbine density from empdnet data: 1 turbine per km²
```{r}
# Offshore wind farm polygons (Operational and planned)
("C:\Users\dlorr\OneDrive - UGent\Documents\PhD\EwE_OWF\Spatial data\OffshoreWindFarms\EMODnet_HA_Energy_WindFarms_pg_20250825.shp")
# Mean hard-substrate density per OWF area (dimensionless)

rho_mean <- A_hard_turb / (dens_km2 * 1e6)  # m² / m²


A_hard_turb
rho_mean
```



Spatial Temporal Data (Temperature and Salinity)
```{r}
library(terra)
library(dplyr)

# Base data folders
base_dir   <- "C:/Users/dlorr/OneDrive - UGent/Documents/PhD/EwE_OWF/Spatial data"
path_temp  <- file.path(base_dir, "Temperature")
path_sal   <- file.path(base_dir, "Salinity")

# Root output folder
out_root <- file.path(base_dir, "Processed")
if (!dir.exists(out_root)) dir.create(out_root)

# Variable- and scenario-specific output dirs
out_temp_baseline <- file.path(out_root, "Temperature", "Baseline")
out_sal_baseline  <- file.path(out_root, "Salinity", "Baseline")

dir.create(out_temp_baseline, recursive = TRUE, showWarnings = FALSE)
dir.create(out_sal_baseline,  recursive = TRUE, showWarnings = FALSE)

# Base grid
template <- depth_base_0
sea_mask <- template > 0
```
Helper function to fetch year and month
```{r}
get_year_month_cmems <- function(f) {
  b <- basename(f)
  m <- regexpr("[0-9]{8}", b)
  date_str <- substr(b, m, m + 7)
  year  <- as.integer(substr(date_str, 1, 4))
  month <- as.integer(substr(date_str, 5, 6))
  list(year = year, month = month)
}
```

Helper function to extract means of 1993-2024 data (available data)
```{r}
compute_annual_means <- function(path_nc, subds_name, template, sea_mask) {
  files <- list.files(path_nc, pattern = "\\.nc$", full.names = TRUE)
  if (length(files) == 0) stop("No NetCDF files found in ", path_nc)
  
  # Parse year/month
  ym <- lapply(files, get_year_month_cmems)
  years  <- sapply(ym, `[[`, "year")
  months <- sapply(ym, `[[`, "month")
  
  df <- data.frame(file = files, year = years, month = months)
  
  # Keep years up to and including 2024
  df <- df[df$year <= 2024, ]
  if (nrow(df) == 0) stop("No files with year <= 2024 in ", path_nc)
  
  annual_list <- list()
  yrs <- sort(unique(df$year))
  
  for (y in yrs) {
    cat("Processing year:", y, "\n")
    df_y <- df[df$year == y, ]
    
    # Init accumulators on template
    sum_r   <- template; values(sum_r)   <- 0
    count_r <- template; values(count_r) <- 0
    
    for (f in df_y$file) {
      cat("  -", basename(f), "\n")
      
      # Read the variable
      r_raw <- try({
        rast(f, subds = subds_name)
      }, silent = TRUE)
      
      if (inherits(r_raw, "try-error")) {
        # Fallback: take first layer if subdataset name fails
        r_raw <- rast(f)
      }
      
      # ⚠️ Ensure we only work with ONE layer (e.g. surface / first level)
      if (nlyr(r_raw) > 1) {
        r_raw <- r_raw[[1]]
      }
      
      # Project onto base grid (still single-layer)
      r_proj <- project(r_raw, template, method = "bilinear")
      
      # Just in case project() returns multiple layers (shouldn't, but be safe)
      if (nlyr(r_proj) > 1) {
        r_proj <- r_proj[[1]]
      }
      
      # Mask to sea cells only
      r_proj[!sea_mask] <- NA
      
      # NA handling
      non_na <- !is.na(r_proj)
      r_proj_na0 <- r_proj
      r_proj_na0[is.na(r_proj_na0)] <- 0
      
      # Accumulate
      sum_r   <- sum_r   + r_proj_na0
      count_r <- count_r + non_na
    }
    
    mean_r <- sum_r / count_r
    mean_r[count_r == 0] <- NA
    
    # Now mean_r is guaranteed to be single-layer
    stopifnot(nlyr(mean_r) == 1)
    
    names(mean_r) <- paste0("year_", y)
    annual_list[[as.character(y)]] <- mean_r
  }
  
  annual_list
}
```

For monthly data:
```{r}
compute_monthly_rasters <- function(path_nc, subds_name, template, sea_mask) {
  files <- list.files(path_nc, pattern = "\\.nc$", full.names = TRUE)
  if (length(files) == 0) stop("No NetCDF files found in ", path_nc)
  
  # Parse year/month
  ym <- lapply(files, get_year_month_cmems)
  years  <- sapply(ym, `[[`, "year")
  months <- sapply(ym, `[[`, "month")
  
  df <- data.frame(file = files, year = years, month = months)
  
  # Only up to and including 2024
  df <- df[df$year <= 2024, ]
  if (nrow(df) == 0) stop("No files with year <= 2024 in ", path_nc)
  
  # Sort by time (year, month)
  df <- df[order(df$year, df$month), ]
  
  monthly_list <- vector("list", nrow(df))
  names(monthly_list) <- character(nrow(df))
  
  for (i in seq_len(nrow(df))) {
    f <- df$file[i]
    y <- df$year[i]
    m <- df$month[i]
    
    cat("Monthly raster:", y, sprintf("%02d", m), "-", basename(f), "\n")
    
    r_raw <- try({
      rast(f, subds = subds_name)
    }, silent = TRUE)
    
    if (inherits(r_raw, "try-error")) {
      r_raw <- rast(f)
    }
    if (nlyr(r_raw) > 1) {
      r_raw <- r_raw[[1]]
    }
    
    r_proj <- project(r_raw, template, method = "bilinear")
    if (nlyr(r_proj) > 1) {
      r_proj <- r_proj[[1]]
    }
    
    # Mask land
    r_proj[!sea_mask] <- NA
    
    nm <- sprintf("Y%04d_M%02d", y, m)
    names(r_proj) <- nm
    monthly_list[[i]] <- r_proj
    names(monthly_list)[i] <- nm
  }
  
  monthly_stack <- rast(monthly_list)
  monthly_stack
}
```


Check variable names
```{r}
# Temperature
ex_temp <- list.files(path_temp, pattern = "\\.nc$", full.names = TRUE)[1]
sds(rast(ex_temp))

# Salinity
ex_sal <- list.files(path_sal, pattern = "\\.nc$", full.names = TRUE)[1]
sds(rast(ex_sal))
```

Annual means for SST and SSS
```{r}
# Adjust these after checking sds()
var_temp <- "thetao"   # or the actual temp variable name
var_sal  <- "so"       # or the actual salinity variable name

# Temperature annual means
temp_annual <- compute_annual_means(
  path_nc    = path_temp,
  subds_name = var_temp,
  template   = template,
  sea_mask   = sea_mask
)

# Salinity annual means
sal_annual <- compute_annual_means(
  path_nc    = path_sal,
  subds_name = var_sal,
  template   = template,
  sea_mask   = sea_mask
)

# Quick checks
length(temp_annual); names(temp_annual)
length(sal_annual);  names(sal_annual)

nlyr(temp_annual[[1]])         # should be 1
plot(temp_annual[[1]])
plot(temp_annual[[15]])
plot(temp_annual[[32]])
nlyr(sal_annual[[1]])         # should be 1
plot(sal_annual[[1]])
plot(sal_annual[[15]])
plot(sal_annual[[32]])

plot(temp_annual[["1993"]], main = "Temp annual mean 1993")
plot(sal_annual[["1993"]],  main = "Sal annual mean 1993")
```
For monthly data
```{r}
# Monthly rasters 1993–2024 (baseline dataset)
temp_monthly_raw <- compute_monthly_rasters(
  path_nc    = path_temp,
  subds_name = var_temp,
  template   = template,
  sea_mask   = sea_mask
)

plot(temp_monthly_raw[[1]])
plot(temp_monthly_raw[[5]])
plot(temp_monthly_raw[[12]])

sal_monthly_raw <- compute_monthly_rasters(
  path_nc    = path_sal,
  subds_name = var_sal,
  template   = template,
  sea_mask   = sea_mask
)

# temp_monthly_raw
# sal_monthly_raw
```

Generate Baseline scenario [YEARLY]:
Baseline rule (per variable):
1991–1992 → use 1993
1993–2024 → use that year’s own raster
2025–2150 → use 2024
```{r}
years_all <- 1991:2150

## ---- Temperature baseline stack ----
temp_baseline_list <- vector("list", length(years_all))
names(temp_baseline_list) <- years_all

for (y in years_all) {
  if (y <= 1992) {
    src_year <- 1993
  } else if (y <= 2024) {
    src_year <- y
  } else {
    src_year <- 2024
  }
  
  src_key <- as.character(src_year)
  temp_baseline_list[[as.character(y)]] <- temp_annual[[src_key]]
}

temp_baseline_stack <- rast(temp_baseline_list)
names(temp_baseline_stack) <- paste0("Temp_", years_all)
temp_baseline_stack


## ---- Salinity baseline stack ----
sal_baseline_list <- vector("list", length(years_all))
names(sal_baseline_list) <- years_all

for (y in years_all) {
  if (y <= 1992) {
    src_year <- 1993
  } else if (y <= 2024) {
    src_year <- y
  } else {
    src_year <- 2024
  }
  
  src_key <- as.character(src_year)
  sal_baseline_list[[as.character(y)]] <- sal_annual[[src_key]]
}

sal_baseline_stack <- rast(sal_baseline_list)
names(sal_baseline_stack) <- paste0("Sal_", years_all)
sal_baseline_stack


# saveRDS(temp_baseline_stack,
#         file = file.path(out_temp_baseline, "Temp_Baseline_1991_2150.rds"))
# 
# saveRDS(sal_baseline_stack,
#         file = file.path(out_sal_baseline, "Sal_Baseline_1991_2150.rds"))
```

Save .asc files per variable [YEARLY]
```{r}
# Temperature
for (i in seq_along(years_all)) {
  y <- years_all[i]
  r <- temp_baseline_stack[[i]]
  
  r0 <- r
  r0[!sea_mask] <- 0   # 0 on land / inactive cells
  
  writeRaster(
    r0,
    filename  = file.path(out_temp_baseline, paste0("Temp_Baseline_", y, ".asc")),
    overwrite = TRUE
  )
}

# Salinity
for (i in seq_along(years_all)) {
  y <- years_all[i]
  r <- sal_baseline_stack[[i]]
  
  r0 <- r
  r0[!sea_mask] <- 0   # 0 on land / inactive cells
  
  writeRaster(
    r0,
    filename  = file.path(out_sal_baseline, paste0("Sal_Baseline_", y, ".asc")),
    overwrite = TRUE
  )
}

```

Generate Baseline scenario [MONTLY]:
Baseline rule (per variable):
1991–1992 → use 1993
1993–2024 → use that year’s own raster
2025–2150 → use 2024
```{r}
years_all  <- 1991:2150
months_all <- 1:12

## ---- Temperature baseline monthly stack ----
temp_baseline_monthly_list <- list()

for (y in years_all) {
  if (y <= 1992) {
    src_year <- 1993
  } else if (y <= 2024) {
    src_year <- y
  } else {
    src_year <- 2024
  }
  
  for (m in months_all) {
    src_name <- sprintf("Y%04d_M%02d", src_year, m)
    if (!src_name %in% names(temp_monthly_raw)) {
      stop("Missing monthly source:", src_name)
    }
    
    r <- temp_monthly_raw[[src_name]]
    out_name <- sprintf("Temp_%d_%02d", y, m)
    
    temp_baseline_monthly_list[[out_name]] <- r
  }
}

temp_baseline_monthly_stack <- rast(temp_baseline_monthly_list)
temp_baseline_monthly_stack
# layer names already like "Temp_1991_01", "Temp_1991_02", ...

## ---- Salinity baseline monthly stack ----
sal_baseline_monthly_list <- list()

for (y in years_all) {
  if (y <= 1992) {
    src_year <- 1993
  } else if (y <= 2024) {
    src_year <- y
  } else {
    src_year <- 2024
  }
  
  for (m in months_all) {
    src_name <- sprintf("Y%04d_M%02d", src_year, m)
    if (!src_name %in% names(sal_monthly_raw)) {
      stop("Missing monthly source:", src_name)
    }
    
    r <- sal_monthly_raw[[src_name]]
    out_name <- sprintf("Sal_%d_%02d", y, m)
    
    sal_baseline_monthly_list[[out_name]] <- r
  }
}

sal_baseline_monthly_stack <- rast(sal_baseline_monthly_list)
sal_baseline_monthly_stack
```
Save .asc files per variable [MONTLY]
```{r}
# Temperature monthly baseline ASCII
for (y in years_all) {
  for (m in months_all) {
    lyr_name <- sprintf("Temp_%d_%02d", y, m)
    r <- temp_baseline_monthly_stack[[lyr_name]]
    
    r0 <- r
    r0[!sea_mask] <- 0  # 0 on land / inactive
    
    fn <- file.path(
      out_temp_baseline,
      sprintf("Temp_Baseline_%d_%02d.asc", y, m)
    )
    
    writeRaster(r0, filename = fn, overwrite = TRUE)
  }
}

# Salinity monthly baseline ASCII
for (y in years_all) {
  for (m in months_all) {
    lyr_name <- sprintf("Sal_%d_%02d", y, m)
    r <- sal_baseline_monthly_stack[[lyr_name]]
    
    r0 <- r
    r0[!sea_mask] <- 0
    
    fn <- file.path(
      out_sal_baseline,
      sprintf("Sal_Baseline_%d_%02d.asc", y, m)
    )
    
    writeRaster(r0, filename = fn, overwrite = TRUE)
  }
}
```


Generate Climate change scenario
```{r}
# Root processed folder from before
out_root <- file.path(base_dir, "Processed")

# Climate change scenario (SSP3-7.0) output dirs
out_temp_cc <- file.path(out_root, "Temperature", "ClimateChange_SSP3-7.0")
out_sal_cc  <- file.path(out_root, "Salinity",   "ClimateChange_SSP3-7.0")

dir.create(out_temp_cc, recursive = TRUE, showWarnings = FALSE)
dir.create(out_sal_cc,  recursive = TRUE, showWarnings = FALSE)


path_sst_csv <- "C:/Users/dlorr/OneDrive - UGent/Documents/PhD/EwE_OWF/R-scripts/Climate_projections_SSP3-7.0/SST_SSP3-7.0_monthly_1991_2099.csv"

path_sss_csv <- "C:/Users/dlorr/OneDrive - UGent/Documents/PhD/EwE_OWF/R-scripts/Climate_projections_SSP3-7.0/SSS_SSP3-7.0_monthly_1991_2099.csv"

```

SST and SSS annual means
```{r}
library(dplyr)
library(readr)

# Helper: compute annual means from monthly CSV
compute_annual_from_csv <- function(path_csv, value_col = "value") {
  df <- read_csv(path_csv, show_col_types = FALSE)
  # Ensure column names exist: adjust if your file uses other names
  
  df_annual <- df %>%
    group_by(year) %>%
    summarise(
      val = mean(.data[[value_col]], na.rm = TRUE),
      .groups = "drop"
    )
  
  df_annual
}

# Temperature (SST) annual means 1991–2099
sst_annual <- compute_annual_from_csv(path_sst_csv, value_col = "value")
write.csv(sst_annual, "SST_annual_1991-2100.csv")
# Salinity (SSS) annual means 1991–2099
sss_annual <- compute_annual_from_csv(path_sss_csv, value_col = "value")
write.csv(sss_annual, "SSS_annual_1991-2100.csv")
```

Calculate anomalies from 2024 onwards
```{r}
ref_year <- 2024

# Temperature anomalies
sst_ref <- sst_annual$val[sst_annual$year == ref_year]
sst_annual <- sst_annual %>%
  mutate(anom_ref24 = val - sst_ref)

# Salinity anomalies
sss_ref <- sss_annual$val[sss_annual$year == ref_year]
sss_annual <- sss_annual %>%
  mutate(anom_ref24 = val - sss_ref)

# Make named vectors: year -> anomaly
anom_sst <- setNames(sst_annual$anom_ref24, sst_annual$year)
anom_sss <- setNames(sss_annual$anom_ref24, sss_annual$year)
```

Build cull CC rasters (1991-2024 + 2025-2150)

1991–2024: CC scenario = baseline (no anomaly)
2025–2099: CC scenario = 2024 baseline grid + annual anomaly(y) 
           (uniform across SBNS)
2100–2150: CC scenario = 2024 baseline grid + anomaly(2099) (held constant)
```{r}
years_all <- 1991:2150

# Helper to extract index of a given year layer in baseline stacks
idx_year <- function(stack, prefix, year) {
  which(names(stack) == paste0(prefix, "_", year))
}

# 2024 base fields
idx_T2024 <- idx_year(temp_baseline_stack, "Temp", 2024)
idx_S2024 <- idx_year(sal_baseline_stack,  "Sal",  2024)

r_T2024 <- temp_baseline_stack[[idx_T2024]]
r_S2024 <- sal_baseline_stack[[idx_S2024]]

## ---- Temperature CC stack ----
temp_cc_list <- vector("list", length(years_all))
names(temp_cc_list) <- years_all

for (i in seq_along(years_all)) {
  y <- years_all[i]
  
  if (y <= 2024) {
    # Historical period: identical to baseline
    r_cc <- temp_baseline_stack[[i]]
    
  } else {
    # Future period: 2024 base field + anomaly
    # cap anomaly years to 2099 (after that: hold 2099 constant)
    if (y <= 2099) {
      y_clim <- y
    } else {
      y_clim <- 2099
    }
    
    # anomaly (may be NA if outside 1991-2099)
    a <- anom_sst[as.character(y_clim)]
    
    # start from 2024 baseline field
    r_cc <- r_T2024
    
    # add anomaly only on sea cells
    r_cc[sea_mask] <- r_cc[sea_mask] + a
  }
  
  temp_cc_list[[as.character(y)]] <- r_cc
}

temp_cc_stack <- rast(temp_cc_list)
names(temp_cc_stack) <- paste0("Temp_CC_", years_all)
temp_cc_stack

## ---- Salinity CC stack ----
sal_cc_list <- vector("list", length(years_all))
names(sal_cc_list) <- years_all

for (i in seq_along(years_all)) {
  y <- years_all[i]
  
  if (y <= 2024) {
    # Historical: identical to baseline
    r_cc <- sal_baseline_stack[[i]]
    
  } else {
    if (y <= 2099) {
      y_clim <- y
    } else {
      y_clim <- 2099
    }
    
    a <- anom_sss[as.character(y_clim)]
    
    r_cc <- r_S2024
    r_cc[sea_mask] <- r_cc[sea_mask] + a
  }
  
  sal_cc_list[[as.character(y)]] <- r_cc
}

sal_cc_stack <- rast(sal_cc_list)
names(sal_cc_stack) <- paste0("Sal_CC_", years_all)
sal_cc_stack


# Check
plot(temp_cc_stack[[which(names(temp_cc_stack) %in% c("Temp_CC_2024","Temp_CC_2050","Temp_CC_2099","Temp_CC_2150"))]])
plot(sal_cc_stack[[which(names(sal_cc_stack)  %in% c("Sal_CC_2024","Sal_CC_2050","Sal_CC_2099","Sal_CC_2150"))]])

```

Save data
```{r}
saveRDS(temp_cc_stack,
        file = file.path(out_temp_cc, "Temp_CC_SSP3-7.0_1991_2150.rds"))

saveRDS(sal_cc_stack,
        file = file.path(out_sal_cc, "Sal_CC_SSP3-7.0_1991_2150.rds"))

# Temperature .asc 
for (i in seq_along(years_all)) {
  y <- years_all[i]
  r <- temp_cc_stack[[i]]
  
  r0 <- r
  r0[!sea_mask] <- 0  # 0 on land/inactive
  
  writeRaster(
    r0,
    filename  = file.path(out_temp_cc, paste0("Temp_CC_SSP3-7.0_", y, ".asc")),
    overwrite = TRUE
  )
}

# Salinity .asc
for (i in seq_along(years_all)) {
  y <- years_all[i]
  r <- sal_cc_stack[[i]]
  
  r0 <- r
  r0[!sea_mask] <- 0
  
  writeRaster(
    r0,
    filename  = file.path(out_sal_cc, paste0("Sal_CC_SSP3-7.0_", y, ".asc")),
    overwrite = TRUE
  )
}

```

Helper function for Montly data:
```{r}
compute_monthly_anoms_from_csv <- function(path_csv, ref_year = 2024, value_col = "value") {
  df <- readr::read_csv(path_csv, show_col_types = FALSE)
  
  # Expect columns 'year' and 'month' (1–12)
  stopifnot(all(c("year", "month") %in% names(df)))
  
  # Select only what's needed
  df <- df %>% dplyr::select(year, month, !!value_col)
  
  # Reference: year == ref_year
  df_ref <- df %>%
    dplyr::filter(year == ref_year) %>%
    dplyr::select(month, ref_val = !!value_col)
  
  if (nrow(df_ref) != 12) {
    warning("Ref year does not have exactly 12 months in ", path_csv)
  }
  
  df_anom <- df %>%
    dplyr::left_join(df_ref, by = "month") %>%
    dplyr::mutate(anom_ref = .data[[value_col]] - .data$ref_val)
  
  # Vector named "YYYY_MM"
  key <- paste0(df_anom$year, "_", sprintf("%02d", df_anom$month))
  anom_vec <- setNames(df_anom$anom_ref, key)
  anom_vec
}
```

Compute montly layers CC
```{r}
anom_sst_month <- compute_monthly_anoms_from_csv(
  path_csv  = path_sst_csv,
  ref_year  = 2024,
  value_col = "value"
)

anom_sss_month <- compute_monthly_anoms_from_csv(
  path_csv  = path_sss_csv,
  ref_year  = 2024,
  value_col = "value"
)
```

Build CC raster stack
```{r}
years_all  <- 1991:2150
months_all <- 1:12

## Convenience: baseline 2024 monthly fields
get_temp_2024_month <- function(m) {
  nm <- sprintf("Temp_2024_%02d", m)
  temp_baseline_monthly_stack[[nm]]
}
get_sal_2024_month <- function(m) {
  nm <- sprintf("Sal_2024_%02d", m)
  sal_baseline_monthly_stack[[nm]]
}

## ---- Temperature CC monthly stack ----
temp_cc_monthly_list <- list()

for (y in years_all) {
  for (m in months_all) {
    
    if (y <= 2024) {
      # Historical: identical to baseline
      nm_base <- sprintf("Temp_%d_%02d", y, m)
      r_cc <- temp_baseline_monthly_stack[[nm_base]]
      
    } else {
      # Future: baseline 2024 month + anomaly
      if (y <= 2099) {
        y_clim <- y
      } else {
        y_clim <- 2099
      }
      
      key <- paste0(y_clim, "_", sprintf("%02d", m))
      a <- anom_sst_month[[key]]
      if (is.null(a) || is.na(a)) {
        stop("Missing SST monthly anomaly for ", key)
      }
      
      r_cc <- get_temp_2024_month(m)
      r_cc[sea_mask] <- r_cc[sea_mask] + a
    }
    
    out_name <- sprintf("Temp_CC_%d_%02d", y, m)
    temp_cc_monthly_list[[out_name]] <- r_cc
  }
}

temp_cc_monthly_stack <- rast(temp_cc_monthly_list)
temp_cc_monthly_stack


## ---- Salinity CC monthly stack ----
sal_cc_monthly_list <- list()

for (y in years_all) {
  for (m in months_all) {
    
    if (y <= 2024) {
      nm_base <- sprintf("Sal_%d_%02d", y, m)
      r_cc <- sal_baseline_monthly_stack[[nm_base]]
      
    } else {
      if (y <= 2099) {
        y_clim <- y
      } else {
        y_clim <- 2099
      }
      
      key <- paste0(y_clim, "_", sprintf("%02d", m))
      a <- anom_sss_month[[key]]
      if (is.null(a) || is.na(a)) {
        stop("Missing SSS monthly anomaly for ", key)
      }
      
      r_cc <- get_sal_2024_month(m)
      r_cc[sea_mask] <- r_cc[sea_mask] + a
    }
    
    out_name <- sprintf("Sal_CC_%d_%02d", y, m)
    sal_cc_monthly_list[[out_name]] <- r_cc
  }
}

sal_cc_monthly_stack <- rast(sal_cc_monthly_list)
sal_cc_monthly_stack
```

Save monthly CC data:
```{r}
# Temperature CC monthly .asc
for (y in years_all) {
  for (m in months_all) {
    lyr_name <- sprintf("Temp_CC_%d_%02d", y, m)
    r <- temp_cc_monthly_stack[[lyr_name]]
    
    r0 <- r
    r0[!sea_mask] <- 0
    
    fn <- file.path(
      out_temp_cc,
      sprintf("Temp_CC_SSP3-7.0_%d_%02d.asc", y, m)
    )
    
    writeRaster(r0, filename = fn, overwrite = TRUE)
  }
}

# Salinity CC monthly .asc
for (y in years_all) {
  for (m in months_all) {
    lyr_name <- sprintf("Sal_CC_%d_%02d", y, m)
    r <- sal_cc_monthly_stack[[lyr_name]]
    
    r0 <- r
    r0[!sea_mask] <- 0
    
    fn <- file.path(
      out_sal_cc,
      sprintf("Sal_CC_SSP3-7.0_%d_%02d.asc", y, m)
    )
    
    writeRaster(r0, filename = fn, overwrite = TRUE)
  }
}
```

Now we have the monthly layers of temperature and salinity from 1991-2150, we can calculate for each year and for each species (with available response data) a suitability raster.Stored in separate folders under "suitability_temp/species-name" and "suitability_sal/species-name".

Helper function: suitability from monthly rasters
```{r}
library(terra)
library(readr)
library(dplyr)

# --- helper: safe species name for filenames
safe_name <- function(x) {
  x %>%
    gsub("[[:space:]]+", "_", .) %>%     # spaces -> _
    gsub("[^A-Za-z0-9_\\-]", "", .)      # drop weird chars
}

# --- compute annual mean response raster from 12 monthly env rasters
# r_month: SpatRaster with 12 layers (months)
# x_grid: numeric vector of env values in response curve (Var_resp$variable)
# y_curve: numeric vector of response (0-1) for ONE species
annual_mean_response <- function(r_month, x_grid, y_curve) {
  terra::app(
    r_month,
    fun = function(v) {
      # v is length-12 vector (one cell, 12 months)
      if (all(is.na(v))) return(NA_real_)
      # interpolate response per month, then average
      mean(approx(x = x_grid, y = y_curve, xout = v, rule = 2)$y, na.rm = TRUE)
    }
  )
}
```

Helper function: build suitability rasters and save per scenario
```{r}
build_suitability_yearly_rasters <- function(
  in_dir,
  out_dir,
  file_prefix,
  resp_csv,
  years = 1991:2150,
  months = 1:12,
  sea_mask = NULL
) {
  dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

  Var_resp <- readr::read_csv(resp_csv, show_col_types = FALSE) |> as.data.frame()
  stopifnot("variable" %in% names(Var_resp))

  x_grid <- Var_resp$variable
  species_names <- setdiff(names(Var_resp), "variable")
  message("Found ", length(species_names), " species in response table.")

  safe_name <- function(x) {
    x |>
      gsub("[[:space:]]+", "_", x = _) |>
      gsub("[^A-Za-z0-9_\\-]", "", x = _)
  }

  annual_mean_response <- function(r_month, x_grid, y_curve) {
    terra::app(
      r_month,
      fun = function(v) {
        if (all(is.na(v))) return(NA_real_)
        mean(approx(x = x_grid, y = y_curve, xout = v, rule = 2)$y, na.rm = TRUE)
      }
    )
  }

  for (y in years) {

    f_month <- file.path(
      in_dir,
      sprintf("%s_%d_%02d.asc", file_prefix, y, months)
    )
    missing <- f_month[!file.exists(f_month)]
    if (length(missing) > 0) {
      stop("Missing monthly rasters for year ", y, ". Example missing file: ", missing[1])
    }

    r_month <- terra::rast(f_month)

    # Prevent land zeros from biasing suitability
    if (!is.null(sea_mask)) {
      r_month[!sea_mask] <- NA
    }

    for (sp in species_names) {

      sp_safe <- safe_name(sp)

      # Create species subfolder
      sp_dir <- file.path(out_dir, sp_safe)
      dir.create(sp_dir, recursive = TRUE, showWarnings = FALSE)

      y_curve <- Var_resp[[sp]]
      r_suit  <- annual_mean_response(r_month, x_grid, y_curve)

      # Write 0 on land
      r_out <- r_suit
      if (!is.null(sea_mask)) {
        r_out[!sea_mask] <- 0
      } else {
        r_out[is.na(r_out)] <- 0
      }

      fn_out <- file.path(
        sp_dir,
        sprintf("Suit_%s_%s_%d.asc", file_prefix, sp_safe, y)
      )

      terra::writeRaster(r_out, filename = fn_out, overwrite = TRUE)
    }

    message("Done year: ", y)
  }

  invisible(TRUE)
}
```

Apply function to baseline and climate change scenario, for both temperature and salinity:
Define directories
```{r}
base_dir <- "C:/Users/dlorr/OneDrive - UGent/Documents/PhD/EwE_OWF/Spatial data"
out_root <- file.path(base_dir, "Processed")

# Monthly env rasters already exist here:
out_temp_baseline <- file.path(out_root, "Temperature", "Baseline")
out_sal_baseline  <- file.path(out_root, "Salinity",   "Baseline")

out_temp_cc <- file.path(out_root, "Temperature", "ClimateChange_SSP3-7.0")
out_sal_cc  <- file.path(out_root, "Salinity",   "ClimateChange_SSP3-7.0")

# Suggested output structure for suitability rasters:
out_suit_temp_baseline <- file.path(out_root, "Suitability", "Temperature", "Baseline")
out_suit_sal_baseline  <- file.path(out_root, "Suitability", "Salinity",   "Baseline")

out_suit_temp_cc <- file.path(out_root, "Suitability", "Temperature", "ClimateChange_SSP3-7.0")
out_suit_sal_cc  <- file.path(out_root, "Suitability", "Salinity",   "ClimateChange_SSP3-7.0")
```

To reduce calculation time: use same approach as for original layer calculation:
Baseline: Calculate species suitabilit for 1993-2024
1991 and 1992 --> use 1993
2025-2150 --> use 2024

Climate Change: Calculate species suitabilit for 1993-2099
1991 and 1992 --> use 1993
2100 - 2150 --> use 2099
Baseline suitabilities for Temp ans Sal:
```{r}
# response curves CSVs (use your actual paths)
resp_temp_csv <- "Temp_Ecosim_format_DL_CC_adapted.csv"
resp_sal_csv  <- "Sal_Ecosim_format_DL_CC_adapted.csv"

years_all <- 1991:2150

# baseline: 1991-1992 same as 1993, and 2025-2150 same as 2024
year_map_baseline <- setNames(years_all, years_all)
year_map_baseline[c("1991","1992")] <- 1993
year_map_baseline[as.character(2025:2150)] <- 2024

# unique source years you actually need to compute
source_years_baseline <- sort(unique(unname(year_map_baseline)))

# ------------------------------------------------------------
# Helper to copy computed source-year suitability rasters
# to all target years (keeps per-year files for EwE)
# ------------------------------------------------------------
copy_suitability_years <- function(out_dir, file_prefix, year_map) {
  # out_dir has species subfolders
  species_dirs <- list.dirs(out_dir, recursive = FALSE, full.names = TRUE)

  for (sp_dir in species_dirs) {
    sp <- basename(sp_dir)

    for (y in as.integer(names(year_map))) {
      y_src <- as.integer(year_map[[as.character(y)]])
      if (y == y_src) next

      src <- file.path(sp_dir, sprintf("Suit_%s_%s_%d.asc", file_prefix, sp, y_src))
      dst <- file.path(sp_dir, sprintf("Suit_%s_%s_%d.asc", file_prefix, sp, y))

      if (!file.exists(src)) {
        stop("Missing source file: ", src)
      }
      if (!file.exists(dst)) {
        file.copy(src, dst, overwrite = TRUE)
      }
    }
  }
  invisible(TRUE)
}

# Temperature baseline
build_suitability_yearly_rasters(
  in_dir      = out_temp_baseline,
  out_dir     = out_suit_temp_baseline,
  file_prefix = "Temp_Baseline",
  resp_csv    = resp_temp_csv,
  years       = source_years_baseline,
  sea_mask    = sea_mask
)

copy_suitability_years(
  out_dir     = out_suit_temp_baseline,
  file_prefix = "Temp_Baseline",
  year_map    = year_map_baseline
)

# Salinity baseline
years_all <- 1991:2150

year_map_baseline <- setNames(years_all, as.character(years_all))
year_map_baseline[c("1991","1992")] <- 1993
year_map_baseline[as.character(2025:2150)] <- 2024

source_years_baseline <- sort(unique(unname(year_map_baseline)))

build_suitability_yearly_rasters(
  in_dir      = out_sal_baseline,
  out_dir     = out_suit_sal_baseline,
  file_prefix = "Sal_Baseline",
  resp_csv    = resp_sal_csv,
  years       = source_years_baseline,
  sea_mask    = sea_mask
)


copy_suitability_years(
  out_dir     = out_suit_sal_baseline,
  file_prefix = "Sal_Baseline",
  year_map    = year_map_baseline
)

```

CC suitabilities for Temp and Sal:
```{r}
# Response curves CSVs (paths)
resp_temp_csv <- "Temp_Ecosim_format_DL_CC_adapted.csv"
resp_sal_csv  <- "Sal_Ecosim_format_DL_CC_adapted.csv"

years_all <- 1991:2150

# ------------------------------------------------------------
# Year mapping for CC
# - 1991 & 1992 identical to 1993 (as you mentioned)
# - 2100–2150 identical to 2099 (because you capped y_clim at 2099)
# ------------------------------------------------------------
year_map_cc <- setNames(years_all, as.character(years_all))
year_map_cc[c("1991","1992")] <- 1993
year_map_cc[as.character(2100:2150)] <- 2099

source_years_cc <- sort(unique(unname(year_map_cc)))

# ------------------------------------------------------------
# Helper to copy computed source-year suitability rasters
# to all target years (keeps per-year files for EwE)
# ------------------------------------------------------------
copy_suitability_years <- function(out_dir, file_prefix, year_map) {
  # out_dir has species subfolders
  species_dirs <- list.dirs(out_dir, recursive = FALSE, full.names = TRUE)

  for (sp_dir in species_dirs) {
    sp <- basename(sp_dir)

    for (y in as.integer(names(year_map))) {
      y_src <- as.integer(year_map[[as.character(y)]])
      if (y == y_src) next

      src <- file.path(sp_dir, sprintf("Suit_%s_%s_%d.asc", file_prefix, sp, y_src))
      dst <- file.path(sp_dir, sprintf("Suit_%s_%s_%d.asc", file_prefix, sp, y))

      if (!file.exists(src)) {
        stop("Missing source file: ", src)
      }
      if (!file.exists(dst)) {
        file.copy(src, dst, overwrite = TRUE)
      }
    }
  }
  invisible(TRUE)
}

# ============================================================
# TEMPERATURE — CC
# ============================================================

build_suitability_yearly_rasters(
  in_dir      = out_temp_cc,
  out_dir     = out_suit_temp_cc,
  file_prefix = "Temp_CC_SSP3-7.0",
  resp_csv    = resp_temp_csv,
  years       = source_years_cc,
  sea_mask    = sea_mask
)

copy_suitability_years(
  out_dir     = out_suit_temp_cc,
  file_prefix = "Temp_CC_SSP3-7.0",
  year_map    = year_map_cc
)

# ============================================================
# SALINITY — CC
# ============================================================

build_suitability_yearly_rasters(
  in_dir      = out_sal_cc,
  out_dir     = out_suit_sal_cc,
  file_prefix = "Sal_CC_SSP3-7.0",
  resp_csv    = resp_sal_csv,
  years       = source_years_cc,
  sea_mask    = sea_mask
)

copy_suitability_years(
  out_dir     = out_suit_sal_cc,
  file_prefix = "Sal_CC_SSP3-7.0",
  year_map    = year_map_cc
)
```


```{r}
## shapefiles:
# Study area (SBNS)
("C:\Users\dlorr\OneDrive - UGent\Documents\PhD\EwE_OWF\Spatial data\SBNS_boundary_epsg3857.shp")


# Habitat layers (Folk 5 - high resolution)
("C:\Users\dlorr\OneDrive - UGent\Documents\PhD\EwE_OWF\Spatial data\SedimentHabitats\seabed_substrate_1k5_multiscale.shp")


# Offshore wind farm polygons (Operational and planned)
("C:\Users\dlorr\OneDrive - UGent\Documents\PhD\EwE_OWF\Spatial data\OffshoreWindFarms\EMODnet_HA_Energy_WindFarms_pg_20250825.shp")

# ----------------------------------------------------
## Rasters:
# Primary production
("C:\Users\dlorr\OneDrive - UGent\Documents\PhD\EwE_OWF\Spatial data\PrimaryProduction_1km_98-25")
# Files with structure: 19980101-19980131_cmems_obs-oc_atl_bgc-pp_my_l4-multi-1km_P1M.nc
# -->This is a file of the year 1998, January
# 20030201-20030228_cmems_obs-oc_atl_bgc-pp_my_l4-multi-1km_P1M.nc
# --> this is a file of the year 2003, February

# Depth
("C:\Users\dlorr\OneDrive - UGent\Documents\PhD\EwE_OWF\Spatial data\Depth")
# 4 Files in this map:
# Depth_Tile_D4.tiff
# Depth_Tile_D5.tiff
# Depth_Tile_E4.tiff
# Depth_Tile_E5.tiff
# --> for study area (SBNS) these have to be merged and clipped to SBNS extend.

# Temperature
("C:\Users\dlorr\OneDrive - UGent\Documents\PhD\EwE_OWF\Spatial data\Temperature")
# Files with structure:
# CMEMS_v6r1_IBI_PHY_MY_NL_01mav_temp_19930101_19930131_R20251125_RE01.nc
# This is a file of the year 1993, January
# CMEMS_v6r1_IBI_PHY_MY_NL_01mav_temp_19970601_19970630_R20251125_RE01.nc
# This is a file of the year 1997, June

# Salinity
("C:\Users\dlorr\OneDrive - UGent\Documents\PhD\EwE_OWF\Spatial data\Salinity")
# Files with structure:
# CMEMS_v6r1_IBI_PHY_MY_NL_01mav_sal_19930101_19930131_R20251125_RE01.nc
# This is a file of the year 1993, January
# CMEMS_v6r1_IBI_PHY_MY_NL_01mav_sal_19961201_19961231_R20251125_RE01.nc
# This is a file of the year 1996, December

# Dissoleved Oxygen
("C:\Users\dlorr\OneDrive - UGent\Documents\PhD\EwE_OWF\Spatial data\OxygenConc")
# Files with structure:
# metoffice_foam1_amm7_NWS_DOXY_mm199301.nc
# This is a file of the year 1993, January
# metoffice_foam1_amm7_NWS_DOXY_mm200005.nc
# This is a file of the year 2000, May
```

Base map grid --> check
- depth map
- 1km x 1km
- 1 field, make sure all grid cells are connected (no lose "islands" in active cells - islands are set to 0 - only keep biggest one (i.e., SBNS study area))
- base grid structure on PP grid (1km x 1km?)
```{r}



```

Using this base map, sample spatial (temporal) data layers for environmental and habitat data inputs

Primary production - mean 1998-2025
```{r}
# unit input = mgC/m²/day
# unit should be mgC/m³/day --> devide by depth (m)



```

Depth - resampled from high resolution grid to base map grid --> done
```{r}
# 0 = land, (positive) values = depth
# unit = meters



```

Distance to land
```{r}
# unit = km



```


Coastal - Marine
```{r}
# 12 nautical miles from shore = Coastal
# further away = Marine

# Check values input Püts et al. (2020)
# probably have to make 2 of these.
# 1) Coastal = 0.75, marine = 1
# 2) Coastal = 1, marine = 0.75



```


Sediment habitats - proportion of habitat in cell
```{r}
# Sum of all habitat layers should thus equal 1 for each cell!
# Shapefile should be transformed in 5 habitat rasters (one for each habitat)



```

Temperature - yearly average SST in cell
[NOTE: Below the historical data. Future projections added later in the script]
```{r}
# unit = °C



```

Salinity - yearly average SSS in cell
[NOTE: Below the historical data. Future projections added later in the script]
```{r}
# unit = psu



```


