---
title: "ENA_processing"
author: "Dries Lorré"
date: "2026-03-16"
output: html_document
---

```{r}
library(enaR)
library(NetIndices)
library(igraph)
library(dplyr)
library(sf)
library(terra)
```
Directories
```{r}
scenario_dirs <- c(
  Baseline  = "C:/Users/dlorr/OneDrive - UGent/Documents/EwE output/SBNS_CC_adapted_spatial_baseline_test",
  CC        = "C:/Users/dlorr/OneDrive - UGent/Documents/EwE output/SBNS_CC_adapted_spatial_CC_test",
  OWF       = "C:/Users/dlorr/OneDrive - UGent/Documents/EwE output/SBNS_CC_adapted_spatial_OWF_test",
  OWF_CC    = "C:/Users/dlorr/OneDrive - UGent/Documents/EwE output/SBNS_CC_adapted_spatial_OWF_CC_test"
)

csv_subfolder <- "Test_Env_resps"

# load pre-cropped land object
land_main <- readRDS(
  "land_main_study_area.rds"
)
land_main_df <- land_main |>
  st_cast("MULTIPOLYGON") |>
  st_cast("POLYGON") |>
  mutate(poly_id = row_number())

land_main_df_plot <- do.call(
  rbind,
  lapply(seq_len(nrow(land_main_df)), function(i) {
    coords <- st_coordinates(land_main_df[i, ])
    data.frame(
      x = coords[, 1],
      y = coords[, 2],
      poly_id = i
    )
  })
)

# 
get_timestep_path <- function(scenario, timestep = "01920") {
  file.path(
    scenario_dirs[[scenario]],
    csv_subfolder,
    "ena_data",
    paste0("TimeStep=", timestep)
  )
}

ts_path <- get_timestep_path("Baseline", "01920")
# Test file 
file_test <- list.files(ts_path, pattern = "\\.txt$", full.names = TRUE)[1]
file_test
```
read SCOR file and run ENA modules
```{r}
net <- read.scor(file_test)

flow <- enaFlow(net)
asc  <- enaAscendency(net)
tro  <- enaTroAgg(net)

```

Test if we can calculate all indicators:
```{r}
FCI <- as.numeric(flow$ns[1, "FCI"])
H   <- as.numeric(asc[1, "H"])

# Candidate 1: total relative overhead
RDC_OH_CAP <- as.numeric(asc[1, "OH.CAP"])

# Candidate 2: internal overhead / total capacity
RDC_internal_CAP <- as.numeric(asc[1, "OH.internal"]) / as.numeric(asc[1, "CAP"])

```

SOI
```{r}
library(network)
# Extract flow matrix to calculate omnivory
F    <- as.matrix.network.adjacency(net, attrname = "flow")   # for SOI

# Check
# dim(F)          OK
# F[1:5, 1:5]     OK
# rownames(F)     OK
# colnames(F)     OK

library(NetIndices)

# dead matter compartments
dead_nodes <- c("Detritus", "Discards")

# trophic levels and node omnivory
ti <- TrophInd(
  Flow = F,
  Dead = dead_nodes
)

# head(ti)
# summary(ti$TL)
# summary(ti$OI) # --> OI for individual species

# weighted system omnivory index
Q <- colSums(F)
consumers <- !(rownames(F) %in% dead_nodes)

valid_exact <- consumers & is.finite(ti$OI) & is.finite(Q) & Q > 0

SOI_exact <- sum(ti$OI[valid_exact] * log(Q[valid_exact])) /
             sum(log(Q[valid_exact])) #correct calculation!
# SOI_unweighted <- mean(ti$OI[consumers & is.finite(ti$OI)])
# valid_Q <- consumers & is.finite(ti$OI) & is.finite(Q) & Q > 0
# 
# SOI_Qweighted <- sum(ti$OI[valid_Q] * Q[valid_Q]) /
#                  sum(Q[valid_Q])
# Combination of all metrics for test cell
one_cell_metrics <- data.frame(
  file = basename(file_test),
  FCI = FCI,
  H = H,
  RDC_OH_CAP = RDC_OH_CAP,
  RDC_internal_CAP = RDC_internal_CAP,
  SOI_exact = SOI_exact #,
  # SOI_unweighted = SOI_unweighted,
  # SOI_Qweighted = SOI_Qweighted
)
one_cell_metrics
```

SCRIPT FOR ALL CELLS AND ALL SCENARIO'S
Helper function for one SCOR file
```{r}
library(enaR)
library(NetIndices)
library(network)
library(dplyr)
library(purrr)
library(stringr)

compute_ENA_indices <- function(file,
                                dead_nodes = c("Detritus", "Discards")) {
  out <- tryCatch({
    net <- read.scor(file)

    flow <- enaFlow(net)
    asc  <- enaAscendency(net)

    F <- as.matrix.network.adjacency(net, attrname = "flow")

    ti <- TrophInd(
      Flow = F,
      Dead = dead_nodes
    )

    Q <- colSums(F)
    consumers <- !(rownames(F) %in% dead_nodes)
    valid <- consumers & is.finite(ti$OI) & is.finite(Q) & Q > 0

    SOI <- sum(ti$OI[valid] * log(Q[valid])) / sum(log(Q[valid]))

    data.frame(
      file = basename(file),
      FCI = as.numeric(flow$ns[1, "FCI"]),
      H = as.numeric(asc[1, "H"]),
      RDC_OH_CAP = as.numeric(asc[1, "OH.CAP"]),
      RDC_internal_CAP = as.numeric(asc[1, "OH.internal"]) / as.numeric(asc[1, "CAP"]),
      SOI = SOI,
      stringsAsFactors = FALSE
    )
  }, error = function(e) {
    data.frame(
      file = basename(file),
      FCI = NA_real_,
      H = NA_real_,
      RDC_OH_CAP = NA_real_,
      RDC_internal_CAP = NA_real_,
      SOI = NA_real_,
      stringsAsFactors = FALSE
    )
  })

  out
}
```

Test function for 1 time step

In total, there are 9739 cells, for each time step, for each scenario for which ENA indicators are calculated.
```{r}
scenario_dirs <- c(
  Baseline  = "C:/Users/dlorr/OneDrive - UGent/Documents/EwE output/SBNS_CC_adapted_spatial_baseline_test",
  CC        = "C:/Users/dlorr/OneDrive - UGent/Documents/EwE output/SBNS_CC_adapted_spatial_CC_test",
  OWF       = "C:/Users/dlorr/OneDrive - UGent/Documents/EwE output/SBNS_CC_adapted_spatial_OWF_test",
  OWF_CC    = "C:/Users/dlorr/OneDrive - UGent/Documents/EwE output/SBNS_CC_adapted_spatial_OWF_CC_test"
)

csv_subfolder <- "Test_Env_resps"

get_timestep_path <- function(scenario, timestep = "01920") {
  file.path(
    scenario_dirs[[scenario]],
    csv_subfolder,
    "ena_data",
    paste0("TimeStep=", timestep)
  )
}

ts_path <- get_timestep_path("Baseline", "01920")

files <- list.files(ts_path, pattern = "\\.txt$", full.names = TRUE)

# Use multisession, using all cores to process the data to speed up.
install.packages("furrr")
library(furrr)
library(future)

plan(multisession, workers = 10)

ena_results_1920 <- future_map_dfr(
  files,
  compute_ENA_indices,
  .progress = TRUE
)

ena_results_1920 <- purrr::map_dfr(files, compute_ENA_indices)

head(ena_results_1920)
summary(ena_results_1920)
```

Full workflow for all scenario's

Install all packages
```{r}
library(enaR)
library(NetIndices)
library(network)
library(dplyr)
library(purrr)
library(furrr)
library(stringr)
library(terra)
library(future)

plan(multisession, workers = 10)
```

Paths and settings
```{r}
scenario_dirs <- c(
  Baseline  = "C:/Users/dlorr/OneDrive - UGent/Documents/EwE output/SBNS_CC_adapted_spatial_baseline_test",
  CC        = "C:/Users/dlorr/OneDrive - UGent/Documents/EwE output/SBNS_CC_adapted_spatial_CC_test",
  OWF       = "C:/Users/dlorr/OneDrive - UGent/Documents/EwE output/SBNS_CC_adapted_spatial_OWF_test",
  OWF_CC    = "C:/Users/dlorr/OneDrive - UGent/Documents/EwE output/SBNS_CC_adapted_spatial_OWF_CC_test"
)

csv_subfolder <- "Test_Env_resps"
timesteps <- sprintf("%04d", 1909:1920)   # because filenames/folders use 01909 etc below
dead_nodes <- c("Detritus", "Discards")

get_timestep_path <- function(scenario, timestep) {
  file.path(
    scenario_dirs[[scenario]],
    csv_subfolder,
    "ena_data",
    paste0("TimeStep=0", timestep)
  )
}
```

Helper function for ENA analysis on one file
```{r}
compute_ENA_indices <- function(file, dead_nodes = c("Detritus", "Discards")) {
  
  out <- tryCatch({
    net <- read.scor(file)
    
    flow <- enaFlow(net)
    asc  <- enaAscendency(net)
    
    F <- as.matrix.network.adjacency(net, attrname = "flow")
    
    ti <- TrophInd(
      Flow = F,
      Dead = dead_nodes
    )
    
    Q <- colSums(F)
    consumers <- !(rownames(F) %in% dead_nodes)
    valid_soi <- consumers & is.finite(ti$OI) & is.finite(Q) & Q > 0
    
    soi_num <- sum(ti$OI[valid_soi] * log(Q[valid_soi]))
    soi_den <- sum(log(Q[valid_soi]))
    SOI <- soi_num / soi_den
    
    fname <- basename(file)
    
    row_id <- as.integer(str_extract(fname, "^\\d{4}"))
    col_id <- as.integer(str_extract(fname, "(?<=^)\\d{4}(?=-\\d{4})"))  # not used
    # better extraction:
    parts <- str_match(fname, "^(\\d{4})-(\\d{4})-")
    row_id <- as.integer(parts[,2])
    col_id <- as.integer(parts[,3])
    
    FCI <- as.numeric(flow$ns[1, "FCI"])
    H   <- as.numeric(asc[1, "H"])
    RDC_OH_CAP <- as.numeric(asc[1, "OH.CAP"])
    RDC_internal_CAP <- as.numeric(asc[1, "OH.internal"]) / as.numeric(asc[1, "CAP"])
    
    # optional filtered versions for mapping/averaging
    FCI_map <- ifelse(is.finite(FCI) & FCI >= 0 & FCI <= 1, FCI, NA_real_)
    SOI_map <- ifelse(is.finite(SOI) & SOI > -1 & SOI < 1, SOI, NA_real_)
    
    data.frame(
      file = fname,
      row = row_id,
      col = col_id,
      FCI = FCI,
      H = H,
      RDC_OH_CAP = RDC_OH_CAP,
      RDC_internal_CAP = RDC_internal_CAP,
      SOI = SOI,
      FCI_map = FCI_map,
      SOI_map = SOI_map,
      stringsAsFactors = FALSE
    )
    
  }, error = function(e) {
    fname <- basename(file)
    parts <- str_match(fname, "^(\\d{4})-(\\d{4})-")
    row_id <- as.integer(parts[,2])
    col_id <- as.integer(parts[,3])
    
    data.frame(
      file = fname,
      row = row_id,
      col = col_id,
      FCI = NA_real_,
      H = NA_real_,
      RDC_OH_CAP = NA_real_,
      RDC_internal_CAP = NA_real_,
      SOI = NA_real_,
      FCI_map = NA_real_,
      SOI_map = NA_real_,
      stringsAsFactors = FALSE
    )
  })
  
  out
}
```

Helper function to run this code for one time step, for one scenario
```{r}
run_one_timestep <- function(scenario, timestep, dead_nodes) {

  ts_path <- get_timestep_path(scenario, timestep)

  files <- list.files(ts_path, pattern = "\\.txt$", full.names = TRUE)

  res <- furrr::future_map_dfr(
    files,
    compute_ENA_indices,
    dead_nodes = dead_nodes
  )

  res$scenario <- scenario
  res$timestep <- timestep

  res
}
```

One scenario, all time steps
```{r}
run_one_scenario <- function(scenario, timesteps, dead_nodes = c("Detritus", "Discards")) {
  
  purrr::map_dfr(timesteps, function(ts) {
    message("Running ", scenario, " timestep ", ts)
    run_one_timestep(scenario, ts, dead_nodes = dead_nodes)
  })
}
```

All scenario's
```{r, echo=FALSE}
all_results <- purrr::map_dfr(names(scenario_dirs), function(sc) {
  message("========== Scenario: ", sc, " ==========")
  run_one_scenario(sc, timesteps, dead_nodes = dead_nodes)
})

# Save restuls
saveRDS(all_results, 
        "C:/Users/dlorr/OneDrive - UGent/Documents/PhD/EwE_OWF/ENA_all_results.rds")

write.csv(
  all_results,
  "C:/Users/dlorr/OneDrive - UGent/Documents/PhD/EwE_OWF/ENA_all_results.csv",
  row.names = FALSE
)
```

```{r}
safe_mean <- function(x) {
  if (all(is.na(x))) NA_real_ else mean(x, na.rm = TRUE)
}

# optional filtering for unstable values
# all_results <- all_results %>%
#   mutate(
#     FCI_map = ifelse(is.finite(FCI) & FCI >= 0 & FCI <= 1, FCI, NA_real_),
#     SOI_map = ifelse(is.finite(SOI) & SOI > -1 & SOI < 1, SOI, NA_real_)
#   )

# average over timesteps per scenario and cell
avg_results <- all_results %>%
  group_by(scenario, row, col) %>%
  summarise(
    FCI = safe_mean(FCI_map),
    H = safe_mean(H),
    RDC_OH_CAP = safe_mean(RDC_OH_CAP),
    RDC_internal_CAP = safe_mean(RDC_internal_CAP),
    SOI = safe_mean(SOI_map),
    .groups = "drop"
  )
```

Conversion to raster
```{r}
# function to write values into the template
df_to_raster_template <- function(df, value_col, template) {
  r <- template
  values(r) <- NA_real_
  cell_ids <- cellFromRowCol(r, df$row, df$col)
  r[cell_ids] <- df[[value_col]]
  r
}
```

```{r}
template_raster <- rast("Temp_Baseline_1991.asc")
baseline_avg <- avg_results %>% filter(scenario == "Baseline")

r_H_baseline <- df_to_raster_template(baseline_avg, "H", template_raster)
r_SOI_baseline <- df_to_raster_template(baseline_avg, "SOI", template_raster)
r_FCI_baseline <- df_to_raster_template(baseline_avg, "FCI", template_raster)
r_RDC_baseline <- df_to_raster_template(baseline_avg, "RDC_internal_CAP", template_raster)
r_RDCOH_baseline <- df_to_raster_template(baseline_avg, "RDC_OH_CAP", template_raster)

plot(r_H_baseline)
plot(r_SOI_baseline)
plot(r_FCI_baseline)
plot(r_RDC_baseline)
plot(r_RDCOH_baseline)

```

Rasters for all scenario's
```{r}
indicators <- c("FCI", "H", "RDC_OH_CAP", "RDC_internal_CAP", "SOI")
scenarios <- unique(avg_results$scenario)

raster_list <- list()

for (sc in scenarios) {
  df_sc <- avg_results %>% filter(scenario == sc)
  raster_list[[sc]] <- list()

  for (ind in indicators) {
    raster_list[[sc]][[ind]] <- df_to_raster_template(df_sc, ind, template_raster)
  }
}

out_dir <- "C:/Users/dlorr/OneDrive - UGent/Documents/PhD/EwE_OWF/R-scripts/ENA_processing/ENA_rasters"
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)

for (sc in names(raster_list)) {
  for (ind in names(raster_list[[sc]])) {
    writeRaster(
      raster_list[[sc]][[ind]],
      file.path(out_dir, paste0("ENA_", ind, "_", sc, "_mean.tif")),
      overwrite = TRUE
    )
  }
}

# plot(raster_list$OWF$SOI)
# plot(raster_list$CC$H)
# plot(raster_list$OWF_CC$RDC_internal_CAP)
```

PLOTTING OF ENA DATA
```{r}
library(terra)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(scales)
library(stringr)
library(tibble)
library(grid)
```


```{r}
# =========================================================
# DIRECTORIES
# =========================================================
ena_dir <- "C:/Users/dlorr/OneDrive - UGent/Documents/PhD/EwE_OWF/R-scripts/ENA_processing/ENA_rasters"
zone_raster_file <- file.path(ena_dir, "OWF_MPA_2150.asc")

out_dir <- file.path(ena_dir, "ENA_plots")
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)
```


```{r}
# =========================================================
# SETTINGS
# =========================================================
indicators <- c("FCI", "H", 
                # "RDC_OH_CAP",
                "RDC_internal_CAP", "SOI")

indicator_label_map <- c(
  "FCI" = "FCI",
  "H" = "H",
  "RDC_internal_CAP" = "RDC",
  "SOI" = "SOI"
)

scenario_order_raw <- c("Baseline", "CC", "OWF", "OWF_CC")

scenario_order_plot <- c("CC", "OWF", "OWF + CC")

scenario_label_map <- c(
  "Baseline" = "Baseline",
  "CC"       = "CC",
  "OWF"      = "OWF",
  "OWF_CC"   = "OWF + CC"
)
```


```{r}
# =========================================================
# HELPER: READ ONE ENA RASTER
# =========================================================
read_ena_raster <- function(indicator, scenario, ena_dir) {
  
  f <- file.path(
    ena_dir,
    paste0("ENA_", indicator, "_", scenario, "_mean.tif")
  )
  
  if (!file.exists(f)) {
    stop("File not found: ", f)
  }
  
  r <- rast(f)
  
  df <- as.data.frame(r, xy = TRUE, cells = TRUE, na.rm = FALSE)
  names(df)[names(df) == names(r)] <- "value"
  
  df %>%
    transmute(
      cell = cell,
      Lon = x,
      Lat = y,
      value = value,
      indicator = indicator,
      scenario_raw = scenario
    )
}
```


```{r}
# =========================================================
# READ ALL RASTERS
# =========================================================
ena_df <- bind_rows(
  lapply(scenario_order_raw, function(sc) {
    bind_rows(
      lapply(indicators, function(ind) {
        read_ena_raster(ind, sc, ena_dir)
      })
    )
  })
) %>%
  mutate(
    scenario = recode(scenario_raw, !!!scenario_label_map),
    indicator = factor(indicator, levels = indicators)
  )
```


```{r}
# =========================================================
# BASELINE AND % CHANGE VS BASELINE
# =========================================================
baseline_ref <- ena_df %>%
  filter(scenario == "Baseline") %>%
  select(cell, indicator, baseline_value = value)

scenario_change_df <- ena_df %>%
  filter(scenario != "Baseline") %>%
  left_join(baseline_ref, by = c("cell", "indicator")) %>%
  mutate(
    pct_change = case_when(
      baseline_value == 0 & value == 0 ~ 0,
      baseline_value == 0 & value != 0 ~ NA_real_,
      is.na(baseline_value) ~ NA_real_,
      TRUE ~ ((value - baseline_value) / baseline_value) * 100
    ),
    scenario = factor(scenario, levels = scenario_order_plot)
  )

baseline_df <- ena_df %>%
  filter(scenario == "Baseline") %>%
  mutate(indicator = factor(indicator, levels = indicators))
```


```{r}
# =========================================================
# EVALUATION ZONES
# =========================================================
owf_rast <- rast(zone_raster_file)

# binary OWF footprint
owf_bin <- ifel(owf_rast > 0, 1, 0)

# same spillover logic as your previous script
w <- matrix(1, nrow = 5, ncol = 5)

owf_buffered <- focal(
  owf_bin,
  w = w,
  fun = max,
  na.policy = "omit"
)

spillover_bin <- ifel(owf_buffered == 1 & owf_bin == 0, 1, 0)
outside_bin   <- ifel(owf_buffered == 0, 1, 0)

zone_df <- tibble(
  cell = 1:ncell(owf_bin),
  in_owf = values(owf_bin, mat = FALSE),
  in_spill = values(spillover_bin, mat = FALSE),
  in_outside = values(outside_bin, mat = FALSE)
) %>%
  mutate(
    zone = case_when(
      in_owf == 1 ~ "OWF",
      in_spill == 1 ~ "Spillover",
      in_outside == 1 ~ "Outside OWF",
      TRUE ~ NA_character_
    )
  ) %>%
  select(cell, zone)

scenario_change_zones <- scenario_change_df %>%
  left_join(zone_df, by = "cell")
```


```{r}
# =========================================================
# ZONAL SUMMARIES
# =========================================================
zone_summary_overall <- scenario_change_zones %>%
  group_by(indicator, scenario) %>%
  summarise(
    mean_pct_change = mean(pct_change, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(zone = "Overall")

zone_summary_zones <- scenario_change_zones %>%
  filter(
    scenario %in% c("OWF", "OWF + CC"),
    !is.na(zone)
  ) %>%
  group_by(indicator, scenario, zone) %>%
  summarise(
    mean_pct_change = mean(pct_change, na.rm = TRUE),
    .groups = "drop"
  )

zone_summary <- bind_rows(zone_summary_overall, zone_summary_zones) %>%
  mutate(
    zone = factor(zone, levels = c("OWF", "Spillover", "Outside OWF", "Overall"))
  )
```


```{r}
# =========================================================
# ZONE CHECK PLOT
# =========================================================
xy_df <- as.data.frame(owf_bin, xy = TRUE, cells = TRUE, na.rm = FALSE) %>%
  transmute(cell = cell, Lon = x, Lat = y)

zone_plot_df <- xy_df %>%
  left_join(zone_df, by = "cell") %>%
  mutate(zone = factor(zone, levels = c("OWF", "Spillover", "Outside OWF")))

zone_check_plot <- ggplot(zone_plot_df, aes(x = Lon, y = Lat, fill = zone)) +
  geom_raster() +
  coord_equal() +
  scale_x_continuous(breaks = pretty(range(zone_plot_df$Lon, na.rm = TRUE), n = 3)) +
  scale_y_continuous(breaks = pretty(range(zone_plot_df$Lat, na.rm = TRUE), n = 3)) +
  scale_fill_manual(
    values = c(
      "OWF" = "#1f78b4",
      "Spillover" = "#33a02c",
      "Outside OWF" = "grey85"
    ),
    na.value = "white"
  ) +
  labs(
    title = "OWF zones",
    x = "Longitude",
    y = "Latitude",
    fill = "Zone"
  ) +
  theme_bw(base_size = 12) +
  theme(
    panel.grid = element_blank(),
    legend.position = "right"
  )
zone_check_plot
ggsave(
  file.path(out_dir, "OWF_zone_check_map.png"),
  zone_check_plot,
  width = 8,
  height = 5,
  dpi = 800
)
```


```{r}
# =========================================================
# BASELINE MAP SUBPLOT
# =========================================================
make_baseline_indicator_subplot <- function(df_panel,
                                            show_x = FALSE,
                                            show_y = FALSE) {
  
  panel_min <- quantile(df_panel$value, probs = 0.01, na.rm = TRUE)
  panel_max <- quantile(df_panel$value, probs = 0.99, na.rm = TRUE)
  
  if (!is.finite(panel_min)) panel_min <- min(df_panel$value, na.rm = TRUE)
  if (!is.finite(panel_max)) panel_max <- max(df_panel$value, na.rm = TRUE)
  
  if (!is.finite(panel_min)) panel_min <- 0
  if (!is.finite(panel_max)) panel_max <- 1
  
  if (panel_min == panel_max) {
    panel_max <- panel_min + 1e-6
  }
  
  p <- ggplot(df_panel, aes(x = Lon, y = Lat, fill = value)) +
    geom_raster() +
    geom_polygon(
      data = land_main_df_plot,
      aes(x = x, y = y, group = poly_id),
      inherit.aes = FALSE,
      fill = "grey85",
      colour = "grey45",
      linewidth = 0.2
      ) +
coord_equal(expand = FALSE) +
    scale_x_continuous(
      breaks = pretty(range(df_panel$Lon, na.rm = TRUE), n = 3)
    ) +
    scale_y_continuous(
      breaks = pretty(range(df_panel$Lat, na.rm = TRUE), n = 3)
    ) +
    scale_fill_distiller(
      palette = "Spectral",
      direction = -1,
      limits = c(panel_min, panel_max),
      oob = scales::squish,
      breaks = pretty(c(panel_min, panel_max), n = 3),
      na.value = "white",
      guide = guide_colorbar(
        title.position = "top",
        barheight = unit(15, "mm"),
        barwidth  = unit(5, "mm"),
        frame.colour = "black",
        ticks.colour = "black"
      )
    ) +
    labs(
      fill = NULL,
      x = if (show_x) "Longitude" else NULL,
      y = if (show_y) "Latitude" else NULL
    ) +
    theme_bw(base_size = 11) +
    theme(
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 12),
      legend.title = element_blank(),
      legend.text = element_text(size = 11),
      panel.grid = element_blank(),
      panel.background = element_rect(fill = "white", colour = NA),
      legend.position = c(0.98, 0.02),
      legend.justification = c(1, 0),
      legend.box.background = element_rect(
        colour = "black",
        fill = scales::alpha("white", 0.75)
      ),
      legend.background = element_rect(
        fill = scales::alpha("white", 0.75),
        colour = NA
      ),
      plot.margin = margin(2, 2, 2, 2)
    )
  
  if (!show_x) {
    p <- p + theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank()
    )
  }
  
  if (!show_y) {
    p <- p + theme(
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    )
  }
  
  p
}
```


```{r}
# =========================================================
# BASELINE FIGURE (2 x 2 grid)
# =========================================================
plot_baseline_figure <- function(df_baseline, out_dir,
                                 filename = "Baseline_ENA_indicators.png",
                                 width = 10,
                                 height = 10) {
  
  inds <- levels(df_baseline$indicator)
  
  if (length(inds) != 4) {
    warning("This function is set up for exactly 4 indicators (2x2 layout).")
  }
  
  ncol_layout <- 2
  nrow_layout <- 2
  
  indicator_blocks <- lapply(seq_along(inds), function(i) {
    ind <- inds[i]
    df_panel <- df_baseline %>% filter(indicator == ind)
    
    row_pos <- ceiling(i / ncol_layout)
    col_pos <- ((i - 1) %% ncol_layout) + 1
    
    ind_title <- ggplot() +
      annotate(
        "text",
        x = 0.02, y = 0.05,
        label = indicator_label_map[[as.character(ind)]],
        hjust = 0,
        fontface = "bold",
        size = 4.6
      ) +
      coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), clip = "off") +
      theme_void() +
      theme(plot.margin = margin(0, 0, -8, 0))
    
    map_plot <- make_baseline_indicator_subplot(
      df_panel,
      show_x = row_pos == nrow_layout,  # only bottom row
      show_y = col_pos == 1             # only left column
    )
    
    ind_title / map_plot + plot_layout(heights = c(0.02, 0.1))
  })
  
  final_plot <- wrap_plots(indicator_blocks, ncol = ncol_layout)
  
  ggsave(
    filename = file.path(out_dir, filename),
    plot = final_plot,
    width = width,
    height = height,
    dpi = 800
  )
  
  final_plot
}

baseline_figure <- plot_baseline_figure(
  df_baseline = baseline_df,
  out_dir = out_dir,
  filename = "Baseline_ENA_indicators.png",
  width = 9,
  height = 7
)

baseline_figure
```


```{r}
# =========================================================
# BARPLOT
# =========================================================
make_zone_barplot <- function(df_zone_panel) {
  
  if (nrow(df_zone_panel) == 0) {
    return(ggplot() + theme_void())
  }
  
  df_zone_panel <- df_zone_panel %>%
    mutate(
      zone_label = case_when(
        zone == "OWF" ~ "OWF",
        zone == "Spillover" ~ "SPILL",
        zone == "Outside OWF" ~ "OUT",
        zone == "Overall" ~ "ALL",
        TRUE ~ as.character(zone)
      ),
      zone_label = factor(zone_label, levels = c("ALL", "OUT", "OWF", "SPILL")),
      change_dir = ifelse(mean_pct_change < 0, "decrease", "increase")
    )
  
  ggplot(df_zone_panel, aes(x = zone_label, y = mean_pct_change, fill = change_dir)) +
    geom_col(width = 0.4) +
    geom_hline(yintercept = 0, linewidth = 0.3) +
    scale_x_discrete(
      limits = c("ALL", "OUT", "OWF", "SPILL"),
      labels = function(x) {
        if (length(unique(df_zone_panel$zone_label)) == 1) {
          ifelse(x == "ALL", "ALL", "")
        } else {
          x
        }
      }
    ) +
    scale_fill_manual(
      values = c(
        "decrease" = "#E67E22",
        "increase" = "#2C7FB8"
      ),
      guide = "none"
    ) +
    labs(x = NULL, y = NULL) +
    theme_bw(base_size = 11) +
    theme(
      axis.title = element_blank(),
      axis.text.x = element_text(size = 12, angle = 0, hjust = 0.5),
      axis.text.y = element_text(size = 12),
      axis.ticks.x = element_blank(),
      panel.grid = element_blank(),
      plot.margin = margin(-12, 2, 0, 2)
    )
}
```


```{r}
# =========================================================
# MAP + BARPLOT PANEL
# =========================================================
make_scenario_indicator_panel <- function(df_panel,
                                          df_zone_panel,
                                          show_x = FALSE,
                                          show_y = FALSE) {
  
  panel_min <- quantile(df_panel$pct_change, probs = 0.01, na.rm = TRUE)
  panel_max <- quantile(df_panel$pct_change, probs = 0.99, na.rm = TRUE)
  
  if (!is.finite(panel_min)) panel_min <- min(df_panel$pct_change, na.rm = TRUE)
  if (!is.finite(panel_max)) panel_max <- max(df_panel$pct_change, na.rm = TRUE)
  
  if (!is.finite(panel_min)) panel_min <- -1
  if (!is.finite(panel_max)) panel_max <- 1
  
  if (panel_min == panel_max) {
    panel_min <- panel_min - 1
    panel_max <- panel_max + 1
  }
  
  map_plot <- ggplot(df_panel, aes(x = Lon, y = Lat, fill = pct_change)) +
    geom_raster() +
    geom_polygon(
      data = land_main_df_plot,
      aes(x = x, y = y, group = poly_id),
      inherit.aes = FALSE,
      fill = "grey85",
      colour = "grey45",
      linewidth = 0.2
      ) +
    coord_equal(expand = FALSE) +
    scale_x_continuous(breaks = pretty(range(df_panel$Lon, na.rm = TRUE), n = 3)) +
    scale_y_continuous(breaks = pretty(range(df_panel$Lat, na.rm = TRUE), n = 3)) +
    scale_fill_gradient2(
      low = "#E67E22",
      mid = "white",
      high = "#2C7FB8",
      midpoint = 0,
      limits = c(panel_min, panel_max),
      oob = scales::squish,
      breaks = pretty(c(panel_min, panel_max), n = 3),
      labels = function(x) paste0(round(x, 1), "%"),
      na.value = "white",
      guide = guide_colorbar(
        title.position = "top",
        barheight = unit(13, "mm"),
        barwidth  = unit(4.5, "mm"),
        frame.colour = "black",
        ticks.colour = "black"
      )
    ) +
    labs(
      fill = NULL,
      x = "Longitude",
      y = "Latitude"
    ) +
    theme_bw(base_size = 11) +
    theme(
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 12),
      legend.title = element_blank(),
      legend.text = element_text(size = 11),
      panel.grid = element_blank(),
      panel.background = element_rect(fill = "white", colour = NA),
      legend.position = c(0.98, 0.02),
      legend.justification = c(1, 0),
      legend.box.background = element_rect(
        colour = "black",
        fill = scales::alpha("white", 0.75)
      ),
      legend.background = element_rect(
        fill = scales::alpha("white", 0.75),
        colour = NA
      ),
      plot.margin = margin(2, 2, -6, 2)
    )
  
  if (!show_x) {
    map_plot <- map_plot + theme(
      axis.title.x = element_text(colour = NA),
      axis.text.x  = element_text(colour = NA),
      axis.ticks.x = element_line(colour = NA)
    )
  }
  
  if (!show_y) {
    map_plot <- map_plot + theme(
      axis.title.y = element_text(colour = NA),
      axis.text.y  = element_text(colour = NA),
      axis.ticks.y = element_line(colour = NA)
    )
  }
  
  bar_plot <- make_zone_barplot(df_zone_panel)
  
  map_plot / bar_plot + plot_layout(heights = c(1, 0.6))
}
```


```{r}
# =========================================================
# SCENARIO FIGURE
# =========================================================
plot_grouped_scenario_figure <- function(df_change, zone_summary, indicators,
                                         out_dir, filename,
                                         width = 14, height = 10) {
  
  df_grp <- df_change %>%
    filter(indicator %in% indicators) %>%
    mutate(indicator = factor(indicator, levels = indicators))
  
  header_row <- wrap_plots(
    list(
      ggplot() + theme_void(),
      ggplot() +
        annotate("text", x = 0.375, y = 0.2, label = "CC",
                 fontface = "bold", size = 6) +
        xlim(0, 1) + ylim(0, 1) + theme_void(),
      ggplot() +
        annotate("text", x = 0.375, y = 0.2, label = "OWF",
                 fontface = "bold", size = 6) +
        xlim(0, 1) + ylim(0, 1) + theme_void(),
      ggplot() +
        annotate("text", x = 0.375, y = 0.2, label = "OWF + CC",
                 fontface = "bold", size = 6) +
        xlim(0, 1) + ylim(0, 1) + theme_void()
    ),
    ncol = 4
  ) +
    plot_layout(widths = c(0.9, 3, 3, 3))
  
  row_blocks <- lapply(seq_along(indicators), function(i) {
    ind <- indicators[i]
    df_ind <- df_grp %>% filter(indicator == ind)
    
    row_title <- ggplot() +
      annotate(
        "text",
        x = 0.02, y = 0.25,
        label = indicator_label_map[[as.character(ind)]],
        hjust = 0,
        fontface = "bold",
        size = 4.6
      ) +
      coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), clip = "off") +
      theme_void() +
      theme(plot.margin = margin(0, 0, -8, 0))
    
    scen_levels <- c("CC", "OWF", "OWF + CC")
    
    row_maps <- lapply(seq_along(scen_levels), function(j) {
      sc <- scen_levels[j]
      
      df_panel <- df_ind %>% filter(scenario == sc)
      
      df_zone_panel <- zone_summary %>%
        filter(indicator == ind, scenario == sc)
      
      make_scenario_indicator_panel(
        df_panel = df_panel,
        df_zone_panel = df_zone_panel,
        show_x = i == length(indicators),
        show_y = j == 1
      )
    })
    
    maps_row <- wrap_plots(row_maps, ncol = 3)
    
    row_title / maps_row + plot_layout(heights = c(0.02, 0.1))
  })
  
  maps_block <- wrap_plots(row_blocks, ncol = 1)
  
  final_plot <- header_row / maps_block +
    plot_layout(heights = c(0.04, 1))
  
  ggsave(
    filename = file.path(out_dir, filename),
    plot = final_plot,
    width = width,
    height = height,
    dpi = 800
  )
  
  final_plot
}
```


```{r}
scenario_figure <- plot_grouped_scenario_figure(
  df_change = scenario_change_df,
  zone_summary = zone_summary,
  indicators = indicators,
  out_dir = out_dir,
  filename = "Scenario_ENA_indicators.png",
  width = 14,
  height = 18.5
)
scenario_figure
```

Only maps
```{r}
make_top_row_panel <- function(plot_obj, label) {
  title_plot <- ggplot() +
    annotate(
      "text",
      x = 0.02, y = 0.3,
      label = label,
      hjust = 0,
      fontface = "bold",
      size = 4.8
    ) +
    xlim(0, 1) + ylim(0, 1) +
    theme_void() +
    theme(plot.margin = margin(0, 0, -6, 0))
  
  title_plot / plot_obj + plot_layout(heights = c(0.12, 1))
}

# =========================================================
# MAP ONLY PANEL
# =========================================================
make_scenario_indicator_map <- function(df_panel,
                                        show_x = FALSE,
                                        show_y = FALSE,
                                        indicator_label = NULL) {
  
  panel_min <- quantile(df_panel$pct_change, probs = 0.01, na.rm = TRUE)
  panel_max <- quantile(df_panel$pct_change, probs = 0.99, na.rm = TRUE)
  
  if (!is.finite(panel_min)) panel_min <- min(df_panel$pct_change, na.rm = TRUE)
  if (!is.finite(panel_max)) panel_max <- max(df_panel$pct_change, na.rm = TRUE)
  
  if (!is.finite(panel_min)) panel_min <- -1
  if (!is.finite(panel_max)) panel_max <- 1
  
  if (panel_min == panel_max) {
    panel_min <- panel_min - 1
    panel_max <- panel_max + 1
  }
  
  map_plot <- ggplot(df_panel, aes(x = Lon, y = Lat, fill = pct_change)) +
    geom_raster() +
    geom_polygon(
      data = land_main_df_plot,
      aes(x = x, y = y, group = poly_id),
      inherit.aes = FALSE,
      fill = "grey85",
      colour = "grey45",
      linewidth = 0.2
    ) +
    coord_equal(expand = FALSE, clip = "off") +
    scale_x_continuous(breaks = pretty(range(df_panel$Lon, na.rm = TRUE), n = 3)) +
    scale_y_continuous(breaks = pretty(range(df_panel$Lat, na.rm = TRUE), n = 3)) +
    scale_fill_gradient2(
      low = "#E67E22",
      mid = "white",
      high = "#2C7FB8",
      midpoint = 0,
      limits = c(panel_min, panel_max),
      oob = scales::squish,
      breaks = pretty(c(panel_min, panel_max), n = 3),
      labels = function(x) paste0(round(x, 1), "%"),
      na.value = "white",
      guide = guide_colorbar(
        title.position = "top",
        barheight = unit(13, "mm"),
        barwidth  = unit(4.5, "mm"),
        frame.colour = "black",
        ticks.colour = "black"
      )
    ) +
    labs(
      fill = NULL,
      x = "Longitude",
      y = "Latitude"
    ) +
    theme_bw(base_size = 11) +
    theme(
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 12),
      legend.title = element_blank(),
      legend.text = element_text(size = 11),
      panel.grid = element_blank(),
      panel.background = element_rect(fill = "white", colour = NA),
      legend.position = c(0.98, 0.02),
      legend.justification = c(1, 0),
      legend.box.background = element_rect(
        colour = "black",
        fill = scales::alpha("white", 0.75)
      ),
      legend.background = element_rect(
        fill = scales::alpha("white", 0.75),
        colour = NA
      ),
      plot.margin = margin(22, 2, 2, 2)
    )
  
  if (!show_x) {
    map_plot <- map_plot + theme(
      axis.title.x = element_text(colour = NA),
      axis.text.x  = element_text(colour = NA),
      axis.ticks.x = element_line(colour = NA)
    )
  }
  
  if (!show_y) {
    map_plot <- map_plot + theme(
      axis.title.y = element_text(colour = NA),
      axis.text.y  = element_text(colour = NA),
      axis.ticks.y = element_line(colour = NA)
    )
  }
  
if (!is.null(indicator_label)) {
  map_plot <- map_plot +
    labs(title = indicator_label) +
    theme(
      plot.title = element_text(
        face = "bold",
        size = 14,
        hjust = 0
      )
    )
}
  
  map_plot
}
```

```{r}
# =========================================================
# SCENARIO FIGURE - MAPS ONLY
# =========================================================
plot_grouped_scenario_maps_only <- function(df_change, indicators,
                                            out_dir, filename,
                                            width = 12, height = 9) {
  
  df_grp <- df_change %>%
    filter(indicator %in% indicators) %>%
    mutate(indicator = factor(indicator, levels = indicators))
  
  scen_levels <- c("CC", "OWF", "OWF + CC")
  
  make_header_cell <- function(label) {
    ggplot() +
      annotate("text", x = 0.5, y = 0.5, label = label,
               fontface = "bold", size = 6) +
      xlim(0, 1) + ylim(0, 1) +
      theme_void()
  }
  
  # -------------------------
  # top header row
  # -------------------------
  plot_list <- list(
    make_header_cell("CC"),
    make_header_cell("OWF"),
    make_header_cell("OWF + CC")
  )
  
  # -------------------------
  # rows per indicator
  # -------------------------
for (i in seq_along(indicators)) {
  ind <- indicators[i]
  df_ind <- df_grp %>% filter(indicator == ind)
  
  for (j in seq_along(scen_levels)) {
    sc <- scen_levels[j]
    df_panel <- df_ind %>% filter(scenario == sc)
    
    p <- make_scenario_indicator_map(
      df_panel = df_panel,
      show_x = i == length(indicators),
      show_y = j == 1,
      indicator_label = if (j == 1) indicator_label_map[[as.character(ind)]] else NULL
    )
    
    plot_list <- c(plot_list, list(p))
  }
}
  
  final_plot <- wrap_plots(
    plotlist = plot_list,
    ncol = 3,
    byrow = TRUE
  ) +
    plot_layout(
      widths = c(1, 1, 1),
      heights = c(0.18, rep(1, length(indicators)))
    )
  
  ggsave(
    filename = file.path(out_dir, filename),
    plot = final_plot,
    width = width,
    height = height,
    dpi = 800
  )
  
  final_plot
}
```

```{r}
scenario_maps_only <- plot_grouped_scenario_maps_only(
  df_change = scenario_change_df,
  indicators = indicators,
  out_dir = out_dir,
  filename = "Scenario_ENA_indicators_maps_only.png",
  width = 14,
  height = 14
)

scenario_maps_only
```


```{r}
# =========================================================
# SUMMARY TABLES FOR SELECTED ENA INDICATORS
# Same structure as for biomass / catch / TL indicators
# =========================================================

library(dplyr)
library(tidyr)
library(readr)

# ---------------------------------------------------------
# 1. Prepare ENA data with clean indicator labels
# ---------------------------------------------------------
ena_table_df <- ena_df %>%
  mutate(
    indicator = recode(
      as.character(indicator),
      "FCI" = "FCI",
      "H" = "H",
      "RDC_internal_CAP" = "RDC",
      "SOI" = "SOI"
    )
  )

# ---------------------------------------------------------
# 2. Attach evaluation zones to ALL scenarios, including Baseline
# ---------------------------------------------------------
ena_table_zones <- ena_table_df %>%
  left_join(zone_df, by = "cell") %>%
  mutate(
    zone = case_when(
      zone == "OWF" ~ "OWF",
      zone == "Spillover" ~ "SPILL",
      zone == "Outside OWF" ~ "OUT",
      TRUE ~ NA_character_
    )
  )

# ---------------------------------------------------------
# 3. Mean absolute ENA value per indicator × scenario × zone
#    Include:
#    - ALL = whole study area
#    - OWF / SPILL / OUT = zonal subsets
# ---------------------------------------------------------
ena_summary_all <- ena_table_zones %>%
  group_by(indicator, scenario) %>%
  summarise(
    abs_value = mean(value, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(zone = "ALL")

ena_summary_zones <- ena_table_zones %>%
  filter(!is.na(zone)) %>%
  group_by(indicator, scenario, zone) %>%
  summarise(
    abs_value = mean(value, na.rm = TRUE),
    .groups = "drop"
  )

ena_summary_abs <- bind_rows(ena_summary_all, ena_summary_zones)

# ---------------------------------------------------------
# 4. Baseline zonal reference
# ---------------------------------------------------------
ena_baseline_ref <- ena_summary_abs %>%
  filter(scenario == "Baseline") %>%
  select(
    indicator,
    zone,
    baseline_abs_value = abs_value
  )

# ---------------------------------------------------------
# 5. Scenario summaries relative to baseline of the SAME zone
# ---------------------------------------------------------
ena_summary_vs_baseline <- ena_summary_abs %>%
  filter(scenario != "Baseline") %>%
  left_join(ena_baseline_ref, by = c("indicator", "zone")) %>%
  mutate(
    abs_change = abs_value - baseline_abs_value,
    pct_change = case_when(
      baseline_abs_value == 0 & abs_value == 0 ~ 0,
      baseline_abs_value == 0 & abs_value != 0 ~ NA_real_,
      is.na(baseline_abs_value) ~ NA_real_,
      TRUE ~ ((abs_value - baseline_abs_value) / baseline_abs_value) * 100
    )
  ) %>%
  mutate(
    indicator = factor(indicator, levels = c("FCI", "H", "RDC", "SOI")),
    scenario = factor(scenario, levels = c("CC", "OWF", "OWF + CC")),
    zone = factor(zone, levels = c("ALL", "OWF", "SPILL", "OUT"))
  ) %>%
  arrange(indicator, scenario, zone)
```


```{r}
# ---------------------------------------------------------
# 6. Formatting helpers
# ---------------------------------------------------------
format_abs_cell <- function(x, digits = 3) {
  ifelse(
    is.na(x),
    "NA",
    formatC(x, format = "f", digits = digits)
  )
}

format_pct_cell <- function(x, digits = 1) {
  ifelse(
    is.na(x),
    "NA",
    paste0(
      ifelse(x > 0, "+", ""),
      formatC(x, format = "f", digits = digits),
      "%"
    )
  )
}

format_abs_change_cell <- function(x, digits = 3) {
  ifelse(
    is.na(x),
    "NA",
    paste0(
      ifelse(x > 0, "+", ""),
      formatC(x, format = "f", digits = digits)
    )
  )
}

format_abs_and_pct_cell <- function(abs_change, pct_change,
                                    abs_digits = 3,
                                    pct_digits = 1) {
  
  abs_txt <- ifelse(
    is.na(abs_change),
    "NA",
    paste0(
      ifelse(abs_change > 0, "+", ""),
      formatC(abs_change, format = "f", digits = abs_digits)
    )
  )
  
  pct_txt <- ifelse(
    is.na(pct_change),
    "NA",
    paste0(
      ifelse(pct_change > 0, "+", ""),
      formatC(pct_change, format = "f", digits = pct_digits),
      "%"
    )
  )
  
  paste0(abs_txt, " (", pct_txt, ")")
}

format_abs_and_rel_cell <- function(abs_value, pct_change,
                                    abs_digits = 3,
                                    pct_digits = 1) {
  
  abs_txt <- ifelse(
    is.na(abs_value),
    "NA",
    formatC(abs_value, format = "f", digits = abs_digits)
  )
  
  pct_txt <- ifelse(
    is.na(pct_change),
    "NA",
    paste0(
      ifelse(pct_change > 0, "+", ""),
      formatC(pct_change, format = "f", digits = pct_digits),
      "%"
    )
  )
  
  paste0(abs_txt, " (", pct_txt, ")")
}
```


```{r}
# ---------------------------------------------------------
# 7. TABLE A: absolute value + relative change
#    Example: 0.345 (+5.6%)
# ---------------------------------------------------------
ena_table_abs_and_rel <- ena_summary_vs_baseline %>%
  mutate(
    cell = format_abs_and_rel_cell(abs_value, pct_change)
  ) %>%
  unite(col_name, scenario, zone, sep = "_") %>%
  select(indicator, col_name, cell) %>%
  pivot_wider(names_from = col_name, values_from = cell) %>%
  select(
    indicator,
    CC_ALL, CC_OWF, CC_SPILL, CC_OUT,
    OWF_ALL, OWF_OWF, OWF_SPILL, OWF_OUT,
    `OWF + CC_ALL`, `OWF + CC_OWF`, `OWF + CC_SPILL`, `OWF + CC_OUT`
  ) %>%
  mutate(indicator = as.character(indicator))

ena_table_abs_and_rel

# ---------------------------------------------------------
# 8. TABLE B: only relative changes (%)
# ---------------------------------------------------------
ena_table_pct <- ena_summary_vs_baseline %>%
  mutate(
    cell = format_pct_cell(pct_change)
  ) %>%
  unite(col_name, scenario, zone, sep = "_") %>%
  select(indicator, col_name, cell) %>%
  pivot_wider(names_from = col_name, values_from = cell) %>%
  select(
    indicator,
    CC_ALL, CC_OWF, CC_SPILL, CC_OUT,
    OWF_ALL, OWF_OWF, OWF_SPILL, OWF_OUT,
    `OWF + CC_ALL`, `OWF + CC_OWF`, `OWF + CC_SPILL`, `OWF + CC_OUT`
  ) %>%
  mutate(indicator = as.character(indicator))

ena_table_pct

# ---------------------------------------------------------
# 9. TABLE C: only absolute changes (+/- X)
# ---------------------------------------------------------
ena_table_abs_change <- ena_summary_vs_baseline %>%
  mutate(
    cell = format_abs_change_cell(abs_change)
  ) %>%
  unite(col_name, scenario, zone, sep = "_") %>%
  select(indicator, col_name, cell) %>%
  pivot_wider(names_from = col_name, values_from = cell) %>%
  select(
    indicator,
    CC_ALL, CC_OWF, CC_SPILL, CC_OUT,
    OWF_ALL, OWF_OWF, OWF_SPILL, OWF_OUT,
    `OWF + CC_ALL`, `OWF + CC_OWF`, `OWF + CC_SPILL`, `OWF + CC_OUT`
  ) %>%
  mutate(indicator = as.character(indicator))

ena_table_abs_change

# ---------------------------------------------------------
# 10. TABLE D: absolute change + relative change
#     Example: +0.012 (+5.6%)
# ---------------------------------------------------------
ena_table_abs_change_and_pct <- ena_summary_vs_baseline %>%
  mutate(
    cell = format_abs_and_pct_cell(abs_change, pct_change)
  ) %>%
  unite(col_name, scenario, zone, sep = "_") %>%
  select(indicator, col_name, cell) %>%
  pivot_wider(names_from = col_name, values_from = cell) %>%
  select(
    indicator,
    CC_ALL, CC_OWF, CC_SPILL, CC_OUT,
    OWF_ALL, OWF_OWF, OWF_SPILL, OWF_OUT,
    `OWF + CC_ALL`, `OWF + CC_OWF`, `OWF + CC_SPILL`, `OWF + CC_OUT`
  ) %>%
  mutate(indicator = as.character(indicator))

ena_table_abs_change_and_pct

# ---------------------------------------------------------
# 11. OPTIONAL: numeric export table
# ---------------------------------------------------------
ena_table_numeric <- ena_summary_vs_baseline %>%
  pivot_wider(
    names_from = c(scenario, zone),
    values_from = c(abs_value, abs_change, pct_change),
    names_glue = "{scenario}_{zone}_{.value}"
  ) %>%
  arrange(indicator)

ena_table_numeric
```


```{r}
# ---------------------------------------------------------
# 12. EXPORT
# ---------------------------------------------------------
write_csv(
  ena_table_abs_and_rel,
  file.path("ENA_summary_table_abs_and_rel.csv")
)

write_csv(
  ena_table_pct,
  file.path("ENA_summary_table_pct_only.csv")
)

write_csv(
  ena_table_abs_change,
  file.path("ENA_summary_table_abs_change_only.csv")
)

write_csv(
  ena_table_abs_change_and_pct,
  file.path("ENA_summary_table_abs_change_and_pct.csv")
)

write_csv(
  ena_table_numeric,
  file.path("ENA_summary_table_numeric.csv")
)
```

