---
title: "Ecospace_Basic_Maps_Plots"
author: "Dries Lorré"
date: "2026-03-19"
output: html_document
---

In this script, we PLOT the basic maps used for ecospace initiation. Initial processing of these data is covered in the RStudio script "Input_layers_processing.rdm" under project "Spatial_Input_layers_processing"

Packages
```{r}
library(terra)
library(sf)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(scales)
library(grid)
```

Directories
```{r}
# HABITATS
hab_coarse <- "C:/Users/dlorr/OneDrive - UGent/Documents/PhD/EwE_OWF/Spatial data/Processed/Folk5_CoarseSed.asc"
hab_mixed  <- "C:/Users/dlorr/OneDrive - UGent/Documents/PhD/EwE_OWF/Spatial data/Processed/Folk5_MixedSed.asc"
hab_mud    <- "C:/Users/dlorr/OneDrive - UGent/Documents/PhD/EwE_OWF/Spatial data/Processed/Folk5_Mud_MuddySand.asc"
hab_rock   <- "C:/Users/dlorr/OneDrive - UGent/Documents/PhD/EwE_OWF/Spatial data/Processed/Folk5_RockBoulders.asc"
hab_sand   <- "C:/Users/dlorr/OneDrive - UGent/Documents/PhD/EwE_OWF/Spatial data/Processed/Folk5_Sand.asc"

# DISTANCE TO LAND
dist_file <- "C:/Users/dlorr/OneDrive - UGent/Documents/PhD/EwE_OWF/Spatial data/Processed/DistCoast_km.asc"

# BATHYMETRY
depth_file <- "C:/Users/dlorr/OneDrive - UGent/Documents/PhD/EwE_OWF/Spatial data/Processed/Depth.asc"

# COASTAL - MARINE
coastmarine_file <- "C:/Users/dlorr/OneDrive - UGent/Documents/PhD/EwE_OWF/Spatial data/Processed/CoastalMarine_Seabirds.asc"

# RELATIVE PRIMARY PRODUCTION
rpp_file <- "C:/Users/dlorr/OneDrive - UGent/Documents/PhD/EwE_OWF/Spatial data/Processed/RPP_relative_yearly/RPP_rel_1991.asc"

# TEMPERATURE
temp_file <- "C:/Users/dlorr/OneDrive - UGent/Documents/PhD/EwE_OWF/Spatial data/Processed/Temperature/Baseline/Temp_Baseline_1991.asc"

# SALINITY
sal_file <- "C:/Users/dlorr/OneDrive - UGent/Documents/PhD/EwE_OWF/Spatial data/Processed/Salinity/Baseline/Sal_Baseline_1991.asc"

# LAND SHAPEFILE / RDS
# Use whichever one you prefer
land_file_rds <- "land_main_study_area.rds"
# land_file_shp <- "LAND_MASS.shp"

# OUTPUT DIRECTORY
out_dir <- "Plots_basic_input_layers"
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

```

Reference raster and land polygon
```{r}
ref_rast <- rast(depth_file)

land_main <- readRDS(land_file_rds)

# If needed, reproject land to raster CRS
if (!st_crs(land_main)$wkt == crs(ref_rast)) {
  land_main <- st_transform(land_main, crs(ref_rast))
}
```


Helper functions
```{r}
raster_to_df <- function(r, varname = "value", zero_to_na = TRUE) {
  df <- as.data.frame(r, xy = TRUE, na.rm = FALSE)
  names(df) <- c("x", "y", varname)

  if (zero_to_na) {
    df[[varname]][df[[varname]] == 0] <- NA
  }

  df
}

base_map_theme <- function() {
  theme_bw(base_size = 11) +
    theme(
      axis.title = element_text(size = 14),
      axis.text = element_text(size = 14),
      legend.title = element_text(size = 13),
      legend.text = element_text(size = 13),
      panel.grid = element_blank(),
      panel.background = element_rect(fill = "white", colour = NA),
      legend.position = "right",

      # optional cleanup
      # legend.key = element_rect(fill = "white", colour = NA),

      plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
      plot.margin = margin(2, 2, 2, 2)
    )
}

add_common_map_layout <- function(p, df) {
  p +
    geom_sf(
      data = land_main,
      inherit.aes = FALSE,
      fill = "grey85",
      colour = "grey45",
      linewidth = 0.2
    ) +
    coord_sf(expand = FALSE) +
    scale_x_continuous(
      breaks = pretty(range(df$x, na.rm = TRUE), n = 3),
      name = "Longitude"
    ) +
    scale_y_continuous(
      breaks = pretty(range(df$y, na.rm = TRUE), n = 3),
      name = "Latitude"
    ) +
    base_map_theme()
}

make_continuous_map <- function(df, value_col, title, colours, legend_title = NULL, labels_fun = waiver()) {
  
  panel_min <- quantile(df[[value_col]], probs = 0.001, na.rm = TRUE)
  panel_max <- quantile(df[[value_col]], probs = 0.999, na.rm = TRUE)
  
  if (!is.finite(panel_min)) panel_min <- min(df[[value_col]], na.rm = TRUE)
  if (!is.finite(panel_max)) panel_max <- max(df[[value_col]], na.rm = TRUE)
  if (!is.finite(panel_min)) panel_min <- 0
  if (!is.finite(panel_max)) panel_max <- 1
  
  if (panel_min == panel_max) panel_max <- panel_min + 1e-6
  
  p <- ggplot(df, aes(x = x, y = y, fill = .data[[value_col]])) +
    geom_raster() +
    scale_fill_gradientn(
      colours = colours,
      limits = c(panel_min, panel_max),
      oob = scales::squish,
      breaks = pretty(c(panel_min, panel_max), n = 3),
      labels = labels_fun,
      na.value = "white",
      name = legend_title,
      guide = guide_colorbar(
        title.position = "top",
        barheight = unit(35, "mm"),
        barwidth = unit(6, "mm"),
        frame.colour = "black",
        ticks.colour = "black"
      )
    ) +
    labs(title = title)
  
  add_common_map_layout(p, df)
}

make_categorical_map <- function(df, value_col, title, values_named) {
  
  p <- ggplot(df, aes(x = x, y = y, fill = .data[[value_col]])) +
    geom_raster() +
    scale_fill_manual(
      values = values_named,
      na.value = "white",
      na.translate = FALSE,
      guide = guide_legend(
        title.position = "top",
        keyheight = unit(6, "mm"),
        keywidth = unit(6, "mm")
      )
    ) +
    labs(title = title)
  
  add_common_map_layout(p, df)
}

```

Load rasters
```{r}
# Habitat rasters
r_coarse <- rast(hab_coarse)
r_mixed  <- rast(hab_mixed)
r_mud    <- rast(hab_mud)
r_rock   <- rast(hab_rock)
r_sand   <- rast(hab_sand)
# Other
r_dist        <- rast(dist_file)
r_depth       <- rast(depth_file)
r_coastmarine <- rast(coastmarine_file)
r_rpp         <- rast(rpp_file)
r_temp        <- rast(temp_file)
r_sal         <- rast(sal_file)
```

Habitat map
```{r}
hab_stack <- c(r_coarse, r_mixed, r_mud, r_rock, r_sand)
names(hab_stack) <- c(
  "Coarse sediment",
  "Mixed sediment",
  "Mud / muddy sand",
  "Rock / boulders",
  "Sand"
)

hab_df <- as.data.frame(hab_stack, xy = TRUE, na.rm = FALSE)

hab_cols <- c(
  "Coarse sediment",
  "Mixed sediment",
  "Mud / muddy sand",
  "Rock / boulders",
  "Sand"
)

# classify strictly on presence == 1
hab_df$habitat <- NA_character_

hab_df$habitat[hab_df$`Coarse sediment` == 1] <- "Coarse sediment"
hab_df$habitat[hab_df$`Mixed sediment`  == 1] <- "Mixed sediment"
hab_df$habitat[hab_df$`Mud / muddy sand` == 1] <- "Mud / muddy sand"
hab_df$habitat[hab_df$`Rock / boulders` == 1] <- "Rock / boulders"
hab_df$habitat[hab_df$`Sand` == 1] <- "Sand"

hab_df$habitat <- factor(hab_df$habitat, levels = hab_cols)

habitat_cols <- c(
  "Coarse sediment" = "#b2182b",
  "Mixed sediment" = "#ef8a62",
  "Mud / muddy sand" = "#99804d",
  "Rock / boulders" = "#636363",
  "Sand" = "#fee08b"
)

p_habitat <- make_categorical_map(
  df = hab_df[, c("x", "y", "habitat")],
  value_col = "habitat",
  title = "Habitat types",
  values_named = habitat_cols
) 

p_habitat

```

Distance to land map
```{r}
dist_df <- raster_to_df(r_dist, "dist_km")

p_dist <- make_continuous_map(
  df = dist_df,
  value_col = "dist_km",
  title = "Distance to land",
  colours = c("khaki1", "goldenrod1", "orange", "tomato", "firebrick"),
  legend_title = "[km]"
)

p_dist

# --> add legend text "[km]"
```

Bathymetry map
```{r}
depth_df <- raster_to_df(r_depth, "depth")

p_depth <- make_continuous_map(
  df = depth_df,
  value_col = "depth",
  title = "Bathymetry",
  colours = c("white", "lightcyan", "lightskyblue", "deepskyblue3", "steelblue4", "midnightblue"),
  legend_title = "[m]"
)

p_depth

# --> add legend text "[m]"
```

Coastal - Marine map
```{r}
coastmarine_df <- raster_to_df(r_coastmarine, "zone") %>%
  mutate(
    zone_class = case_when(
      is.na(zone) ~ NA_character_,
      zone == 1 ~ "Coastal",
      zone == 0.75 ~ "Marine",
      TRUE ~ as.character(zone)
    )
  )

coastmarine_df$zone_class <- factor(coastmarine_df$zone_class, levels = c("Coastal", "Marine"))

p_coastmarine <- make_categorical_map(
  df = coastmarine_df %>% select(x, y, zone_class),
  value_col = "zone_class",
  title = "Coastal–marine classification",
  values_named = c(
    "Coastal" = "#fee07b",
    "Marine" = "#888eb8"
  )
) + theme(legend.title = element_blank())

p_coastmarine 
```

Relative primary production map
```{r}
rpp_df <- raster_to_df(r_rpp, "rpp")

p_rpp <- make_continuous_map(
  df = rpp_df,
  value_col = "rpp",
  title = "Relative primary production",
  colours = c("white", "palegreen", "yellowgreen", "forestgreen", "darkgreen")
)

p_rpp
```

Temperature map 
```{r}
temp_df <- raster_to_df(r_temp, "temp")

p_temp <- make_continuous_map(
  df = temp_df,
  value_col = "temp",
  title = "Sea Surface Temperature (1991 - average)",
  colours = c("navy", "dodgerblue3", "deepskyblue", "gold", "orange", "red3"),
  legend_title = "[°C]"
)

p_temp

# --> add legend text "[°C]"
```

Salinity map
```{r}
sal_df <- raster_to_df(r_sal, "sal")

p_sal <- make_continuous_map(
  df = sal_df,
  value_col = "sal",
  title = "Sea Surface Salinity (1991 - average)",
  colours = c("lightyellow", "khaki1", "gold", "orange", "brown4"),
  legend_title = "[psu]"
)

p_sal

# --> add legend text "[psu]"
```

Overview plot
```{r}
overview_plot <- (
  p_habitat + p_dist +
  p_depth + p_coastmarine +
  p_rpp + p_temp +
  p_sal + plot_spacer()
) +
  plot_layout(ncol = 2) +
  plot_annotation(
    title = NULL,
    theme = theme(
      plot.title = element_text(face = "bold", size = 15, hjust = 0.5)
    )
  )

overview_plot

# Save overview figure
ggsave(
  file.path(out_dir, "Overview_basic_Ecospace_input_layers.png"),
  overview_plot,
  width = 14,
  height = 17,
  dpi = 800
)
```

